---
title: "AFSC Case Study: Converting a Bering Sea Alaska Plaice SS3 Model to FIMS"
stock: "AFSC Alaska Plaice"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
---

## Introduction

This document serves as a case study and guide for converting a stock assessment model from SS3 to FIMS framework. We use the Bering Sea and Aleutian Islands (BSAI) Alaska Plaice assessment in this example.

The goal is to demonstrate a reproducible workflow, highlight common challenges, and provide solutions for simplifying a complex SS3 model to a structure compatible with FIMS. This process often involves making simplifying assumptions, and we will walk through these decisions.

We will cover two main pathways for conversion, based on how the model handles growth:

1.  **Using Growth Parameters**: The model estimates growth using a von Bertalanffy growth function, and weight is derived from a length-weight relationship.
2.  **Using Empirical Weight-at-Age (EWAA)**: The model uses observed mean weight-at-age data directly, bypassing the estimation of growth parameters.

For each pathway, we will follow these general steps:
1.  Simplify the original SS3 data and control files.
2.  Run the simplified SS3 model to generate necessary output files.
3.  Convert the simplified SS3 output into a FIMS data object.
4.  Set up and run the corresponding FIMS model.
5.  Compare the results between the original SS3 model, the simplified SS3 model, and the FIMS model.

## The Setup

This section loads necessary libraries and defines some common variables that will be used throughout the analysis.

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Alaska plaice"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

**A Note on This Demonstration:** The models presented in this case study are changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration of the FIMS conversion process and nothing more.

## Part 1: Preparing a Simplified SS3 Model for FIMS Conversion

FIMS is a powerful but newer framework and does not yet have all the features of SS3 (e.g., two-sex models, certain selectivity patterns). Therefore, a critical first step is often to simplify the existing SS3 model to a "FIMS-compatible" structure. This also helps in making a more direct, "apples-to-apples" comparison between the two modeling platforms.

The original Alaska plaice model is a two-sex model. As FIMS currently only supports single-sex (or combined-sex) models, our primary simplification will be to convert the model to a female-only model.

To keep our code organized and reduce repetition, we'll create a helper function to perform these simplifications below.

```{r}
#| label: simplification-function

#' Simplify SS3 data and control files for FIMS conversion
#'
#' This function takes an SS3 input list (from r4ss::SS_read) and applies
#' a series of simplifications to make it more compatible with FIMS and allow
#' for a cleaner comparison.
#'
#' @param input_list An SS3 input list from r4ss::SS_read.
#' @param use_ewaa A logical flag. If TRUE, the model will be configured to use
#'   Empirical Weight-at-Age and length comps will be turned off.
#' @param output_dir The directory to write the new model files to.
#'
#' @return The modified input_list.

simplify_ss3_for_fims <- function(input_list, use_ewaa, output_dir) {
  
  dat <- input_list$dat
  ctl <- input_list$ctl
  fore <- input_list$fore
  
  # --- DATA FILE SIMPLIFICATIONS ---
  # Convert to a single-sex model (female-only)
  dat$Nsexes <- -1
  
  # Remove male age compositions
  dat$agecomp <- dat$agecomp |>
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    dplyr::mutate(sex = -1)
  
  if (use_ewaa) {
    # If using EWAA, we don't use length data
    dat$use_lencomp <- 0
  } else {
    # If not using EWAA, remove male length compositions
    dat$lencomp <- dat$lencomp |>
      dplyr::select(!dplyr::matches("^m\\d+$")) |>
      dplyr::mutate(sex = -1)
    ctl$EmpiricalWAA <- 0
  }
  
  # --- CONTROL FILE SIMPLIFICATIONS ---
  ctl$Nsexes <- dat$Nsexes
  
  # Remove all male-specific mortality and growth parameters
  ctl$MG_parms <- ctl$MG_parms |>
    dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))
  
  # Change selectivity to logistic (FIMS option) instead of double-normal.
  # Set Male to 0 as we no longer have a male-specific curve.
  ctl$age_selex_types <- ctl$age_selex_types |>
    dplyr::mutate(
      Pattern = 12, # 12 = logistic
      Male = 0
    )
  
  # Replace double-normal selectivity parameters with logistic parameters.
  # Remove male offset parameters.
  ctl$age_selex_parms <- cbind(
    data.frame(
      LO = 1, 
      HI = 15, 
      INIT = c(10, 2, 7, 2), 
      PRIOR = 0,
      PR_SD = 99, 
      PR_type = 0, 
      PHASE = 2
    ),
    matrix(0, nrow = 4, ncol = 7) # Fills remaining columns with 0
  )
  names(ctl$age_selex_parms) <- names(ctl$MG_parms)
  rownames(ctl$age_selex_parms) <- paste0(
    "AgeSel_", 
    c("P_1_", "P_2_"), 
    c("Fishery", "Fishery", "Survey", "Survey")
  )
  
  # Remove blocks and variance adjustments for simplicity
  ctl$N_Block_Designs <- 0
  ctl$blocks_per_pattern <- NULL
  ctl$Block_Design <- NULL
  ctl$Variance_adjustment_list <- NULL
  ctl$DoVar_adjust <- 0
  ctl$sd_offset <- 0
  
  # Adjust recruitment deviation years
  ctl$MainRdevYrFirst <- dat$styr
  ctl$MainRdevYrLast <- dat$endyr
  ctl$recdev_early_start <- dat$styr - dat$Nages
  
  # --- FORECAST FILE ---
  # Turn off forecasting for this comparison
  fore$Forecast <- 0
  
  # --- WRITE MODIFIED FILES ---
  new_input <- input_list
  new_input$dat <- dat
  new_input$ctl <- ctl
  new_input$fore <- fore
  if (use_ewaa) {
    new_input$wtatage <- input_list$wtatage
  }
  
  # Name and create output directory
  new_input$dir <- output_dir
  if (!dir.exists(new_input$dir)) {
    dir.create(new_input$dir)
  }
  
  # Write to output directory
  r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)
  
  return(new_input)
}
```

Now we can use this function to create our two simplified SS3 models.

### Scenario 1: Simplified Model using Growth Parameters (No EWAA)

Here, we simplify the SS3 model and configure it to estimate growth internally rather than using empirical weight-at-age.

```{r}
library(FIMS)
#| label: prep-no-ewaa-model
# Read the original assessment model from GitHub
plaice_input_base <- r4ss::SS_read(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    read_wtatage = TRUE, # Read wtatage file even if not used
    ss_new = TRUE
)

# Create and write the simplified model files
simplified_no_ewaa <- simplify_ss3_for_fims(
  input_list = plaice_input_base,
  use_ewaa = FALSE,
  output_dir = "plaice_model_noEWAA_FIMS"
)

# Run the simplified SS3 model
# For the converter to work robustly, especially when length comps are present,
# it's best to run the model to generate output files like Report.sso.
r4ss::get_ss3_exe(dir = simplified_no_ewaa$dir)
r4ss::run(
    dir = simplified_no_ewaa$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```

### Scenario 2: Simplified Model using Empirical Weight-at-Age (EWAA)

In this scenario, we simplify the model but configure it to use the empirical weight-at-age data directly. This means we will turn off the length composition data.

```{r}
#| label: prep-ewaa-model

# We can re-use the plaice_input_base object we read in the previous chunk.
# Create and write the simplified model files for the EWAA case.
simplified_ewaa <- simplify_ss3_for_fims(
  input_list = plaice_input_base,
  use_ewaa = TRUE,
  output_dir = "plaice_model_EWAA_FIMS"
)

# Run the simplified SS3 model
# When using EWAA, the ss3_data_to_fims converter requires the Report.sso file
# to correctly process the weight-at-age information.
r4ss::get_ss3_exe(dir = simplified_ewaa$dir)
r4ss::run(
    dir = simplified_ewaa$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```

## Part 2: Converting the Simplified SS3 Model to FIMS

With our simplified and successfully run SS3 models in hand, we can now use `r4ss::ss3_data_to_fims` to convert them into FIMS data objects.

For the remainder of this walkthrough, we will proceed with the **"No EWAA"** and **"EWAA"** models to demonstrate the FIMS setup process. The process for the EWAA model has the main difference being that growth parameters don't need to be specified in the FIMS model.

```{r}
#| label: convert-to-fims-no-ewaa
#| warning: false

# Convert the simplified "No EWAA" SS3 model output into the FIMS format
# and then create a FIMSFrame object.
data_4_fims <- r4ss::ss3_data_to_fims(
    ss3_dir = simplified_no_ewaa$dir,
    ss_new = TRUE
) |> FIMS::FIMSFrame()

# To proceed, we also need to read the output of our simplified SS3 run
# to get parameter estimates for starting values in FIMS.
rep_simple_ss3 <- r4ss::SS_output(
    simplified_no_ewaa$dir,
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)
```


```{r}
#| label: convert-to-fims-ewaa
#| warning: false
data_4_fims_ewaa <- r4ss::ss3_data_to_fims(
    ss3_dir = simplified_ewaa$dir,
    ss_new = TRUE
) |>
    FIMS::FIMSFrame()

# To proceed, we also need to read the output of our simplified SS3 run
# to get parameter estimates for starting values in FIMS.
rep_ewaa_ss3 <- r4ss::SS_output(
    simplified_ewaa$dir,
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)
```

## Part 3: Building and Running the FIMS Model

Now we have the data for FIMS in the `data_4_fims` object. The next step is to define the model structure and provide initial parameter values.

### Setting Up Parameters

We start by creating a data frame of default parameters based on our data. We then modify this data frame with better starting values, often derived from the model we are converting from (our simplified SS3 model).

```{r}
#| label: setup-fims-params
#| max.height: '100px'
#| attr.output: '.numberLines'

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
  data = data_4_fims
)

# View fixed params (constants) that we NEED to make sure are right for our species
default_parameters |>
  dplyr::filter(estimation_type == "constant") |>
  dplyr::count(module_name, fleet, label, value)

# View estimated params
default_parameters |>
  dplyr::filter(estimation_type == "fixed_effects") |>
  dplyr::count(module_name, fleet, label, value)

# Get number of years and ages from the data to make sense out of parameter counts
n_years <- FIMS::get_n_years(data_4_fims)
n_ages <- FIMS::get_ages(data_4_fims)
```

### Populating Parameters from the Simplified SS3 Model

This is a critical step. Good starting values are essential for model convergence. We will (hopefully) eventually create a helper function to make it easier to pull values from the SS3 report object that is already in the `{r4ss}` package but for now we will create a helper function to help us get these values.

```{r}
#| label: get-ss3-param-helper

# Helper function to get an estimated parameter value from the SS3 report object
get_ss3_estimate <- function(string, report = rep_simple_ss3) {
    report$parameters |>
        dplyr::filter(grepl(
            pattern = string,
            Label,
            fixed = TRUE
        )) |>
        dplyr::pull(Value)
}
```


In FIMS, initial numbers-at-age (NAA) are a free parameter for every age. In SS3, they are derived from R0, M and the early recruitment deviations. Estimating all 26 of them freely does not work for plaice. The fish that are 17 and older in 1975 have died out before most of the composition data begin, so nothing in the data informs them. They run off toward zero, the Hessian becomes singular, and every standard error is `NaN`. For this bridging exercise we instead fix the initial NAA at the start-year numbers from the simplified SS3 run. FIMS does not yet have an equilibrium-based option for the initial age structure.

```{r}
#| label: naa
log_m_init <- log(get_ss3_estimate("NatM_uniform_Fem_GP_1"))
log_rzero_init <- get_ss3_estimate("SR_LN(R0)")
# Beginning-of-year numbers-at-age (1000s) in the first model year
start_year_naa <- rep_simple_ss3$natage |>
    dplyr::filter(Yr == min(Yr[Era == "TIME"]), `Beg/Mid` == "B") |>
    dplyr::select(dplyr::all_of(as.character(FIMS::get_ages(data_4_fims))))
init_naa <- log(unlist(start_year_naa))
```

SS3 reports growth in two ways. The parameters (`L_a_A1`, `L_a_A2`, `CVmin`, `CVmax`) are not what FIMS expects:

- `A2` is 999, SS3's flag meaning `L_a_A2` is Linf rather than a length at a modeled age.
- `CVmin` and `CVmax` are coefficients of variation. FIMS wants the standard deviation of length-at-age in cm.

So we take beginning-of-year mean length and SD at ages 1 and 25 from `rep_simple_ss3$endgrowth`. SS3 interpolates the CV linearly in length, while FIMS interpolates the SD linearly in mean length, so SDs at intermediate ages differ somewhat. We use beginning-of-year values because FIMS uses one growth curve for both weight-at-age and the age-length key. That keeps spawning biomass comparable to SS3, whose length compositions are sampled at month 7.

```{r}
#| label: ss3-growth-at-reference-ages
growth_reference_ages <- c(1, 25)
growth_at_reference_ages <- rep_simple_ss3$endgrowth |>
    dplyr::filter(Age_Beg %in% growth_reference_ages)
length_at_reference_ages <- growth_at_reference_ages$Len_Beg
length_sd_at_reference_ages <- growth_at_reference_ages$SD_Beg

# SS3 reads maturity-at-age directly (maturity option 3). FIMS uses a
# logistic curve in age, so fit one to the SS3 vector.
ss3_maturity <- rep_simple_ss3$endgrowth |>
    dplyr::select(age = Age_Beg, maturity = Age_Mat)
maturity_fit <- stats::nls(
    maturity ~ 1 / (1 + exp(-slope * (age - inflection_point))),
    data = ss3_maturity,
    start = list(inflection_point = 9, slope = 1)
)
stats::coef(maturity_fit)
```


Now, let's create our final parameter set for FIMS by updating the defaults.

```{r}
# The default parameter tibble sets up EWAA growth, which does not require any parameters
default_parameters |> dplyr::filter(module_name == "Growth")

# This adds all the growth parameters that we will add later, you first need to replace the 
# EWAA row with parameters for a default growth model:
default_parameters <- default_parameters |>
      dplyr::filter(module_name != "Growth") |>
      dplyr::bind_rows(
        FIMS::setup_default_Growth(
          data = data_4_fims,
          module_type = "VonBertalanffySchnute"
        )
      )

# Now the EWAA row is replace with a set of 
# VonBertalanffySchnute row
default_parameters |> dplyr::filter(module_name == "Growth")
# `dplyr::rows_update()` requires a key column that is
# present in both the tibble being updated and the tibble 
# being used to update. The name of this column or 
# combination of columns is passed to the `by` argument.
```

```{r}
#| label: populate-fims-params
parameters <- default_parameters |>
    # Update Growth parameters. These are treated as fixed constants in this example,
    # using the estimates from the SS3 model. If using EWAA, this block would be removed.
    # If you want to look up what these parameters are, you can do the following:
    # default_parameters |> dplyr::filter(module_name == "Growth") |> dplyr::pull(label) |> unique()
    dplyr::rows_update(
        tibble::tibble(
        module_name = "Growth",
        label = c(
            "mean_length_young", "mean_length_old", "growth_coefficient",
            "reference_age_for_length_young", "reference_age_for_length_old",
            "length_weight_a", "length_weight_b"
        ),
        value = c(
            length_at_reference_ages,
            rep_simple_ss3$Growth_Parameters$K,
            growth_reference_ages,
            rep_simple_ss3$Growth_Parameters$WtLen1,
            rep_simple_ss3$Growth_Parameters$WtLen2
        ),
        estimation_type = "constant"
        ),
        by = c("module_name", "label")
    ) |>
    # change first reference age for sd values from 0 to 1
    dplyr::mutate(
    age = dplyr::if_else(
      module_name == "Growth" &
        label == "length_at_age_sd_at_reference_ages" &
        age == 0,
      1,
      age
    )) |>
    # Update SD of length-at-age (cm, not a CV) at the growth reference ages
    # If you want to look at what these parameters are, and how many there are, you can do the following:
    # default_parameters |> dplyr::filter(module_name == "Growth") |> dplyr::filter(label == "length_at_age_sd_at_reference_ages")
    dplyr::rows_update(
        tibble::tibble(
        module_name = "Growth",
        label = "length_at_age_sd_at_reference_ages",
        age = growth_reference_ages,
        value = length_sd_at_reference_ages
        ),
        by = c("module_name", "label", "age")
    ) |>
    # Update Maturity parameters from the logistic fit to the SS3 maturity-at-age
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Maturity",
            label = c("inflection_point", "slope"),
            value = unname(stats::coef(maturity_fit)[c("inflection_point", "slope")])
        ),
        by = c("module_name", "label")
    ) |>
    # Update Recruitment (Stock-Recruit Relationship) parameters
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                log_rzero_init,
                # Steepness (h) is a parameter bounded between 0.2 and 1.0.
                # Optimization algorithms work best on an unconstrained scale (-Inf to Inf).
                # The logit transform maps the value from its bounded range to an
                # unconstrained one. FIMS provides a helper for this common transformation.
                FIMS::logit(0.2, 1.0, 0.99),
                log(get_ss3_estimate("SR_sigmaR"))
            )
        ),
        by = c("module_name", "label")
    ) |>
    # Update survey catchability (log_q)
    dplyr::rows_update(
        tibble::tibble(
            fleet = "Survey",
            label = "log_q",
            value = get_ss3_estimate("LnQ_base_Survey(2)")
        ),
        by = c("fleet", "label")
    ) |>
    # Fix initial numbers-at-age (NAA) at the SS3 start-year values
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = n_ages,
            value = init_naa,
            estimation_type = "constant"
        ),
        by = c("module_name", "label", "age")
    ) |>
    # Update Natural Mortality (M)
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log_m_init,
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )
```

### Sanity Checks

Before running the model, it's good practice to check how our final parameter set (`parameters`) differs from the defaults. This helps catch mistakes.

```{r}
#| label: fims-param-checks
# This is a great practice to ensure your parameter setup is what you expect.
# Should this be a built-in FIMS function? Maybe! For now, this custom
# check is very helpful for debugging.

# Confirm that all identifying columns are identical
if (!identical(
    parameters |> dplyr::select(-age, -value, -estimation_type),
    default_parameters |> dplyr::select(-age, -value, -estimation_type)
)) {
    cli::cli_alert_danger("Parameter data frames are not structurally identical!")
}

# Show which parameters have been updated from their default values
cli::cli_alert_info("Rows with updated initial values from defaults:")
updated_parameter_counts <- parameters |>
    dplyr::filter(value != default_parameters$value) |>
    dplyr::select(module_name, fleet, label, value) |>
    dplyr::count(module_name, label)
print(updated_parameter_counts, n = Inf, width = Inf)
```

### Running the FIMS Model

Finally, we can run the model. It's often a good idea to first run without optimizing to check for any immediate errors in the model setup (`test_fit`). Then, run with optimization to get the final fit.

```{r}
#| label: run-fims-model
# Run the model without optimization to check for errors
test_fit <- parameters |>
    FIMS::initialize_fims(data = data_4_fims) |>
    FIMS::fit_fims(optimize = FALSE)

# need to fix error message to say "fit" or whatever instead of FIMSFit

# Run the model with optimization
fit <- parameters |>
    FIMS::initialize_fims(data = data_4_fims) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)

get_report(fit)$nll_components
```

The length compositions need a FIMS version that includes NOAA-FIMS/FIMS#1751. Earlier versions build expected length compositions from the age compositions after rescaling them to the observed age sample size. Most plaice years have length compositions but no age compositions, so the expected length compositions in those years were 0/0 = `NaN`, and `nlminb()` stopped with "NA/NaN function evaluation".

Setting `get_sd = TRUE` needs NOAA-FIMS/FIMS#1775. Without it, `TMB::sdreport()` builds a dense covariance matrix of all 46,384 derived quantities (about 17 GB), which exhausts memory.

### Running the EWAA FIMS Model

The EWAA model uses the same setup without growth parameters. One unit detail matters here:

- `ss3_data_to_fims()` divides SS3 weight-at-age by 1000, assuming kg to mt, so weights are per fish rather than per thousand fish.
- SS3 numbers are in thousands, so we multiply SS3's R0 and start-year numbers-at-age by 1000 to match, as in the petrale case study.
- Plaice weights in SS3 happen to be in grams, but only the product of numbers and weight has to match SS3's catch units. That product is unchanged.

The no-EWAA model avoids this by taking the SS3 length-weight parameters directly.

```{r}
#| label: run-fims-ewaa-model
get_ss3_ewaa_estimate <- function(string) {
    get_ss3_estimate(string, report = rep_ewaa_ss3)
}
ages_ewaa <- FIMS::get_ages(data_4_fims_ewaa)
start_year_naa_ewaa <- rep_ewaa_ss3$natage |>
    dplyr::filter(Yr == min(Yr[Era == "TIME"]), `Beg/Mid` == "B") |>
    dplyr::select(dplyr::all_of(as.character(ages_ewaa)))

parameters_ewaa <- FIMS::setup_default_parameters(data = data_4_fims_ewaa) |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Maturity",
            label = c("inflection_point", "slope"),
            value = unname(stats::coef(maturity_fit)[c("inflection_point", "slope")])
        ),
        by = c("module_name", "label")
    ) |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                # SS3 recruitment is in thousands of fish
                log(1000) + get_ss3_ewaa_estimate("SR_LN(R0)"),
                FIMS::logit(0.2, 1.0, 0.99),
                log(get_ss3_ewaa_estimate("SR_sigmaR"))
            )
        ),
        by = c("module_name", "label")
    ) |>
    dplyr::rows_update(
        tibble::tibble(
            fleet = "Survey",
            label = "log_q",
            value = get_ss3_ewaa_estimate("LnQ_base_Survey(2)")
        ),
        by = c("fleet", "label")
    ) |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = ages_ewaa,
            # SS3 numbers are in thousands of fish
            value = log(1000 * unlist(start_year_naa_ewaa)),
            estimation_type = "constant"
        ),
        by = c("module_name", "label", "age")
    ) |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log(get_ss3_ewaa_estimate("NatM_uniform_Fem_GP_1")),
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )

fit_ewaa <- parameters_ewaa |>
    FIMS::initialize_fims(data = data_4_fims_ewaa) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)

get_report(fit_ewaa)$nll_components
```

## Part 4: Comparing Model Outputs

A crucial final step is to compare the results. Here, we'll use `r4ss::SSplotComparisons` to visually compare the output of our original, complex SS3 model with our simplified SS3 model. This helps us understand the impact of the simplifications we made. A full comparison would also include plotting the FIMS results against the SS3 models.

```{r}
#| label: compare-models
# Read output from the original SS3 model on GitHub for comparison
plaice_output_original <- r4ss::SS_output(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

# Read output from our simplified SS3 model (already in rep_simple_ss3)

# Create a summary object for comparison plotting
summary_list <- r4ss::SSsummarize(list(
    "Original SS3" = plaice_output_original,
    "Simplified SS3" = rep_simple_ss3
))

# Plot the comparisons
r4ss::SSplotComparisons(summary_list)
```

The comparison plots show that removing the male data and changing the selectivity and growth options had a noticeable effect on the estimated spawning output, highlighting the importance of understanding the impact of such simplifying assumptions.

Next we compare each FIMS model with the simplified SS3 model it was converted from. Spawning biomass from both FIMS models is on the same scale as SS3 (see the unit note in the EWAA section). EWAA recruitment is in fish, so we divide it by 1000 to match SS3.

```{r}
#| label: compare-fims-ss3
years <- rep_simple_ss3$timeseries |>
    dplyr::filter(Era == "TIME") |>
    dplyr::pull(Yr)
n_years_plot <- length(years)

get_ss3_timeseries <- function(report, model) {
    report$timeseries |>
        dplyr::filter(Era == "TIME") |>
        dplyr::transmute(
            year = Yr,
            spawning_biomass = SpawnBio,
            recruitment = Recruit_0,
            model = model
        )
}
get_fims_timeseries <- function(fit, model, recruitment_scale = 1) {
    report <- get_report(fit)
    tibble::tibble(
        year = years,
        spawning_biomass = report$spawning_biomass[[1]][seq_len(n_years_plot)],
        recruitment = report$expected_recruitment[[1]][seq_len(n_years_plot)] /
            recruitment_scale,
        model = model
    )
}

comparison <- dplyr::bind_rows(
    get_ss3_timeseries(rep_simple_ss3, "SS3 no EWAA"),
    get_fims_timeseries(fit, "FIMS no EWAA"),
    get_ss3_timeseries(rep_ewaa_ss3, "SS3 EWAA"),
    get_fims_timeseries(fit_ewaa, "FIMS EWAA", recruitment_scale = 1000)
) |>
    tidyr::separate_wider_delim(model, delim = " ", names = c("platform", "scenario"), too_many = "merge") |>
    tidyr::pivot_longer(c(spawning_biomass, recruitment), names_to = "quantity")

ggplot2::ggplot(comparison, ggplot2::aes(year, value, color = platform)) +
    ggplot2::geom_line() +
    ggplot2::facet_grid(quantity ~ scenario, scales = "free_y") +
    ggplot2::labs(x = "Year", y = NULL, color = NULL)

# Ratio of FIMS to SS3 spawning biomass by scenario
comparison |>
    dplyr::filter(quantity == "spawning_biomass") |>
    tidyr::pivot_wider(names_from = platform, values_from = value) |>
    dplyr::mutate(ratio = FIMS / SS3) |>
    dplyr::group_by(scenario) |>
    dplyr::summarise(min_ratio = min(ratio), max_ratio = max(ratio))
```

## Appendix: Notes on Steepness

For Nanaimo-specific models, there is the spiny dogfish assessment (https://github.com/pbs-assess/dogfish-assess/tree/main/ss3/A1) and this other assessment that Ian recommended: https://github.com/pbs-software/pbs-synth/tree/master/PBSsynth/inst/input/2024/YTR/BC/Base/B1.R02v2/02.01`

For Beverton-Holt steepness, the value is often difficult to estimate within the model and is frequently fixed based on a meta-analysis or prior knowledge. The example code shows how you would insert a fixed value for steepness (e.g., `h = 0.283`) into the FIMS parameter setup.

```{r}
#| label: steepness-example
#| eval: false
#| echo: true
# Example of setting a fixed steepness value
h_fixed <- 0.283
parameters |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = "logit_steep",
            # Use the logit transform on the fixed value
            value = FIMS::logit(0.2, 1.0, h_fixed),
            # Set as a constant so it is not estimated
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )
```