---
title: AFSC Case Study for Bering Sea and Aleutian Islands Alaska Plaice
stock: "AFSC Alaska Plaice"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Alaska plaice"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model a simpler SS3 model was developed using code in https://github.com/NOAA-FIMS/case-studies/blob/test-plaice/content/R/AFSC-plaice_simplify_assessment.R

These changes included: 
* Remove data
  * Remove male lengths and ages
  * Change to not use EWAA
* 

## Simplify data (Try without re-running model and no EWAA)
COMMENT: I really could use some more info on what other types of ss3 parameters and options would need to change before converting to FIMS ... I'm so confused on what exactly needs to change
```{r}
plaice_input <- r4ss::SS_read(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/")

# The alaska plaice model is strange in a way that only a few AFSC seem to be. It uses both empirical weight-at-age data for most calculations AND growth # parameters to fit to the length data, which can be internally inconsistent approaches. For illustrating FIMS, I think it would be better to focus on one # option or the other. If you choose the weight-at-age approach, you should remove the length comp data. If you choose the parametric growth option, you 
# should remove the weight_at_age and length_to_age_conversion elements from the FIMSFrame (or wait for fixed r4ss converter).

# You could also switch gears to a Nanaimo-based model. I'm familiar with this dogfish model https://github.com/pbs-assess/dogfish-assess/tree/main/ss3/A1 although it has stock-recruit options that aren't available in FIMS but I could help translate the parameters.

##################################################################
## SIMPLIFY SS3 DATA FILE
# All of the changes to the data file are what is needed to be able to convert SS3 -> FIMS
dat <- plaice_input$dat

# change dimensions
dat$Nsexes <- 1 # previously 2, but FIMS is only single sex

# FIMS can only do single sex right now, so remove male age comp data
dat$agecomp <-
    dat$agecomp |>
    # exclude male compositions (columns matching pattern "m" followed
    # by numeric value for length bin, but doesn't match "month")
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    # assign observations to sex (Gender in 2024_BSAI_AKP.dat) 1 instead of 3
    dplyr::mutate(sex = -1)

# simplify length comps
dat$lencomp <-
    dat$lencomp |>
    # exclude male compositions (columns matching pattern "m" followed
    # by numeric value for length bin, but doesn't match "month")
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    dplyr::mutate(sex = -1)

##################################################################
## SIMPLIFY SS3 CONTROL FILE
## All the changes to the control file are what is needed to be able to better compare a FIMS and SS3 file
ctl <- plaice_input$ctl
ctl$EmpiricalWAA <- 0

ctl$Nsexes <- dat$Nsexes # previously 2

new_input <- plaice_input
new_input$dat <- dat
new_input$ctl <- ctl

new_input$dir <- "plaice_model_FIMS"
if (!dir.exists(new_input$dir)) {dir.create(new_input$dir)}

r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)

# This is still requiring a report file which doesn't make sense if you shouldn't HAVE to run the model prior to doing anything
data_4_model <- r4ss::ss3_data_to_fims(
  ss3_dir = new_input$dir,
  ss_new = FALSE,
)
```



## Re-run but no EWAA

```{r}
plaice_input <- r4ss::SS_read(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/")

##################################################################
## SIMPLIFY SS3 DATA FILE
# All of the changes to the data file are what is needed to be able to convert SS3 -> FIMS
dat <- plaice_input$dat

# change dimensions
dat$Nsexes <- 1 # previously 2, but FIMS is only single sex

# FIMS can only do single sex right now, so remove male age comp data
dat$agecomp <-
    dat$agecomp |>
    # exclude male compositions (columns matching pattern "m" followed
    # by numeric value for length bin, but doesn't match "month")
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    # assign observations to sex (Gender in 2024_BSAI_AKP.dat) 1 instead of 3
    dplyr::mutate(sex = -1)

# simplify length comps
dat$lencomp <-
    dat$lencomp |>
    # exclude male compositions (columns matching pattern "m" followed
    # by numeric value for length bin, but doesn't match "month")
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    dplyr::mutate(sex = -1)

##################################################################
## SIMPLIFY SS3 CONTROL FILE
## All the changes to the control file are what is needed to be able to better compare a FIMS and SS3 file
ctl <- plaice_input$ctl
ctl$EmpiricalWAA <- 0

ctl$Nsexes <- dat$Nsexes # previously 2

# remove all male mortality and growth parameters
ctl$MG_parms <- ctl$MG_parms |>
    dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))

# Change selectivity to be logistic instead of double normal since FIMS doesn't have that option and the selectivity curves looked logistic
# Also change Male to 0 because we will only have 1 sex and so we don't need any male offset
ctl$age_selex_types <- ctl$age_selex_types |>
    dplyr::mutate(
      Pattern = c(12, 12),
      Male = c(0, 0)
    )

ctl$age_selex_parms <- ctl$age_selex_parms |>
    dplyr::filter(!grepl("_PMalOff_", rownames(ctl$age_selex_parms)))

# remove blocks
ctl$N_Block_Designs <- 0
ctl$blocks_per_pattern <- NULL
ctl$Block_Design <- NULL

# remove variance adjustment
ctl$Variance_adjustment_list <- NULL
ctl$DoVar_adjust <- 0
ctl$sd_offset <- 0

ctl$MainRdevYrFirst <- dat$styr 
ctl$MainRdevYrLast <- dat$endyr
# Need to subtract Nages because of something to do with initial numbers at age
ctl$recdev_early_start <- dat$styr - dat$Nages

# Turn off forecast file
fore <- plaice_input$fore
fore$Forecast <- 0 # turn off forecast

# write modified data and control files
new_input <- plaice_input
new_input$dat <- dat
new_input$ctl <- ctl
new_input$fore <- fore

new_input$dir <- "plaice_model_noEWAA_FIMS"
if (!dir.exists(new_input$dir)) {dir.create(new_input$dir)}

r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)

# For the converter to work, if are using empirical weight-at-age and length comps are present or age-to-length conversion then you need output files and thus would need to run an ss3 model once any changes have been made before you can proceed

data_4_model <- r4ss::ss3_data_to_fims(
  ss3_dir = new_input$dir,
  ss_new = FALSE
)

# If you need to get the version of ss3 that the model originally was run with, you can do that
# For simplicity's sake, I'm not doing that here.
r4ss::get_ss3_exe(dir = new_input$dir)

# run model and compare output
r4ss::run(
    new_input$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```


## Using EWAA
```{r}
plaice_input <- r4ss::SS_read(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    read_wtatage = TRUE)

##################################################################
## SIMPLIFY SS3 DATA FILE
# All of the changes to the data file are what is needed to be able to convert SS3 -> FIMS
dat <- plaice_input$dat

# change dimensions
dat$Nsexes <- 1 # previously 2, but FIMS is only single sex

# FIMS can only do single sex right now, so remove male age comp data
dat$agecomp <-
    dat$agecomp |>
    # exclude male compositions (columns matching pattern "m" followed
    # by numeric value for length bin, but doesn't match "month")
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    # assign observations to sex (Gender in 2024_BSAI_AKP.dat) 1 instead of 3
    dplyr::mutate(sex = -1)

# simplify length comps
dat$use_lencomp <- 0

##################################################################
## SIMPLIFY SS3 CONTROL FILE
## All the changes to the control file are what is needed to be able to better compare a FIMS and SS3 file
ctl <- plaice_input$ctl

ctl$Nsexes <- dat$Nsexes # previously 2

# remove all male mortality and growth parameters
ctl$MG_parms <- ctl$MG_parms |>
    dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))

# Change selectivity to be logistic instead of double normal since FIMS doesn't have that option and the selectivity curves looked logistic
# Also change Male to 0 because we will only have 1 sex and so we don't need any male offset
ctl$age_selex_types <- ctl$age_selex_types |>
    dplyr::mutate(
      Pattern = c(12, 12),
      Male = c(0, 0)
    )

ctl$age_selex_parms <- ctl$age_selex_parms |>
    dplyr::filter(!grepl("_PMalOff_", rownames(ctl$age_selex_parms)))

# remove blocks
ctl$N_Block_Designs <- 0
ctl$blocks_per_pattern <- NULL
ctl$Block_Design <- NULL

# remove variance adjustment
ctl$Variance_adjustment_list <- NULL
ctl$DoVar_adjust <- 0
ctl$sd_offset <- 0

ctl$MainRdevYrFirst <- dat$styr 
ctl$MainRdevYrLast <- dat$endyr
# Need to subtract Nages because of something to do with initial numbers at age
ctl$recdev_early_start <- dat$styr - dat$Nages

wtatage <- plaice_input$wtatage
# wtatage for fleets -1:0 (population), and fleets 1:4 (real fleets)
wtatage_pop <- wtatage |>
    dplyr::select(1:6, paste(0:dat$Nages)) |>
    dplyr::filter(year == 1876 & sex == 1 & fleet == -1) |>
    dplyr::mutate(year = -year)
# original model had a fecundity relationship
# so recalculate the maturity * fecundity based on the population weight-at-age
# and an age-based maturity curve (logistic with 50% maturity at age 6 and slope of 1.5)
# this is a rough approximation to the length-based maturity converted to age
# within the original model, but should be sufficient for testing the simplified model
maturity <- 1 / (1 + exp(-1.5 * (0:dat$Nages - 6)))
matfec <- maturity * as.numeric(dplyr::select(wtatage_pop, paste0(0:dat$Nages)))
wtatage_matfec <- wtatage_pop
wtatage_matfec[, paste(0:dat$Nages)] <- matfec


wtatage_simple <- rbind(
    wtatage_matfec, # fleet = -2
    wtatage_pop, # fleet = -1
    wtatage_pop, # fleet = 0
    wtatage_pop, # fleet = 1
    wtatage_pop, # fleet = 2
    wtatage_pop, # fleet = 3
    wtatage_pop # fleet = 4
  ) |>
    dplyr::mutate(fleet = -2:4)

# Turn off forecast file
fore <- plaice_input$fore
fore$Forecast <- 0 # turn off forecast

# write modified data and control files
new_input <- plaice_input
new_input$dat <- dat
new_input$ctl <- ctl
new_input$fore <- fore

new_input$dir <- "plaice_model_EWAA_FIMS"
if (!dir.exists(new_input$dir)) {dir.create(new_input$dir)}

r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)

# For the converter to work, if are using empirical weight-at-age and length comps are present or age-to-length conversion then you need output files and thus would need to run an ss3 model once any changes have been made before you can proceed

data_4_model <- r4ss::ss3_data_to_fims(
  ss3_dir = new_input$dir,
  ss_new = FALSE
)

# If you need to get the version of ss3 that the model originally was run with, you can do that
# For simplicity's sake, I'm not doing that here.
r4ss::get_ss3_exe(dir = new_input$dir)

# run model and compare output
r4ss::run(
    new_input$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```

## Setting up the data
QUESTION: How am I even supposed to know what I need here, or what the options are for everything?
QUESTION: If you aren't converting your model from ss3, trying to load in data, really probably long lists of parameter values, seems like a nightmare and I would just give up and cry
COMMENT: I know it's not "cool" to use real data in examples, but we need 1) a vignette with a step-by-step narration on what decisions were made and what things to take into account when converting a model to FIMS 2) we need a vignette with step-by-step narration for creating a new model in FIMS using real data for a real species. All this FAKE DATA makes zero sense to me
```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# convert SS3 data into FIMS format using function in r4ss and convert to FIMSFrame object
data_4_model <- r4ss::ss3_data_to_fims(
    ss3_dir = "plaice_model_FIMS",
    ss_new = TRUE
) |>
    FIMS::FIMSFrame()

# read SS3 output files from plaice assessment
rep <- r4ss::SS_output(
    "plaice_model_FIMS",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)
```


## Run FIMS Model

### Setup default parameters first
```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
  data = data_4_model
)

# View fixed params that we NEED to make sure are right about for our species
default_parameters |>
  dplyr::filter(estimation_type == "constant") |>
  dplyr::count(module_name, fleet, label, value)

# View estimated params
default_parameters |>
  dplyr::filter(estimation_type == "fixed_effects") |>
  dplyr::count(module_name, fleet, label, value)

# Get number of years and ages from the data to make sense out of parameter counts
n_years <- FIMS::get_n_years(data_4_model)
n_years

n_ages <- FIMS::get_ages(data_4_model)
n_ages
```



```{r}
nat_M <- rep$parameters |>
                dplyr::filter(grepl("NatM.*Fem", Label)) |>
                dplyr::pull(Value)
# This gives you a table similar to what is given for setup default parameters and you will filter those out and rbind these in
init_naa <- FIMS:::setup_default_log_init_naa(n_ages = FIMS::get_n_ages(data_4_model),
                            log_M = nat_M,
                            # Might need to add 1000 the ss3 R0, might not, no one knows
                            log_rzero = rep$parameters |> dplyr::filter(grepl("R0", Label)) |> dplyr::pull(Value)
                        )

init_naa <- FIMS:::setup_default_log_init_naa(n_ages = FIMS::get_n_ages(data_4_model),
                            log_M = 0.2,
                            log_rzero = 12
                        )

# Update the NAA with what we calculated
default_parameters <- default_parameters |> 
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = FIMS::get_ages(data_4_model),
            value = init_naa,
        ),
        by = c("module_name", "label", "age")
    )



# NAA doesn't exist in SS3, FIMS estimates initial NAA as parameters (one param for each age) and you can start this whenever you want,
# This is similar to what WHAM does
# You can calculate what you think your initial numbers at age should be based on your R0 parameter and natural mortality
# Bai 


# function to get estimated parameter value from SS3 model
get_ss3_estimate <- function(string) {
    rep$parameters |>
        dplyr::filter(grepl(
            pattern = string,
            Label,
            fixed = TRUE
        )) |>
        dplyr::pull(Value)
}

parameters <- default_parameters |>
    # COMMENT: All of this stuff for growth is chaotic and I don't like it
    dplyr::rows_update(
        tibble::tibble(
        module_name = "Growth",
        label = c(
            "mean_length_young",
            "mean_length_old",
            "von_bertalanffy_coefficient_K",
            "reference_age_for_length_1",
            "reference_age_for_length_2",
            "length_weight_a", # added length-weight relationship parameters
            "length_weight_b"
        ),
        # values below are SS3 estimates, code won't work for 2-sex model
        # or model with time-varying growth parameters
        value = c(
            rep$Growth_Parameters$L_a_A1,
            rep$Growth_Parameters$L_a_A2,
            rep$Growth_Parameters$K,
            rep$Growth_Parameters$A1,
            rep$Growth_Parameters$A2,
            rep$Growth_Parameters$WtLen1,
            rep$Growth_Parameters$WtLen2
        ),
        estimation_type = "constant"
        ),
        by = c("module_name", "label")
    ) |>
    # change first reference age for sd values from 0 to 1
    dplyr::mutate(
        age = dplyr::if_else(
        module_name == "Growth" &
            label == "length_at_age_sd_at_ref_ages" &
            age == 0,
        1,
        age
        )
    ) |>
    dplyr::rows_update(
        # Note: this could probably be combined with Growth block above and put NAs for age for the other parameters
        tibble::tibble(
        module_name = "Growth",
        label = "length_at_age_sd_at_ref_ages",
        # Go back and extract these from rep$ rather than hard coding them
        age = c(1, 25),
        value = c(rep$Growth_Parameters$CVmin, rep$Growth_Parameters$CVmax)
        ),
        by = c("module_name", "label", "age")
    ) |>
    # QUESTION: Where are the inflection_point & slope values coming from?
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Maturity",
            label = c("inflection_point", "slope"),
            value = c(7.0, 1.0), # maturity based on the derived age-based pattern from original length-based SS3 model, arbitrary guess of slope
        ),
        by = c("module_name", "label")
    ) |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                get_ss3_estimate("SR_LN(R0)"),
                # QUESTION: Why is logit used here, how are you supposed to know you need to do this?
                # QUESTION: Why were these logit numbers chosen? Also, if it has to be a part of the conversion for everyone, shouldn't it just be integrated into the get_ss3_estimate() function?
                # For steepness we are always trying to the transformation between 0.2 and 1
                FIMS::logit(0.2, 1.0, get_ss3_estimate("SR_BH_steep")),
                log(get_ss3_estimate("SR_sigmaR"))
            )
        ),
        by = c("module_name", "label")
    ) |>
    # set initial value for log_q based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            fleet = "Survey",
            label = "log_q",
            value = get_ss3_estimate("LnQ_base_Survey(2)")
            # QUESTION: Why is there no estimation_type for this?
        ),
        by = c("fleet", "label")
    ) |>
    # set initial numbers at age based on SS3 output calculated above
    # TODO:will replace this with what I did up top
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = FIMS::get_ages(data_4_model),
            value = log(init_naa),
            estimation_type = "fixed_effects"
        ),
        by = c("module_name", "label", "age")
    ) |>
    # set initial M based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log(get_ss3_estimate("NatM_uniform_Fem_GP_1")),
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )
```

## Checks
QUESTION: If these are all useful checks and such, should there be a function in FIMS that is like data_check() to do all of this?
```{r}
# confirm that all columns except "value" and "estimation_type" are identical between parameters and default_parameters
if (!identical(
    parameters |>
        dplyr::select(-value, -estimation_type),
    default_parameters |>
        dplyr::select(-value, -estimation_type)
)) {
    cli::cli_alert_danger("Columns other than 'value' and 'estimation_type' are not identical between parameters and default_parameters")
}
# figure out if any rows have different estimate_type
cli::cli_alert_info("Rows with different estimation_type between default_parameters parameters:")
diff_estimation_type <- parameters |>
    dplyr::filter(estimation_type != default_parameters$estimation_type) |>
    dplyr::select(module_name, fleet, label, estimation_type)

if (nrow(diff_estimation_type) == 0) {
    cli::cli_alert_success("No rows with different estimation_type.")
} else {
    print(diff_estimation_type, n = Inf, width = Inf)
}

# parameters which have been updated from defaults
cli::cli_alert_info("Rows with updated initial values from default_parameters to parameters:")
updated_parameter_counts <- parameters |>
    dplyr::filter(value != default_parameters$value) |>
    dplyr::select(module_name, fleet, label, value) |>
    dplyr::count(module_name, label)

if (nrow(updated_parameter_counts) == 0) {
    cli::cli_alert_success("No rows with updated initial values.")
} else {
    print(updated_parameter_counts, n = Inf, width = Inf)
}
```

## Actually run the model
```{r}
# Run the model without optimization to help ensure a viable model
test_fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = FALSE)

# Run the model with optimization
fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)
fit_est <- fit # backup in case we want to set fit <- test_fit
```

## Additional model comparison work

FIMS doesn't have all of the bells and whistles that SS3 has yet, so if we want to make a more accurate comparison to FIMS, we will need to make a SS3 model that is more simplified. We already simplified the data file to only have one sex at the top of this file, now we will simplify the control and turn off the forecast file, and re-run the model
```{r}
##################################################################
## SIMPLIFY SS3 CONTROL FILE
## All the changes to the control file are what is needed to be able to better compare a FIMS and SS3 file
ctl <- plaice_input$ctl
ctl$Nsexes <- dat$Nsexes # previously 2

# remove all male mortality and growth parameters
ctl$MG_parms <- ctl$MG_parms |>
    dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))

# Change selectivity to be logistic instead of double normal since FIMS doesn't have that option and the selectivity curves looked logistic
# Also change Male to 0 because we will only have 1 sex and so we don't need any male offset
ctl$age_selex_types <- ctl$age_selex_types |>
    dplyr::mutate(
      Pattern = c(12, 12),
      Male = c(0, 0)
    )

ctl$age_selex_parms <- ctl$age_selex_parms |>
    dplyr::filter(!grepl("_PMalOff_", rownames(ctl$age_selex_parms)))

# remove blocks
ctl$N_Block_Designs <- 0
ctl$blocks_per_pattern <- NULL
ctl$Block_Design <- NULL

# remove variance adjustment
ctl$Variance_adjustment_list <- NULL
ctl$DoVar_adjust <- 0
ctl$sd_offset <- 0


ctl$MainRdevYrFirst <- dat$styr 
ctl$MainRdevYrLast <- dat$endyr
# Need to subtract Nages because of something to do with initial numbers at age
ctl$recdev_early_start <- dat$styr - dat$Nages


# Turn off forecast file
fore <- plaice_input$fore
fore$Forecast <- 0 # turn off forecast

# write modified data and control files
new_input <- plaice_input
new_input$dat <- dat
new_input$ctl <- ctl
new_input$fore <- fore

new_input$dir <- "plaice_model_FIMS"
  if (!dir.exists(new_input$dir)) {
    dir.create(new_input$dir)
  }

r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)

# run model and compare output
r4ss::run(
    new_input$dir,
    show_in_console = TRUE,
    skipfinished = FALSE
)

# QUESTION: Will I need this?
# r4ss::run(
#     new_input$dir,
#     show_in_console = TRUE,
#     skipfinished = FALSE,
#     extras = "-hess_step" # was exceeding default maxfn 100 in early phases
# )

# read output from model that was just run
plaice_simple_output <- r4ss::SS_output(
    new_input$dir,
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

r4ss::SS_plots(plaice_simple_output)

# read SS3 output files from plaice assessment on GitHub
plaice_output1 <- r4ss::SS_output(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

# compare outputs from simplified and original SS3 models
# (fairly large changes in spawning output, perhaps due to removal of the older age classes)
SSplotComparisons(SSsummarize(list(
    plaice_simple_output,
    plaice_output1
    )))
```