---
title: "AFSC Case Study: Converting a Bering Sea Alaska Plaice SS3 Model to FIMS"
stock: "AFSC Alaska Plaice"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
    toc: true
    number-sections: true
---

## Introduction

This document serves as a case study and guide for converting a stock assessment model from SS3 to FIMS framework. We use the Bering Sea and Aleutian Islands (BSAI) Alaska Plaice assessment in this example.

The goal is to demonstrate a reproducible workflow, highlight common challenges, and provide solutions for simplifying a complex SS3 model to a structure compatible with FIMS. This process often involves making simplifying assumptions, and we will walk through these decisions.

We will cover two main pathways for conversion, based on how the model handles growth:

1.  **Using Growth Parameters**: The model estimates growth using a von Bertalanffy growth function, and weight is derived from a length-weight relationship.
2.  **Using Empirical Weight-at-Age (EWAA)**: The model uses observed mean weight-at-age data directly, bypassing the estimation of growth parameters.

For each pathway, we will follow these general steps:
1.  Simplify the original SS3 data and control files.
2.  Run the simplified SS3 model to generate necessary output files.
3.  Convert the simplified SS3 output into a FIMS data object.
4.  Set up and run the corresponding FIMS model.
5.  Compare the results between the original SS3 model, the simplified SS3 model, and the FIMS model.

## The Setup

This section loads necessary libraries and defines some common variables that will be used throughout the analysis.

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Alaska plaice"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

**A Note on This Demonstration:** The models presented in this case study are changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration of the FIMS conversion process and nothing more.

## Part 1: Preparing a Simplified SS3 Model for FIMS Conversion

FIMS is a powerful but newer framework and does not yet have all the features of SS3 (e.g., two-sex models, certain selectivity patterns). Therefore, a critical first step is often to simplify the existing SS3 model to a "FIMS-compatible" structure. This also helps in making a more direct, "apples-to-apples" comparison between the two modeling platforms.

The original Alaska plaice model is a two-sex model. As FIMS currently only supports single-sex (or combined-sex) models, our primary simplification will be to convert the model to a female-only model.

To keep our code organized and reduce repetition, we'll create a helper function to perform these simplifications below.

```{r}
#| label: simplification-function

#' Simplify SS3 data and control files for FIMS conversion
#'
#' This function takes an SS3 input list (from r4ss::SS_read) and applies
#' a series of simplifications to make it more compatible with FIMS and allow
#' for a cleaner comparison.
#'
#' @param input_list An SS3 input list from r4ss::SS_read.
#' @param use_ewaa A logical flag. If TRUE, the model will be configured to use
#'   Empirical Weight-at-Age and length comps will be turned off.
#' @param output_dir The directory to write the new model files to.
#'
#' @return The modified input_list.

simplify_ss3_for_fims <- function(input_list, use_ewaa, output_dir) {
  
  dat <- input_list$dat
  ctl <- input_list$ctl
  fore <- input_list$fore
  
  # --- DATA FILE SIMPLIFICATIONS ---
  # Convert to a single-sex model (female-only)
  dat$Nsexes <- -1
  
  # Remove male age compositions
  dat$agecomp <- dat$agecomp |>
    dplyr::select(!dplyr::matches("^m\\d+$")) |>
    dplyr::mutate(sex = -1)
  
  if (use_ewaa) {
    # If using EWAA, we don't use length data
    dat$use_lencomp <- 0
  } else {
    # If not using EWAA, remove male length compositions
    dat$lencomp <- dat$lencomp |>
      dplyr::select(!dplyr::matches("^m\\d+$")) |>
      dplyr::mutate(sex = -1)
    ctl$EmpiricalWAA <- 0
  }
  
  # --- CONTROL FILE SIMPLIFICATIONS ---
  ctl$Nsexes <- dat$Nsexes
  
  # Remove all male-specific mortality and growth parameters
  ctl$MG_parms <- ctl$MG_parms |>
    dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))
  
  # Change selectivity to logistic (FIMS option) instead of double-normal.
  # Set Male to 0 as we no longer have a male-specific curve.
  ctl$age_selex_types <- ctl$age_selex_types |>
    dplyr::mutate(
      Pattern = 12, # 12 = logistic
      Male = 0
    )
  
  # Replace double-normal selectivity parameters with logistic parameters.
  # Remove male offset parameters.
  ctl$age_selex_parms <- cbind(
    data.frame(
      LO = 1, 
      HI = 15, 
      INIT = c(10, 2, 7, 2), 
      PRIOR = 0,
      PR_SD = 99, 
      PR_type = 0, 
      PHASE = 2
    ),
    matrix(0, nrow = 4, ncol = 7) # Fills remaining columns with 0
  )
  names(ctl$age_selex_parms) <- names(ctl$MG_parms)
  rownames(ctl$age_selex_parms) <- paste0(
    "AgeSel_", 
    c("P_1_", "P_2_"), 
    c("Fishery", "Fishery", "Survey", "Survey")
  )
  
  # Remove blocks and variance adjustments for simplicity
  ctl$N_Block_Designs <- 0
  ctl$blocks_per_pattern <- NULL
  ctl$Block_Design <- NULL
  ctl$Variance_adjustment_list <- NULL
  ctl$DoVar_adjust <- 0
  ctl$sd_offset <- 0
  
  # Adjust recruitment deviation years
  ctl$MainRdevYrFirst <- dat$styr
  ctl$MainRdevYrLast <- dat$endyr
  ctl$recdev_early_start <- dat$styr - dat$Nages
  
  # --- FORECAST FILE ---
  # Turn off forecasting for this comparison
  fore$Forecast <- 0
  
  # --- WRITE MODIFIED FILES ---
  new_input <- input_list
  new_input$dat <- dat
  new_input$ctl <- ctl
  new_input$fore <- fore
  if (use_ewaa) {
    new_input$wtatage <- input_list$wtatage
  }
  
  # Name and create output directory
  new_input$dir <- output_dir
  if (!dir.exists(new_input$dir)) {
    dir.create(new_input$dir)
  }
  
  # Write to output directory
  r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)
  
  return(new_input)
}
```

Now we can use this function to create our two simplified SS3 models.

### Scenario 1: Simplified Model using Growth Parameters (No EWAA)

Here, we simplify the SS3 model and configure it to estimate growth internally rather than using empirical weight-at-age.

```{r}
#| label: prep-no-ewaa-model
# Read the original assessment model from GitHub
plaice_input_base <- r4ss::SS_read(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    read_wtatage = TRUE, # Read wtatage file even if not used
    ss_new = TRUE
)

# Create and write the simplified model files
simplified_no_ewaa <- simplify_ss3_for_fims(
  input_list = plaice_input_base,
  use_ewaa = FALSE,
  output_dir = "plaice_model_noEWAA_FIMS"
)

# Run the simplified SS3 model
# For the converter to work robustly, especially when length comps are present,
# it's best to run the model to generate output files like Report.sso.
r4ss::get_ss3_exe(dir = simplified_no_ewaa$dir)
r4ss::run(
    dir = simplified_no_ewaa$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```

### Scenario 2: Simplified Model using Empirical Weight-at-Age (EWAA)

In this scenario, we simplify the model but configure it to use the empirical weight-at-age data directly. This means we will turn off the length composition data.

```{r}
#| label: prep-ewaa-model

# We can re-use the plaice_input_base object we read in the previous chunk.
# Create and write the simplified model files for the EWAA case.
simplified_ewaa <- simplify_ss3_for_fims(
  input_list = plaice_input_base,
  use_ewaa = TRUE,
  output_dir = "plaice_model_EWAA_FIMS"
)

# Run the simplified SS3 model
# When using EWAA, the ss3_data_to_fims converter requires the Report.sso file
# to correctly process the weight-at-age information.
r4ss::get_ss3_exe(dir = simplified_ewaa$dir)
r4ss::run(
    dir = simplified_ewaa$dir,
    exe = "ss3",
    show_in_console = TRUE,
    skipfinished = FALSE
)
```

## Part 2: Converting the Simplified SS3 Model to FIMS

With our simplified and successfully run SS3 models in hand, we can now use `r4ss::ss3_data_to_fims` to convert them into FIMS data objects.

For the remainder of this walkthrough, we will proceed with the **"No EWAA"** model to demonstrate the FIMS setup process. The process for the EWAA model would be similar, with the main difference being that growth parameters would not need to be specified in the FIMS model.

```{r}
#| label: convert-to-fims
#| output: false
#| warning: false

# Convert the simplified "No EWAA" SS3 model output into the FIMS format
# and then create a FIMSFrame object.
data_4_fims <- r4ss::ss3_data_to_fims(
    ss3_dir = simplified_no_ewaa$dir,
    ss_new = TRUE
) |>
    FIMS::FIMSFrame()

# To proceed, we also need to read the output of our simplified SS3 run
# to get parameter estimates for starting values in FIMS.
rep_simple_ss3 <- r4ss::SS_output(
    simplified_no_ewaa$dir,
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)
```


```{r}
data_4_fims <- r4ss::ss3_data_to_fims(
    ss3_dir = simplified_ewaa$dir,
    ss_new = TRUE
) 
|>
    FIMS::FIMSFrame()

# To proceed, we also need to read the output of our simplified SS3 run
# to get parameter estimates for starting values in FIMS.
rep_ewaa_ss3 <- r4ss::SS_output(
    simplified_ewaa$dir,
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)
```

## Part 3: Building and Running the FIMS Model

Now we have the data for FIMS in the `data_4_fims` object. The next step is to define the model structure and provide initial parameter values.

### Setting Up Parameters

We start by creating a data frame of default parameters based on our data. We then modify this data frame with better starting values, often derived from the model we are converting from (our simplified SS3 model).

```{r}
#| label: setup-fims-params
#| max.height: '100px'
#| attr.output: '.numberLines'

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
  data = data_4_fims
)

# View fixed params (constants) that we NEED to make sure are right for our species
default_parameters |>
  dplyr::filter(estimation_type == "constant") |>
  dplyr::count(module_name, fleet, label, value)

# View estimated params
default_parameters |>
  dplyr::filter(estimation_type == "fixed_effects") |>
  dplyr::count(module_name, fleet, label, value)

# Get number of years and ages from the data to make sense out of parameter counts
n_years <- FIMS::get_n_years(data_4_fims)
n_ages <- FIMS::get_ages(data_4_fims)
```

### Populating Parameters from the Simplified SS3 Model

This is a critical step. Good starting values are essential for model convergence. We will (hopefully) eventually create a helper function to make it easier to pull values from the SS3 report object that is already in the `{r4ss}` package but for now we will create a helper function to help us get these values.

```{r}
#| label: get-ss3-param-helper

# Helper function to get an estimated parameter value from the SS3 report object
get_ss3_estimate <- function(string, report = rep_simple_ss3) {
    report$parameters |>
        dplyr::filter(grepl(
            pattern = string,
            Label,
            fixed = TRUE
        )) |>
        dplyr::pull(Value)
}
```


In FIMS, initial numbers-at-age (NAA) are estimated parameters. In SS3, they are derived from R0 and M. We can calculate initial values for FIMS based on the SS3 equilibrium assumption using `setup_default_log_init_naa()`.

```{r}
#| label: naa
log_m_init <- log(get_ss3_estimate("NatM_uniform_Fem_GP_1"))
log_rzero_init <- get_ss3_estimate("SR_LN(R0)")
init_naa <- FIMS:::setup_default_log_init_naa(
    n_ages = FIMS::get_ages(data_4_fims),
    log_M = log_m_init,
    log_rzero = log_rzero_init
)
```


Now, let's create our final parameter set for FIMS by updating the defaults.

```{r}
#| label: populate-fims-params.
parameters <- default_parameters |>
    # Update Growth parameters. These are treated as fixed constants in this example,
    # using the estimates from the SS3 model. If using EWAA, this block would be removed.
    dplyr::rows_update(
        tibble::tibble(
        module_name = "Growth",
        label = c(
            "mean_length_young", "mean_length_old", "von_bertalanffy_coefficient_K",
            "reference_age_for_length_1", "reference_age_for_length_2",
            "length_weight_a", "length_weight_b"
        ),
        value = c(
            rep_simple_ss3$Growth_Parameters$L_a_A1, rep_simple_ss3$Growth_Parameters$L_a_A2,
            rep_simple_ss3$Growth_Parameters$K, rep_simple_ss3$Growth_Parameters$A1,
            rep_simple_ss3$Growth_Parameters$A2, rep_simple_ss3$Growth_Parameters$WtLen1,
            rep_simple_ss3$Growth_Parameters$WtLen2
        ),
        estimation_type = "constant"
        ),
        by = c("module_name", "label")
    ) |>
    # Update SD of length-at-age
    dplyr::rows_update(
        tibble::tibble(
        module_name = "Growth",
        label = "length_at_age_sd_at_ref_ages",
        age = c(rep_simple_ss3$Growth_Parameters$A1, rep_simple_ss3$Growth_Parameters$A2),
        value = c(rep_simple_ss3$Growth_Parameters$CVmin, rep_simple_ss3$Growth_Parameters$CVmax)
        ),
        by = c("module_name", "label", "age")
    ) |>
    # Update Maturity parameters.
    # QUESTION: Where are the inflection_point & slope values coming from?
    # ANSWER: These were derived from the original length-based maturity in the SS3
    # model and converted to an age-based approximation. The slope was an arbitrary guess.
    # In a real assessment, these would be carefully chosen based on biological data.
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Maturity",
            label = c("inflection_point", "slope"),
            value = c(7.0, 1.0),
        ),
        by = c("module_name", "label")
    ) |>
    # Update Recruitment (Stock-Recruit Relationship) parameters
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                log_rzero_init,
                # QUESTION: Why is logit used here?
                # ANSWER: Steepness (h) is a parameter bounded between 0.2 and 1.0.
                # Optimization algorithms work best on an unconstrained scale (-Inf to Inf).
                # The logit transform maps the value from its bounded range to an
                # unconstrained one. FIMS provides a helper for this common transformation.
                FIMS::logit(0.2, 1.0, get_ss3_estimate("SR_BH_steep")),
                log(get_ss3_estimate("SR_sigmaR"))
            )
        ),
        by = c("module_name", "label")
    ) |>
    # Update survey catchability (log_q)
    dplyr::rows_update(
        tibble::tibble(
            fleet = "Survey",
            label = "log_q",
            value = get_ss3_estimate("LnQ_base_Survey(2)")
        ),
        by = c("fleet", "label")
    ) |>
    # Update initial numbers-at-age (NAA)
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = n_ages,
            value = init_naa,
            estimation_type = "fixed_effects"
        ),
        by = c("module_name", "label", "age")
    ) |>
    # Update Natural Mortality (M)
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log_m_init,
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )
```

### Sanity Checks

Before running the model, it's good practice to check how our final parameter set (`parameters`) differs from the defaults. This helps catch mistakes.

```{r}
#| label: fims-param-checks
# This is a great practice to ensure your parameter setup is what you expect.
# Should this be a built-in FIMS function? Maybe! For now, this custom
# check is very helpful for debugging.

# Confirm that all identifying columns are identical
if (!identical(
    parameters |> dplyr::select(-value, -estimation_type),
    default_parameters |> dplyr::select(-value, -estimation_type)
)) {
    cli::cli_alert_danger("Parameter data frames are not structurally identical!")
}

# Show which parameters have been updated from their default values
cli::cli_alert_info("Rows with updated initial values from defaults:")
updated_parameter_counts <- parameters |>
    dplyr::filter(value != default_parameters$value) |>
    dplyr::select(module_name, fleet, label, value) |>
    dplyr::count(module_name, label)
print(updated_parameter_counts, n = Inf, width = Inf)
```

### Running the FIMS Model

Finally, we can run the model. It's often a good idea to first run without optimizing to check for any immediate errors in the model setup (`test_fit`). Then, run with optimization to get the final fit.

```{r}
#| label: run-fims-model
# Run the model without optimization to check for errors
test_fit <- parameters |>
    FIMS::initialize_fims(data = data_4_fims) |>
    FIMS::fit_fims(optimize = FALSE)

# Run the model with optimization
fit <- parameters |>
    FIMS::initialize_fims(data = data_4_fims) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)
```

## Part 4: Comparing Model Outputs

A crucial final step is to compare the results. Here, we'll use `r4ss::SSplotComparisons` to visually compare the output of our original, complex SS3 model with our simplified SS3 model. This helps us understand the impact of the simplifications we made. A full comparison would also include plotting the FIMS results against the SS3 models.

```{r}
#| label: compare-models
# Read output from the original SS3 model on GitHub for comparison
plaice_output_original <- r4ss::SS_output(
    "https://raw.githubusercontent.com/afsc-assessments/BSAI_Alaska_plaice/main/2024/Models/Base-3/",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

# Read output from our simplified SS3 model (already in rep_simple_ss3)

# Create a summary object for comparison plotting
summary_list <- r4ss::SSsummarize(list(
    "Original SS3" = plaice_output_original,
    "Simplified SS3" = rep_simple_ss3
))

# Plot the comparisons
r4ss::SSplotComparisons(summary_list)
```

The comparison plots show that removing the male data and changing the selectivity and growth options had a noticeable effect on the estimated spawning output, highlighting the importance of understanding the impact of such simplifying assumptions.

## Appendix: Notes on Steepness

The user had a note about steepness from a spiny dogfish assessment. This is a good example of borrowing information from other assessments or literature, which is common practice.

For Beverton-Holt steepness, the value is often difficult to estimate within the model and is frequently fixed based on a meta-analysis or prior knowledge. The example code shows how you would insert a fixed value for steepness (e.g., `h = 0.283`) into the FIMS parameter setup.

```{r}
#| label: steepness-example
#| eval: false
#| echo: true
# Example of setting a fixed steepness value
h_fixed <- 0.283
parameters |>
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = "logit_steep",
            # Use the logit transform on the fixed value
            value = FIMS::logit(0.2, 1.0, h_fixed),
            # Set as a constant so it is not estimated
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )
```