---
title: AFSC Case Study BSAI Atka mackerel
editor_options: 
  chunk_output_type: console
---


## The setup

```{r}
#| warning: false
#| label: startup
#| output: false

packages <- c("dplyr", "tidyr", "ggplot2", "TMB", "remotes")
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], repos = "http://cran.us.r-project.org")
}
#remotes::install_github("NOAA-FIMS/FIMS", ref='isNA')
remotes::install_github("kaskr/TMB_contrib_R/TMBhelper")

library(ggplot2)
library(tidyr)
library(dplyr)
library(FIMS)
library(TMB)
library(TMBhelper)
## devtools::install_github('afsc-assessments/GOApollock', ref='v0.1.2')
## library(GOApollock)
theme_set(theme_bw())


R_version <- version$version.string
TMB_version <- packageDescription("TMB")$Version
FIMS_commit <- substr(packageDescription("FIMS")$GithubSHA1, 1, 7)
```

-   R version: `r R_version`\
-   TMB version: `r TMB_version`\
-   FIMS commit: `r FIMS_commit`\
-   Stock name: BSAI Atka mackerel\
-   Region: AFSC\
-   Analyst: Jim Ianelli \

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the BSAI Atka mackerel stock. 

To get the opertional model to more closely match FIMS I:
* stumbled lots

## Script to prepare data for building FIMS object
```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false

## define the dimensions and global variables
#  Opent the original AMAK file
atka_dat <- readRDS(here::here("content","data_files","atka_dat.RDS"))
atka_rep <- readRDS(here::here("content","data_files","atka_rep.RDS"))
## define the dimensions and global variables
years <- atka_dat$styr:atka_dat$endyr
nyears <- length(years)
nseasons <- 1
nages <- 11
ages <- 1:nages
## nfleets <- 1
## This will fit the models bridging to FIMS (simplifying)
## source("fit_bridge_models.R")
## compare changes to model
#pkfitfinal <- readRDS("data_files/pkfitfinal.RDS")
#pkfit0 <- readRDS("data_files/pkfit0.RDS")
#parfinal <- pkfitfinal$obj$env$parList()
#pkinput0 <- readRDS('data_files/pkinput0.RDS')
#fimsdat <- pkdat0 <- pkinput0$dat
#pkinput <- readRDS('data_files/pkinput.RDS')
```




How I simplified my assessment:
* Haven't got there yet...


## Script that sets up and runs the model
```{r}
#| warning: false

# clear memory
clear()
clear_logs()
estimate_fish_selex <- TRUE
estimate_survey_selex <- TRUE
estimate_q2 <- TRUE
estimate_q3 <- TRUE
estimate_q6 <- TRUE
estimate_F <- TRUE
estimate_recdevs <- TRUE

# read the AMAK rdat files
# First generate them from the original AMAK file
#rdat <- dget(file.path("data_files", "NEFSC_YT_SIMPLIFIED.RDAT")) # to be used in FIMS, lots of modifications from original
#orig <- dget(file.path("data_files", "NEFSC_YT_ORIGINAL.RDAT"))   # where started before modifications for use in FIMS

# function to create equivalent of data_mile1, basic catch and survey data
# need to think about how to deal with multiple fleets and indices - only use 1 of each for now
#remotes::install_github("mlverse/chattr")
# q: What is the definition of standard error?
# a: The standard error is the standard deviation of the log-transformed data. This is the standard deviation of the log-transformed data, not the standard deviation of the data itself. The standard error is the standard deviation of the data itself.

# q: What is the definition of the standard deviation of the age composition data?
# a: The standard deviation of the age composition data is the standard deviation of the log-transformed age composition data. This is the standard deviation of the log-transformed age composition data, not the standard deviation of the age composition data itself. The standard deviation of the age composition data is the standard deviation of the age composition data itself.

# q: What is the definition of the standard deviation of the index data?
# a: The standard deviation of the index data is the standard deviation of the log-transformed index data. This is the standard deviation of the log-transformed index data, not the standard deviation of the index data itself. The standard deviation of the index data is the standard deviation of the index data itself.
get_amak_data <- function(rdat,rrep){
## put into fims friendly form
  res <- data.frame(type = character(),
                  name = character(),
                  age = integer(),
                  datestart = character(),
                  dateend = character(),
                  value = double(),
                  unit = character(),
                  uncertainty = double())
  landings <- data.frame(type = "landings",
                       name = "fleet1",
                       age = NA,
                       datestart = paste0(seq(rdat$styr, rdat$endyr), "-01-01"),
                       dateend = paste0(seq(rdat$styr, rdat$endyr), "-12-31"),
                       value = as.numeric(rdat$catch),
                       unit = "t",
                       uncertainty = rdat$catch_cv)
##  need to fill missing years with -999 so it's ignored in FIMS
  indtmp <- 0*rdat$catch-999
  indtmp[which(years %in% rdat$yrs_ind )] <- rdat$biom_ind
#indtmp
  CVtmp <- rep(1, length=nyears) # actually SE in log space
  CVtmp[which(years %in% rdat$yrs_ind)] <- rdat$biom_std/rdat$biom_ind
## repeat with fish catch at age, using expected in missing years
#names(atka_rep)
caa <- 0*rrep$N[,-1]-999
caa[which(years %in% rdat$yrs_ages_fsh), ] <- rdat$page_fsh
Ncaa <- rep(1, nyears)
Ncaa[which(years %in% rdat$yrs_ages_fsh)] <- rdat$sample_ages_fsh
paa2 <-  0*rrep$N[,-1]-999
paa2[which(years %in% rdat$yrs_ages_ind), ] <- rdat$page_ind
Npaa2 <- rep(1, nyears)
Npaa2[which(years %in% rdat$yrs_ages_ind)] <- rdat$sample_ages_ind
index <- data.frame(type = "index",
                    name = "survey",
                    age = NA,
                    datestart = paste0(seq(rdat$styr, rdat$endyr), "-01-01"),
                    dateend = paste0(seq(rdat$styr, rdat$endyr), "-12-31"),
                    value = ifelse(indtmp>0, indtmp, indtmp),
                    unit = "",
                    uncertainty = CVtmp)
## these have -999 for missing data years
catchage <- data.frame(type = "age",
                       name = "fleet1",
                       age = rep(seq(1,nages), nyears),
                       datestart = rep(paste0(seq(rdat$styr, rdat$endyr), "-01-01"), each=nages),
                       dateend = rep(paste0(seq(rdat$styr, rdat$endyr), "-12-31"), each=nages),
                       value = as.numeric(t(caa)),
                       unit = "",
                       uncertainty = rep(Ncaa, each=nages))
indexage <- data.frame(
          type = "age",
          name = "survey",
          age = rep(seq(1, nages), nyears),
          datestart = rep(paste0( seq(rdat$styr, rdat$endyr), "-01-01" ), each = nages),
          dateend = rep(paste0( seq(rdat$styr, rdat$endyr), "-12-31" ), each = nages),
          value = as.numeric(t(paa2)),
          unit = "",
          uncertainty = rep(Npaa2, each = nages)
)
## indexage=indexage2
## index=index2
res <- rbind(res, landings, index, catchage, indexage)
return(res)
}
fimsdat<-get_amak_data(atka_dat,atka_rep)


age_frame <- FIMS::FIMSFrameAge(fimsdat)
fishery_catch <- FIMS::m_landings(age_frame)
fishery_agecomp <- FIMS::m_agecomp(age_frame, "fleet1")
survey_index <- FIMS::m_index(age_frame, "survey")
survey_agecomp <- FIMS::m_agecomp(age_frame, "survey")
# need to think about how to deal with multiple fleets - only using 1 fleeet for now
fish_index <- methods::new(Index, nyears)
fish_age_comp <- methods::new(AgeComp, nyears, nages)
fish_index$index_data <- fishery_catch
fish_age_comp$age_comp_data <- fishery_agecomp * catchage$uncertainty#rep(Ncaa, each=nages)


### set up fishery
## fleet selectivity: converted from time-varying ascending
## slope/intercept to constant double-logistic
## methods::show(DoubleLogisticSelectivity)
fish_selex <- methods::new(DoubleLogisticSelectivity)
fish_selex$inflection_point_asc$value <- 3.0 #parfinal$inf1_fsh_mean
fish_selex$inflection_point_asc$is_random_effect <- FALSE
fish_selex$inflection_point_asc$estimated <- estimate_fish_selex
fish_selex$inflection_point_desc$value <- 3.0 #parfinal$inf2_fsh_mean
fish_selex$inflection_point_desc$is_random_effect <- FALSE
fish_selex$inflection_point_desc$estimated <- estimate_fish_selex
fish_selex$slope_asc$value <- 1. #exp(parfinal$log_slp1_fsh_mean)
fish_selex$slope_asc$is_random_effect <- FALSE
fish_selex$slope_asc$estimated <- estimate_fish_selex
fish_selex$slope_desc$value <- 1. #exp(parfinal$log_slp2_fsh_mean)
fish_selex$slope_desc$is_random_effect <- FALSE
fish_selex$slope_desc$estimated <- estimate_fish_selex
## create fleet object
fish_fleet <- methods::new(Fleet)
fish_fleet$nages <- nages
fish_fleet$nyears <- nyears
fish_fleet$log_Fmort <- log(atka_rep$F_age_1[,-1])
fish_fleet$estimate_F <- estimate_F
fish_fleet$random_F <- FALSE
fish_fleet$log_q <- 0 # why is this length two in Chris' case study?
fish_fleet$estimate_q <- FALSE
fish_fleet$random_q <- FALSE
fish_fleet$log_obs_error <- log(landings$uncertainty)
## fish_fleet$log_obs_error$estimated <- FALSE
# Next two lines not currently used by FIMS
fish_fleet$SetAgeCompLikelihood(1)
fish_fleet$SetIndexLikelihood(1)
# Set Index, AgeComp, and Selectivity using the IDs from the modules defined above
fish_fleet$SetObservedIndexData(fish_index$get_id())
fish_fleet$SetObservedAgeCompData(fish_age_comp$get_id())
fish_fleet$SetSelectivity(fish_selex$get_id())


## Setup survey 2
survey_fleet_index <- methods::new(Index, nyears)
survey_age_comp <- methods::new(AgeComp, nyears, nages)
survey_fleet_index$index_data <- survey_index
survey_age_comp$age_comp_data <- survey_agecomp * indexage$uncertainty
## survey selectivity: ascending logistic
## methods::show(DoubleLogisticSelectivity)
survey_selex <- methods::new(DoubleLogisticSelectivity)
survey_selex$inflection_point_asc$value <- 3.0 #parfinal$inf1_srv
survey_selex$inflection_point_asc$is_random_effect <- FALSE
survey_selex$inflection_point_asc$estimated <- estimate_survey_selex
survey_selex$slope_asc$value <- 3.0 #exp(parfinal$log_slp1_srv2)
survey_selex$slope_asc$is_random_effect <- FALSE
survey_selex$slope_asc$estimated <- estimate_survey_selex
## not estimated to make it ascending only, fix at input values
survey_selex$inflection_point_desc$value <- 3.0 #parfinal$inf2_srv2
survey_selex$inflection_point_desc$is_random_effect <- FALSE
survey_selex$inflection_point_desc$estimated <- FALSE
survey_selex$slope_desc$value <- 3.0 #exp(parfinal$log_slp2_srv2)
survey_selex$slope_desc$is_random_effect <- FALSE
survey_selex$slope_desc$estimated <- FALSE
survey_fleet <- methods::new(Fleet)
survey_fleet$is_survey <- TRUE
survey_fleet$nages <- nages
survey_fleet$nyears <- nyears
survey_fleet$estimate_F <- FALSE
survey_fleet$random_F <- FALSE
survey_fleet$log_q <- 1.0 #parfinal$log_q2_mean
survey_fleet$estimate_q <- estimate_q2
survey_fleet$random_q <- FALSE
# sd = sqrt(log(cv^2 + 1)), sd is log transformed
survey_fleet$log_obs_error <- log(index$uncertainty)
## survey_fleet$log_obs_error$estimated <- FALSE
survey_fleet$SetAgeCompLikelihood(1)
survey_fleet$SetIndexLikelihood(1)
survey_fleet$SetSelectivity(survey_selex$get_id())
survey_fleet$SetObservedIndexData(survey_fleet_index$get_id())
survey_fleet$SetObservedAgeCompData(survey_age_comp$get_id())

# Population module
# recruitment
recruitment <- methods::new(BevertonHoltRecruitment)
## methods::show(BevertonHoltRecruitment)
recruitment$log_sigma_recruit$value <- 0.6 #log(parfinal$sigmaR)
recruitment$log_rzero$value <- 11. #parfinal$mean_log_recruit + log(1e9)
recruitment$log_rzero$is_random_effect <- FALSE
recruitment$log_rzero$estimated <- TRUE
## note: do not set steepness exactly equal to 1, use 0.99 instead in ASAP run
recruitment$logit_steep$value <- -log(1.0 - .99999) + log(.99999 - 0.2)
recruitment$logit_steep$is_random_effect <- FALSE
recruitment$logit_steep$estimated <- FALSE
recruitment$estimate_log_devs <- estimate_recdevs
recruitment$log_devs <-  0. #parfinal$dev_log_recruit[-1]

## growth  -- assumes single WAA vector for everything, based on
## Srv1 above
waa <- atka_rep$wt_fsh_1
ewaa_growth <- methods::new(EWAAgrowth)
ewaa_growth$ages <- ages
# NOTE: FIMS currently cannot use matrix of WAA, so have to ensure constant WAA over time in ASAP file for now
ewaa_growth$weights <- waa[1,]
## NOTE: FIMS assumes SSB calculated at the start of the year, so
## need to adjust ASAP to do so as well for now, need to make
## timing of SSB calculation part of FIMS later
## maturity
## NOTE: for now tricking FIMS into thinking age 0 is age 1, so need to adjust A50 for maturity because FIMS calculations use ages 0-5+ instead of 1-6
maturity <- new(LogisticMaturity)
maturity$inflection_point$value <- 4.5
maturity$inflection_point$is_random_effect <- FALSE
maturity$inflection_point$estimated <- FALSE
maturity$slope$value <- 1.5
maturity$slope$is_random_effect <- FALSE
maturity$slope$estimated <- FALSE

# population
population <- new(Population)
population$log_M <- log( atka_rep$M) 
population$estimate_M <- FALSE
population$log_init_naa <- 11. #c(log(pkfitfinal$rep$recruit[1]), log(pkfitfinal$rep$initN)) + log(1e9)
population$estimate_init_naa <- FALSE # TRUE , NOTE: fixing at ASAP estimates to test SSB calculations
population$nages <- nages
population$ages <- ages
population$nfleets <- 2 # 1 fleet and 1 survey
population$nseasons <- nseasons
population$nyears <- nyears
## population$prop_female <- 1.0 # ASAP assumption
population$SetMaturity(maturity$get_id())
population$SetGrowth(ewaa_growth$get_id())
population$SetRecruitment(recruitment$get_id())

# make FIMS model
success <- CreateTMBModel()
parameters <- list(p = get_fixed())
obj <- MakeADFun(data = list(), parameters, DLL = "FIMS", silent = TRUE)


# fitting the model
#opt <- nlminb(start=obj$par, objective=obj$fn, gradient=obj$gr,
#              control = list(eval.max = 8000, iter.max = 800))

#  method = "BFGS",
#   control = list(maxit=1000000, reltol = 1e-15))
#print(opt)


#max(abs(obj$gr())) # from Cole, can use TMBhelper::fit_tmb to get val to <1e-10

#opt <- TMBhelper::fit_tmb(obj, newtonsteps=3, quiet = TRUE) # don't understand why quiet flag does not work in Quarto

#max(abs(obj$gr()))

#sdr <- TMB::sdreport(obj)
#sdr_fixed <- summary(sdr, "fixed")
#report <- obj$report()

### Plotting
