---
title: AFSC Case Study BSAI Atka mackerel
format:
  html:
    code-fold: true
editor_options: 
  chunk_output_type: console
---

## The setup

{{< include setup.qmd >}}

```{r}
#| warning: false
#| label: startup
#| output: false

theme_set(theme_bw())


# R_version <- version$version.string
# TMB_version <- packageDescription("TMB")$Version
# FIMS_commit <- substr(packageDescription("FIMS")$GithubSHA1, 1, 7)
```

-   R version: `r R_version`\
-   TMB version: `r TMB_version`\
-   FIMS commit: `r FIMS_commit`\
-   Stock name: BSAI Atka mackerel\
-   Region: AFSC\
-   Analyst: Jim Ianelli \

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the BSAI Atka mackerel stock. 

To get the opertional model to more closely match FIMS I:
* stumbled lots

## Script to prepare data for building FIMS object
```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false

## define the dimensions and global variables
#  Opent the original AMAK file
atka_dat <- readRDS(here::here("content","data_files","atka_dat.RDS"))
atka_rep <- readRDS(here::here("content","data_files","atka_rep.RDS"))
## define the dimensions and global variables
years <- atka_dat$styr:atka_dat$endyr
nyears <- length(years)
nseasons <- 1
nages <- 11
ages <- 1:nages
## nfleets <- 1
## This will fit the models bridging to FIMS (simplifying)
## source("fit_bridge_models.R")
## compare changes to model
#pkfitfinal <- readRDS("data_files/pkfitfinal.RDS")
#pkfit0 <- readRDS("data_files/pkfit0.RDS")
#parfinal <- pkfitfinal$obj$env$parList()
#pkinput0 <- readRDS('data_files/pkinput0.RDS')
#fimsdat <- pkdat0 <- pkinput0$dat
#pkinput <- readRDS('data_files/pkinput.RDS')
```




How I simplified my assessment:
* Haven't got there yet...


## Script that sets up and runs the model

```{r}
#| label: run-FIMS
#| output: false
#| eval: true

# clear memory
clear()
clear_logs()
estimate_fish_selex <- TRUE
estimate_survey_selex <- TRUE
estimate_q2 <- TRUE
estimate_q3 <- TRUE
estimate_q6 <- TRUE
estimate_F <- TRUE
estimate_recdevs <- TRUE

get_amak_data <- function(rdat,rrep){
  ## put into fims friendly form
  res <- data.frame(type = character(),
                    name = character(),
                    age = integer(),
                    datestart = character(),
                    dateend = character(),
                    value = double(),
                    unit = character(),
                    uncertainty = double())
  landings <- data.frame(type = "landings",
                         name = "fleet1",
                         age = NA,
                         datestart = paste0(seq(rdat$styr, rdat$endyr), "-01-01"),
                         dateend = paste0(seq(rdat$styr, rdat$endyr), "-12-31"),
                         value = as.numeric(rdat$catch),
                         unit = "t",
                         uncertainty = rdat$catch_cv)
  ##  need to fill missing years with -999 so it's ignored in FIMS
  indtmp <- 0*rdat$catch-999
  indtmp[which(years %in% rdat$yrs_ind )] <- rdat$biom_ind
  #indtmp
  CVtmp <- rep(1, length=nyears) # actually SE in log space
  CVtmp[which(years %in% rdat$yrs_ind)] <- rdat$biom_std/rdat$biom_ind
  ## repeat with fish catch at age, using expected in missing years
  #names(atka_rep)
  caa <- 0*rrep$N[,-1]-999
  caa[which(years %in% rdat$yrs_ages_fsh), ] <- rdat$page_fsh
  Ncaa <- rep(1, nyears)
  Ncaa[which(years %in% rdat$yrs_ages_fsh)] <- rdat$sample_ages_fsh
  paa2 <-  0*rrep$N[,-1]-999
  paa2[which(years %in% rdat$yrs_ages_ind), ] <- rdat$page_ind
  Npaa2 <- rep(1, nyears)
  Npaa2[which(years %in% rdat$yrs_ages_ind)] <- rdat$sample_ages_ind
  index <- data.frame(type = "index",
                      name = "survey",
                      age = NA,
                      datestart = paste0(seq(rdat$styr, rdat$endyr), "-01-01"),
                      dateend = paste0(seq(rdat$styr, rdat$endyr), "-12-31"),
                      value = ifelse(indtmp>0, indtmp, indtmp),
                      unit = "",
                      uncertainty = CVtmp)
  ## these have -999 for missing data years
  catchage <- data.frame(type = "age",
                         name = "fleet1",
                         age = rep(seq(1,nages), nyears),
                         datestart = rep(paste0(seq(rdat$styr, rdat$endyr), "-01-01"), each=nages),
                         dateend = rep(paste0(seq(rdat$styr, rdat$endyr), "-12-31"), each=nages),
                         value = as.numeric(t(caa)),
                         unit = "",
                         uncertainty = rep(Ncaa, each=nages))
  indexage <- data.frame(
    type = "age",
    name = "survey",
    age = rep(seq(1, nages), nyears),
    datestart = rep(paste0( seq(rdat$styr, rdat$endyr), "-01-01" ), each = nages),
    dateend = rep(paste0( seq(rdat$styr, rdat$endyr), "-12-31" ), each = nages),
    value = as.numeric(t(paa2)),
    unit = "",
    uncertainty = rep(Npaa2, each = nages)
  )
  ## indexage=indexage2
  ## index=index2
  res <- rbind(res, landings, index, catchage, indexage)
  return(res)
}
fimsdat<-get_amak_data(atka_dat,atka_rep)

```

```{r}
#| label: run-FIMS2
#| output: false
#| eval: true
#| 

age_frame <- FIMS::FIMSFrameAge(fimsdat)
fishery_catch <- FIMS::m_landings(age_frame)
fishery_agecomp <- FIMS::m_agecomp(age_frame, "fleet1")
survey_index <- FIMS::m_index(age_frame, "survey")
survey_agecomp <- FIMS::m_agecomp(age_frame, "survey")
# need to think about how to deal with multiple fleets - only using 1 fleeet for now
fish_index <- methods::new(Index, nyears)
fish_age_comp <- methods::new(AgeComp, nyears, nages)
fish_index$index_data <- fishery_catch
fish_age_comp$age_comp_data <- fishery_agecomp * fimsdat$catchage$uncertainty#rep(Ncaa, each=nages)

```
