---
title: NEFSC Case Study Southern New England-Mid Atlantic Yellowtail Flounder
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Southern New England-Mid Atlantic Yellowtail Flounder"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NEFSC
* Analyst: Chris Legault

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* End year 2021 to 2019 due to missing 2020 survey
* 5 indices to 2
* Filled a survey missing year for one index (see below)
* Varying weight at age to constant over time
* Different weight at age matrices for catch, SSB, Jan-1 to same for all 3
* SSB calculated in April to Jan-1
* Aggregate index in numbers to weight
* Fishery 2 selectivity blocks to 1
* Catch and Index selectivity at age to logistic
* Index timing from April to Jan-1
* Varying Index CV to constant over time
* Varying Catch ESS to constant over time
* SR unexploited scaler SSB to recruitment


How I simplified my assessment:

To fill the missing year of survey data, I first ran the model in ASAP with the missing data treated as missing. I then filled the missing data with the expected value of the survey, both in aggregate and for catch at age in place of the missing data.

The varying weights at age, index CVs, and ESS values were replaced by the time series mean in all years.

The catch weight at age matrix was used as the weight at age for all sources.

## Setting up the data

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# clear memory
FIMS::clear()

# read the ASAP rdat files
rdat <- dget(file.path(data_directory, "NEFSC_YT_SIMPLIFIED.RDAT")) # to be used in FIMS, lots of modifications from original
orig <- dget(file.path(data_directory, "NEFSC_YT_ORIGINAL.RDAT"))   # where started before modifications for use in FIMS

# function to create equivalent of data_mile1, basic catch and survey data
# need to think about how to deal with multiple fleets and indices - only use 1 of each for now
get_asap_data <- function(asap_input) {
  res <- data.frame(type = character(),
                name = character(),
                age = integer(),
                timing = double(),
                value = double(),
                unit = character(),
                uncertainty = double())
  years_in_model <- seq(asap_input$parms$styr, asap_input$parms$endyr)
  
  landings <- data.frame(type = "landings",
                     name = "fishery",
                     age = NA,
                     timing = years_in_model,
                     value = as.numeric(asap_input$catch.obs[1,]),
                     unit = "mt",
                     # This was
                     # asap_input$control.parms$catch.tot.cv[,1]
                     # but it was just reset to the following later in the parameters so using internal functionality to set it now
                     # uncertainty = rep(log(sqrt(log(as.numeric(mean(asap_input$control.parms$catch.tot.cv[,1], na.rm=TRUE)^2) + 1))), asap_input[["parms"]][["nyears"]])
                     uncertainty = rep(sqrt(log(as.numeric(mean(asap_input$control.parms$catch.tot.cv[,1], na.rm=TRUE)^2) + 1)), asap_input[["parms"]][["nyears"]])
  )
  
  # loop over all indices
  for (i in seq(asap_input$parms$nindices)) {
    index <- data.frame(
      type = "index",
      name = paste0("survey", i),
      age = NA_integer_,
      timing = years_in_model[asap_input$index.year.counter[[i]]],
      value = as.numeric(asap_input$index.obs[[i]]),
      unit = "mt",
      # uncertainty = rep(log(sqrt(log(as.numeric(mean(asap_input$index.cv[[i]], na.rm=TRUE)^2 + 1)))), asap_input[["parms"]][["nyears"]]))
      uncertainty = rep(
        sqrt(log(as.numeric(mean(asap_input$index.cv[[i]], na.rm=TRUE)^2 + 1))),
        length(asap_input$index.year.counter[[i]])
      )
    )
    if (i == 1){
      allinds <- index
    }else{
      allinds <- rbind(allinds, index)
    }
  }
  
  catchage <- data.frame(
    type = "age_comp",
    name = "fishery",
    age = rep(seq(1,asap_input$parms$nages), asap_input$parms$nyears),
    timing = rep(seq(asap_input$parms$styr, asap_input$parms$endyr), each = asap_input$parms$nages),
    value = as.numeric(t(asap_input$catch.comp.mats$catch.fleet1.ob)),
    unit = "",
    uncertainty = rep(asap_input$fleet.catch.Neff.init[1,], each = asap_input$parms$nages)
  )
  
  # loop over all indices
  for (i in 1:asap_input$parms$nindices){
    indexage <- data.frame(type = "age_comp",
                           name = paste0("survey", i),
                           age = rep(seq(1,asap_input$parms$nages), asap_input$parms$nyears),
                           timing = rep(seq(asap_input$parms$styr, asap_input$parms$endyr), each = asap_input$parms$nages),
                           value = as.numeric(t(asap_input$index.comp.mats[[i*2-1]])),
                           unit = "",
                           uncertainty = rep(asap_input$index.Neff.init[i,], each = asap_input$parms$nages))
    if (i == 1){
      allindsage <- indexage
    }else{
      allindsage <- rbind(allindsage, indexage)
    }
  }
  
  # Weight at age is in asap_input$WAA.mats$WAA.catch.all but for yellowtail the data
  # is time-invariant. Could just use the first row but the code below shows how
  # to use values for every year/age combination. Additionally, it adds the
  # extra year of data needed for FIMS to predict spawning biomass after all
  # catches are removed
  weight_at_age_matrix <- rbind(
      asap_input[["WAA.mats"]][["WAA.catch.all"]],
      asap_input[["WAA.mats"]][["WAA.catch.all"]][
        NROW(asap_input[["WAA.mats"]][["WAA.catch.all"]]), 
      ]
    )
  
  weight_at_age <- data.frame(
    type = "weight_at_age",
    name = NA_character_,
    age = rep(
      seq(asap_input[["parms"]][["nages"]]),
      length(seq(asap_input[["parms"]][["styr"]], asap_input[["parms"]][["endyr"]] + 1))
    ),
    timing = rep(
      seq(asap_input[["parms"]][["styr"]], asap_input[["parms"]][["endyr"]] + 1),
      each = asap_input[["parms"]][["nages"]]
    ),
    value = c(t(weight_at_age_matrix)),
    unit = "mt",
    uncertainty = NA
  )

  res <- rbind(res, landings, allinds, catchage, allindsage, weight_at_age)
  return(res)
}

# set up FIMS data objects
data_4_model <- FIMS::FIMSFrame(get_asap_data(rdat))
# define the dimensions
ages <- get_ages(data_4_model)
n_ages <- get_n_ages(data_4_model)
years <- seq(get_start_year(data_4_model), get_end_year(data_4_model))
n_years <- get_n_years(data_4_model)
```

## Run FIMS model

```{r}
#| label: setup-model
parameters <- create_default_configurations(data = data_4_model) |>
  create_default_parameters(data = data_4_model) |>
  tidyr::unnest(data) |>
  # ASAP assumes Fmult devs = 0
  dplyr::rows_update(
    tibble::tibble(
      fleet_name = c("survey1", "survey2"),
      label = "log_q",
      value = log(rdat$initial.guesses$q.year1.init)
    ),
    by = c("fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      # ASAP can enter either R0 or SB0, use R0 in input file
      label = c("log_rzero", "logit_steep"),
      value = c(
        log(rdat$initial.guesses$SR.inits$SR.scaler.init),
        FIMS::logit(0.2, 1.0, rdat$initial.guesses$SR.inits$SR_steepness.init)
      ),
      estimation_type = c("fixed_effects", "constant")
    ),
    by = c("label")
  ) |>
  # Maturity
  # NOTE, to match FIMS for a protogynous stock,
  # maturity values were obtained by fitting a logistic fcn to the age vector,
  # mat.female*prop.female + mat.male*prop.male
  # assuming an all female population 
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(1.8, 4)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = ages,
      value = log(rdat$N.age[1, ]),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = rep("log_M", n_years * n_ages),
      time = rep(years, each = n_ages),
      age = rep(ages, n_years),
      # M is vector of age1 M X nyrs then age2 M X nyrs
      value = log(as.numeric(t(rdat$M.age)))
    ),
    by = c("module_name", "label", "age", "time"),
    # TODO: fix this
    unmatched = "ignore"
  ) |>
    # Update log_sd for log_devs in the Recruitment module
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_sd",
      # ASAP takes one value per year but that value is often just repeated
      value = mean(rdat$control.parms$recruit.cv)
    ),
    by = c("module_name", "label")
  )

# Run the  model with optimization
fit <- parameters |>
  initialize_fims(data = data_4_model) |>
  fit_fims(optimize = TRUE)
output <- get_estimates(fit)

# eventually change to allow multiple fishing fleets similar to multiple indices - only using 1 fishing fleet for now
# NOTE: FIMS assumes SB calculated at the start of the year, so need to adjust ASAP to do so as well for now, need to make timing of SSB calculation part of FIMS later
# NOTE: for now tricking FIMS into thinking age 0 is age 1, so need to adjust A50 for maturity because FIMS calculations use ages 0-5+ instead of 1-6

sdr <- get_sdreport(fit)
sdr_fixed <- summary(sdr, "fixed")
report <- get_report(fit)
```

### Plotting

```{r}
#| label: comparison-plots

convert_output_asap <- function(report, fleet_names) {
  years <- seq(report$parms$styr, report$parms$endyr)
  catch <- data.frame(
    module_name = "Fleet",
    module_id = rep(seq(NROW(report$catch.pred)), each = length(years)),
    label = "landings_expected",
    year_i = rep(seq(NCOL(report$catch.pred)), NROW(report$catch.pred)),
    estimate = c(report$catch.pred),
    estimation_type = "derived_quantity"
  )
  index <- data.frame(
    module_name = "Fleet",
    module_id = rep(seq(report$index.pred) + NROW(report$catch.pred), each = length(years)),
    label = "indices_expected",
    year_i = unlist(report$index.year.counter),
    estimate = unlist(report$index.pred),
    estimation_type = "derived_quantity"
  )
  index_observed <- index |>
    dplyr::mutate(
      label = "indices_observed",
      estimate = unlist(report$index.obs)
    )
  spawning_biomass <- data.frame(
    module_name = "Population",
    label = "spawning_biomass",
    year_i = seq(years),
    estimate = report$SSB * 0.5,
    estimation_type = "derived_quantity"
  )
  expected_recruitment <- data.frame(
    module_name = "Population",
    label = "expected_recruitment",
    year_i = seq(years),
    estimate = as.numeric(report$N.age[, 1]),
    estimation_type = "derived_quantity"
  )
  # I don't know the structure of F.report when more than one fleet is present
  log_Fmort <- data.frame(
    module_name = "Fleet",
    module_id = 1,
    label = "log_Fmort",
    type = "vector",
    year_i = seq(years),
    estimate = log(report$F.report),
    estimation_type = "fixed_effects"
  )
  log_init_naa <- data.frame(
    module_name = "Population",
    label = "log_init_naa",
    type = "vector",
    age_i = rep(
      seq(NCOL(report$N.age)),
      NROW(report$N.age)
    ),
    year_i = rep(
      seq(NROW(report$N.age)),
      each = NCOL(report$N.age)
    ),
    estimate = log(c(t(report$N.age))),
    estimation_type = "fixed_effects"
  )
  dplyr::bind_rows(
    catch, index, index_observed,
    spawning_biomass, expected_recruitment, log_Fmort, log_init_naa
  ) |>
    dplyr::left_join(
      data.frame(
        year = years,
        year_i = seq(years)
      ),
      by = "year_i"
    ) |>
    dplyr::mutate(
      uncertainty = NA_real_,
      uncertainty_label = "se",
      fleet = as.character(factor(module_id, labels = fleet_names)),
      era = "time"
    )
}
output_asap <- convert_output_asap(rdat, get_fleets(data_4_model))
output_fims <- output |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + FIMS::get_start_year(data_4_model) - 1,
    estimate = estimated,
    era = "time"
  )
stockplotr::plot_spawning_biomass(
  list(
    "ASAP" = output_asap,
    "FIMS" = output_fims
  )
)
stockplotr::plot_recruitment(
  list(
    "ASAP" = output_asap,
    "FIMS" = output_fims
  ),
  facet = "model"
)

# Not showing this plot because the hard-coded one is better
# stockplotr::plot_indices(
#   dat = list(
#     "ASAP" = output_asap,
#     "FIMS" = dplyr::mutate(output_fims,
#       label = dplyr::if_else(label == "index_expected", "indices_expected", label),
#       fleet = as.character(factor(module_id, labels = get_fleets(data_4_model)))
#     ) |>
#       dplyr::filter(!is.na(observed))
#   )
# )

mycols <- c("FIMS" = "blue", "ASAP" = "red", "ASAP_orig" = "darkgreen")
for (i in seq(rdat$parms$nindices)) {
  index_results <- data.frame(
    survey = i,
    year = years,
    observed = rdat$index.obs[[i]],
    FIMS = report$index_expected[[rdat$parms$nfleet+i]],
    ASAP = rdat$index.pred[[i]]
  )
  if (i==1){
    allinds_results <- index_results
  }else{
    allinds_results <- rbind(allinds_results, index_results)
  }
}
#print(allinds_results)

comp_index <- ggplot2::ggplot(allinds_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::geom_line(ggplot2::aes(x = year, y = FIMS), color = "blue") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = ASAP), color = "red") +
  ggplot2::facet_wrap(~survey, scales = "free_y", nrow = 2) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Index") +
  ggplot2::ggtitle("Blue=FIMS, Red=ASAP") +
  ggplot2::theme_bw()
print(comp_index)

catch_results <- data.frame(
  observed = get_data(data_4_model) |>
    dplyr::filter(type == "landings", name == "fishery") |>
    dplyr::pull(value),
  FIMS = report$landings_expected[[1]],
  ASAP = as.numeric(rdat$catch.pred[1,])
)
#print(catch_results)

comp_catch <- ggplot2::ggplot(catch_results, ggplot2::aes(x = years, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = years, y = FIMS), color = "blue") +
  ggplot2::geom_line(ggplot2::aes(x = years, y = ASAP), color = "red") +
  ggplot2::ggtitle("Blue=FIMS, Red=ASAP") +
  ggplot2::theme_bw()
print(comp_catch)

pop_results <- data.frame(
  Year = c(years, max(years)+1, years, years, years, years, max(years)+1, years),
  Metric = c(rep("Spawning biomass", 2*n_years+1), rep("Fishing mortality", 2*n_years), rep("Recruitment", 2*n_years+1)),
  Model = c(rep("FIMS", n_years+1), rep("ASAP", n_years), rep(c("FIMS", "ASAP"), each = n_years), 
             rep("FIMS", n_years+1), rep("ASAP", n_years)),
  Value = c(
    report$spawning_biomass[[1]],
    rdat$SSB / 2,
    output |>
      dplyr::filter(label == "log_Fmort", module_id == 1) |>
      dplyr::pull(estimated) |>
      exp(),
    rdat$F.report,
    report$expected_recruitment[[1]],
    as.numeric(rdat$N.age[,1])
  )
)
#print(pop_results)

ggplot2::ggplot(dplyr::filter(pop_results, Year <=2019), ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Metric, ncol=1, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)

orig_years <- seq(orig$parms$styr, orig$parms$endyr)
orig_pop_results <- data.frame(
  Year = rep(orig_years, 3),
  Metric = rep(c("Spawning biomass", "Fishing mortality", "Recruitment"), each = length(orig_years)),
  Model = "ASAP_orig",
  Value = c(orig$SSB / 2, orig$F.report, as.numeric(orig$N.age[,1]))
)

pop_results_3 <- rbind(pop_results, orig_pop_results)
#print(pop_results_3)

comp_FRSSB3 <- ggplot2::ggplot(pop_results_3, ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Metric, ncol=1, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)
print(comp_FRSSB3)

FIMS_naa_results <- data.frame(
  Year = rep(c(years, max(years)+1), each = n_ages),
  Age = rep(ages, n_years+1),
  Metric = "NAA",
  Model = "FIMS",
  Value = report$numbers_at_age[[1]]
)

ASAP_naa_results <- data.frame(
  Year = rep(years, each = n_ages),
  Age = rep(ages, n_years),
  Metric = "NAA",
  Model = "ASAP",
  Value = as.numeric(t(rdat$N.age))
)

orig_naa_results <- data.frame(
  Year = rep(orig_years, each = n_ages),
  Age = rep(ages, length(orig_years)),
  Metric = "NAA",
  Model = "ASAP_orig",
  Value = as.numeric(t(orig$N.age))
)
naa_results <- rbind(FIMS_naa_results, ASAP_naa_results, orig_naa_results)
#print(naa_results)

comp_naa3 <- ggplot2::ggplot(dplyr::filter(naa_results, Year <= 2019), ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Age, ncol=1, scales = "free_y") +
  ggplot2::ylab("NAA") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)
print(comp_naa3)
```

## Comparison table

The likelihood components from FIMS and ASAP for the same data are shown in the table below. Note that the ASAP file had to turn on the use likelihood constants option to enable this comparison (this option should not be used when recruitment deviations are estimated).

```{r}
#| label: comparison-tables
nll_components <- report[["nll_components"]]
jnlltab <- data.frame(
  Component=c("Total","Index","Age Comp", "Rec"),
  FIMS = c(
    report[["jnll"]],
    sum(nll_components[seq(2, length(nll_components), by = 2)]),
    sum(nll_components[seq(3, length(nll_components), by = 2)]),
    nll_components[1]
  ),
  ASAP = c(
    rdat$like$lk.total,
    (rdat$like$lk.catch.total + rdat$like$lk.index.fit.total),
    (rdat$like$lk.catch.age.comp + rdat$like$lk.index.age.comp),
    rdat$like$lk.Recruit.devs
  )
)
print(jnlltab)
```

## List any issues that you ran into or found

Please [open an issue](https://github.com/NOAA-FIMS/FIMS/issues/new/choose) if you found something new.

* SB calculations in FIMS assume 0.5 multiplier, which differs from ASAP [Issue #521](https://github.com/NOAA-FIMS/FIMS/issues/521)
* Separate weights at age for catch, SSB, Jan-1 population, indices, etc.
* Fishery selectivity blocks or random effects
* Allow time-varying CVs and ESS (or alternative functions)
* Timing of Index and SSB calculations within the year
* One-step-ahead residuals
* Reference points

The following items are available in FIMS but have yet to be added to this case study

* Projections
* Option for Index in numbers
* Push-button retro
* Inclusion of the other 3 indices because missing values are automatically filled in now
* Include the true time-varying weight-at-age data

```{r}
#| label: cleanup
#| echo: false
# Clear C++ objects from memory
FIMS::clear()
```
