---
title: Southern New England-Mid Atlantic Yellowtail Flounder
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
#| include: false
#| echo: false
common_name <- "Southern New England-Mid Atlantic Yellowtail Flounder"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NEFSC
* Analyst: Chris Legault

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational assessment model that was presented to management in 2022 for `r common_name` ([NEFSC, 2022](https://doi.org/10.25923/380j-t283)) and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration for comparing the results of a FIMS model to the results of an Age-Structured Assessment Program (ASAP) model.

To get the operational model to more closely match what is capable in a FIMS model, the following changes were made:

* Changed the timing of the surveys from from April to January 1.
* Changed the timing of when spawning biomass is calculated from April to January 1.
* Used the catch weight-at-age matrix for weight-at-age data and ignored the matrices for spawning biomass and the timing of the surveys.
* Changed the selectivity for both the fishing fleet and survey fleets to logistic.
* Changed from an unexploited scaler on spawning biomass to recruitment.
* Spawning biomass calculations in FIMS assume 0.5 multiplier, which differs from ASAP [Issue #521](https://github.com/NOAA-FIMS/FIMS/issues/521).
* Turned on the use likelihood constants in ASAP to enable the comparison of likelihood components, an option that should not be used because recruitment deviations are estimated.
* One-step-ahead residuals do not exist yet in FIMS.
* Reference points are not automatically calculated and could not be compared.

The following additional changes were made as well though they are not necessary:

* End year 2021 to 2019 due to missing 2020 survey.
* Removed 3 index of abundance fleets because they had years with missing data.
* Removed the discard fishing fleet.
* Filled a survey missing year for one index using fitted estimate from ASAP.
* Used the mean weight at age across all years instead of allowing weight-at-age data to vary with time.
* Changed the aggregate index unit from numbers to weight.
* Removed the time blocks on selectivity for the fishing fleet.
* Used the time-series mean of the coefficient of variation for each index as the measurement of uncertainty.
* Used the time-series mean of effective sample size for age-composition data for each fleet.
* Projections were removed from the comparison.

Missing years of survey data were filled in using predictions from the simplified ASAP model. The ASAP model was first ran with the missing data treated as missing. Then the missing data was filled in with the expected value of the survey, both in aggregate and for catch at age in place of the missing data.

## Setting up the data

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# read the ASAP rdat files
rdat <- dget(file.path(data_directory, "NEFSC_YT_SIMPLIFIED.RDAT")) # to be used in FIMS, lots of modifications from original
orig <- dget(file.path(data_directory, "NEFSC_YT_ORIGINAL.RDAT"))   # where started before modifications for use in FIMS

# set up FIMS data objects
data_4_model <- FIMS::FIMSFrame(get_asap_data(rdat))
# define the dimensions
years <- seq(get_start_year(data_4_model), get_end_year(data_4_model))
```

## Run FIMS model

```{r}
#| label: setup-model
# clear memory
FIMS::clear()

parameters <- create_default_configurations(data = data_4_model) |>
  create_default_parameters(data = data_4_model) |>
  tidyr::unnest(data) |>
  # ASAP assumes Fmult devs = 0
  dplyr::rows_update(
    tibble::tibble(
      fleet_name = c("survey1", "survey2"),
      label = "log_q",
      value = log(rdat$initial.guesses$q.year1.init)
    ),
    by = c("fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      # ASAP can enter either R0 or SB0, use R0 in input file
      label = c("log_rzero", "logit_steep"),
      value = c(
        log(rdat$initial.guesses$SR.inits$SR.scaler.init),
        FIMS::logit(0.2, 1.0, rdat$initial.guesses$SR.inits$SR_steepness.init)
      ),
      estimation_type = c("fixed_effects", "constant")
    ),
    by = c("label")
  ) |>
  # Maturity
  # NOTE, to match FIMS for a protogynous stock,
  # maturity values were obtained by fitting a logistic fcn to the age vector,
  # mat.female*prop.female + mat.male*prop.male
  # assuming an all female population 
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(1.8, 4)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = get_ages(data_4_model),
      value = log(rdat$N.age[1, ]),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = rep("log_M", get_n_years(data_4_model) * get_n_ages(data_4_model)),
      time = rep(years, each = get_n_ages(data_4_model)),
      age = rep(get_ages(data_4_model), get_n_years(data_4_model)),
      # M is vector of age1 M X nyrs then age2 M X nyrs
      value = log(as.numeric(t(rdat$M.age)))
    ),
    by = c("module_name", "label", "age", "time"),
    # TODO: fix this
    unmatched = "ignore"
  ) |>
    # Update log_sd for log_devs in the Recruitment module
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_sd",
      # ASAP takes one value per year but that value is often just repeated
      value = mean(rdat$control.parms$recruit.cv)
    ),
    by = c("module_name", "label")
  )

# Run the  model with optimization
fit <- parameters |>
  initialize_fims(data = data_4_model) |>
  fit_fims(optimize = TRUE)
output <- get_estimates(fit)

# eventually change to allow multiple fishing fleets similar to multiple indices - only using 1 fishing fleet for now
# NOTE: FIMS assumes SB calculated at the start of the year, so need to adjust ASAP to do so as well for now, need to make timing of SSB calculation part of FIMS later
# NOTE: for now tricking FIMS into thinking age 0 is age 1, so need to adjust A50 for maturity because FIMS calculations use ages 0-5+ instead of 1-6

sdr <- get_sdreport(fit)
sdr_fixed <- summary(sdr, "fixed")
report <- get_report(fit)
```

### Plotting

```{r}
#| label: comparison-plots

convert_output_asap <- function(report, fleet_names) {
  years <- seq(report$parms$styr, report$parms$endyr)
  catch <- data.frame(
    module_name = "Fleet",
    module_id = rep(seq(NROW(report$catch.pred)), each = length(years)),
    label = "landings_expected",
    year_i = rep(seq(NCOL(report$catch.pred)), NROW(report$catch.pred)),
    estimate = c(report$catch.pred),
    estimation_type = "derived_quantity"
  )
  index <- data.frame(
    module_name = "Fleet",
    module_id = rep(seq(report$index.pred) + NROW(report$catch.pred), each = length(years)),
    label = "indices_expected",
    year_i = unlist(report$index.year.counter),
    estimate = unlist(report$index.pred),
    estimation_type = "derived_quantity"
  )
  index_observed <- index |>
    dplyr::mutate(
      label = "indices_observed",
      estimate = unlist(report$index.obs)
    )
  spawning_biomass <- data.frame(
    module_name = "Population",
    label = "spawning_biomass",
    year_i = seq(years),
    estimate = report$SSB * 0.5,
    estimation_type = "derived_quantity"
  )
  expected_recruitment <- data.frame(
    module_name = "Population",
    label = "expected_recruitment",
    year_i = seq(years),
    estimate = as.numeric(report$N.age[, 1]),
    estimation_type = "derived_quantity"
  )
  # I don't know the structure of F.report when more than one fleet is present
  log_Fmort <- data.frame(
    module_name = "Fleet",
    module_id = 1,
    label = "log_Fmort",
    type = "vector",
    year_i = seq(years),
    estimate = log(report$F.report),
    estimation_type = "fixed_effects"
  )
  log_init_naa <- data.frame(
    module_name = "Population",
    label = "log_init_naa",
    type = "vector",
    age_i = rep(
      seq(NCOL(report$N.age)),
      NROW(report$N.age)
    ),
    year_i = rep(
      seq(NROW(report$N.age)),
      each = NCOL(report$N.age)
    ),
    estimate = log(c(t(report$N.age))),
    estimation_type = "fixed_effects"
  )
  dplyr::bind_rows(
    catch, index, index_observed,
    spawning_biomass, expected_recruitment, log_Fmort, log_init_naa
  ) |>
    dplyr::left_join(
      data.frame(
        year = years,
        year_i = seq(years)
      ),
      by = "year_i"
    ) |>
    dplyr::mutate(
      uncertainty = NA_real_,
      uncertainty_label = "se",
      fleet = as.character(factor(module_id, labels = fleet_names)),
      era = "time"
    )
}
output_asap <- convert_output_asap(rdat, get_fleets(data_4_model))
output_fims <- output |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + FIMS::get_start_year(data_4_model) - 1,
    estimate = estimated,
    era = "time"
  )
stockplotr::plot_spawning_biomass(
  list(
    "ASAP" = output_asap,
    "FIMS" = output_fims
  )
)
stockplotr::plot_recruitment(
  list(
    "ASAP" = output_asap,
    "FIMS" = output_fims
  ),
  facet = "model"
)

# Not showing this plot because the hard-coded one is better
# stockplotr::plot_indices(
#   dat = list(
#     "ASAP" = output_asap,
#     "FIMS" = dplyr::mutate(output_fims,
#       label = dplyr::if_else(label == "index_expected", "indices_expected", label),
#       fleet = as.character(factor(module_id, labels = get_fleets(data_4_model)))
#     ) |>
#       dplyr::filter(!is.na(observed))
#   )
# )

mycols <- c("FIMS" = "blue", "ASAP" = "red", "ASAP_orig" = "darkgreen")
for (i in seq(rdat$parms$nindices)) {
  index_results <- data.frame(
    survey = i,
    year = years,
    observed = rdat$index.obs[[i]],
    FIMS = report$index_expected[[rdat$parms$nfleet+i]],
    ASAP = rdat$index.pred[[i]]
  )
  if (i==1){
    allinds_results <- index_results
  }else{
    allinds_results <- rbind(allinds_results, index_results)
  }
}
#print(allinds_results)

comp_index <- ggplot2::ggplot(allinds_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::geom_line(ggplot2::aes(x = year, y = FIMS), color = "blue") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = ASAP), color = "red") +
  ggplot2::facet_wrap(~survey, scales = "free_y", nrow = 2) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Index") +
  ggplot2::ggtitle("Blue=FIMS, Red=ASAP") +
  ggplot2::theme_bw()
print(comp_index)

catch_results <- data.frame(
  observed = get_data(data_4_model) |>
    dplyr::filter(type == "landings", name == "fishery") |>
    dplyr::pull(value),
  FIMS = report$landings_expected[[1]],
  ASAP = as.numeric(rdat$catch.pred[1,])
)
#print(catch_results)

comp_catch <- ggplot2::ggplot(catch_results, ggplot2::aes(x = years, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = years, y = FIMS), color = "blue") +
  ggplot2::geom_line(ggplot2::aes(x = years, y = ASAP), color = "red") +
  ggplot2::ggtitle("Blue=FIMS, Red=ASAP") +
  ggplot2::theme_bw()
print(comp_catch)

pop_results <- data.frame(
  Year = c(years, max(years)+1, years, years, years, years, max(years)+1, years),
  Metric = c(rep("Spawning biomass", 2*get_n_years(data_4_model)+1), rep("Fishing mortality", 2*get_n_years(data_4_model)), rep("Recruitment", 2*get_n_years(data_4_model)+1)),
  Model = c(rep("FIMS", get_n_years(data_4_model)+1), rep("ASAP", get_n_years(data_4_model)), rep(c("FIMS", "ASAP"), each = get_n_years(data_4_model)), 
             rep("FIMS", get_n_years(data_4_model)+1), rep("ASAP", get_n_years(data_4_model))),
  Value = c(
    report$spawning_biomass[[1]],
    rdat$SSB / 2,
    output |>
      dplyr::filter(label == "log_Fmort", module_id == 1) |>
      dplyr::pull(estimated) |>
      exp(),
    rdat$F.report,
    report$expected_recruitment[[1]],
    as.numeric(rdat$N.age[,1])
  )
)
#print(pop_results)

ggplot2::ggplot(dplyr::filter(pop_results, Year <=2019), ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Metric, ncol=1, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)

orig_years <- seq(orig$parms$styr, orig$parms$endyr)
orig_pop_results <- data.frame(
  Year = rep(orig_years, 3),
  Metric = rep(c("Spawning biomass", "Fishing mortality", "Recruitment"), each = length(orig_years)),
  Model = "ASAP_orig",
  Value = c(orig$SSB / 2, orig$F.report, as.numeric(orig$N.age[,1]))
)

pop_results_3 <- rbind(pop_results, orig_pop_results)
#print(pop_results_3)

comp_FRSSB3 <- ggplot2::ggplot(pop_results_3, ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Metric, ncol=1, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)
print(comp_FRSSB3)

FIMS_naa_results <- data.frame(
  Year = rep(c(years, max(years)+1), each = get_n_ages(data_4_model)),
  Age = rep(get_ages(data_4_model), get_n_years(data_4_model)+1),
  Metric = "NAA",
  Model = "FIMS",
  Value = report$numbers_at_age[[1]]
)

ASAP_naa_results <- data.frame(
  Year = rep(years, each = get_n_ages(data_4_model)),
  Age = rep(get_ages(data_4_model), get_n_years(data_4_model)),
  Metric = "NAA",
  Model = "ASAP",
  Value = as.numeric(t(rdat$N.age))
)

orig_naa_results <- data.frame(
  Year = rep(orig_years, each = get_n_ages(data_4_model)),
  Age = rep(get_ages(data_4_model), length(orig_years)),
  Metric = "NAA",
  Model = "ASAP_orig",
  Value = as.numeric(t(orig$N.age))
)
naa_results <- rbind(FIMS_naa_results, ASAP_naa_results, orig_naa_results)
#print(naa_results)

comp_naa3 <- ggplot2::ggplot(dplyr::filter(naa_results, Year <= 2019), ggplot2::aes(x=Year, y=Value, color=Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Age, ncol=1, scales = "free_y") +
  ggplot2::ylab("NAA") +
  ggplot2::theme_bw() +
  ggplot2::scale_color_manual(values = mycols)
print(comp_naa3)
```

## Comparison table

The likelihood components from FIMS and ASAP for the same data are shown in the table below. Note that the ASAP file had to turn on the use likelihood constants option to enable this comparison (this option should not be used when recruitment deviations are estimated).

```{r}
#| label: comparison-tables
nll_components <- report[["nll_components"]]
jnlltab <- data.frame(
  Component=c("Total","Index","Age Comp", "Rec"),
  FIMS = c(
    report[["jnll"]],
    sum(nll_components[seq(2, length(nll_components), by = 2)]),
    sum(nll_components[seq(3, length(nll_components), by = 2)]),
    nll_components[1]
  ),
  ASAP = c(
    rdat$like$lk.total,
    (rdat$like$lk.catch.total + rdat$like$lk.index.fit.total),
    (rdat$like$lk.catch.age.comp + rdat$like$lk.index.age.comp),
    rdat$like$lk.Recruit.devs
  )
)
print(jnlltab)
```

```{r}
#| label: cleanup
#| include: false
#| echo: false
# Clear C++ objects from memory
FIMS::clear()
```
