---
title: NWFSC Case Study Petrale Sole
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Petrale sole"


# read SS3 input files from petrale sole assessment on GitHub
petrale_input <- r4ss::SS_read(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a034.001/"
)
# add wtatage file to list of input files
petrale_input[["wtatage"]] <- r4ss::SS_readwtatage(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/refs/heads/main/models/2023.a034.001/wtatage.ss_new"
)
# read SS3 output files from petrale sole assessment on GitHub
petrale_output <- r4ss::SS_output(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a034.001",
  printstats = FALSE,
  verbose = FALSE,
  covar = FALSE
)
output_petrale <- stockplotr::convert_output(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a034.001/Report.sso",
  model = "ss3"
)
# years <- seq(petrale_input$dat$styr, petrale_input$dat$endyr)
years <- seq(1896, petrale_input$dat$endyr)
n_years <- length(years)
# ages <- 0:petrale_input$dat$Nages # population ages in SS3, starts at age 0
ages <- 1:17 # same as data bins
n_ages <- length(ages)
lengths <- petrale_output[["lbins"]]
nlengths <- length(lengths)
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* Remove data
  * Remove male lengths and ages
  * Remove discard fractions and discard comps
* Simplify selectivity
  * Remove length-based retention functions
  * Convert to age-based logistic (from length-based double-normal, fixed asymptotic)
* Remove priors
  * Natural mortality
  * Steepness
* Use female mean weight at age from the SS3 weight-at-age file

## Setting up the data

```{r}
#| eval: false
#| label: local-production-model
#| echo: false

## SIMPLIFY SS3 DATA FILE
dat <- petrale_input$dat

# change dimensions
dat$Nsexes <- 1 # previously 2
dat$Nages <- n_ages # previously 40

# mirroring selectivity and keeping both fishing fleets instead of aggregating
# catch aggregate landings across fleets
# dat$catch <- dat$catch |>
#     dplyr::group_by(year) |>
#     dplyr::summarize(catch = sum(catch), catch_se = mean(catch_se)) |>
#     dplyr::mutate(seas = 1, fleet = 1, .after = year) |>
#     as.data.frame() # write_fwf4() can't write tibles

# filter indices (just include WCGBTS)
dat$CPUE <- dat$CPUE |>
  dplyr::filter(index == 4)

# remove discard info
dat$N_discard_fleets <- 0
dat$discard_fleet_info <- NULL
dat$discard_data <- NULL

# remove mean body weight
dat$use_meanbodywt <- 0
dat$meanbodywt <- NULL
dat$DF_for_meanbodywt <- NULL

# remove length comps
dat$use_lencomp <- 0
dat$len_info <- NULL
dat$lencomp <- NULL

# simplify ageing error matrix
dat$N_ageerror_definitions <- 1
dat$ageerror <- dat$ageerror[1:2, paste0("age", 0:17)]

# simplify age comps
dat$agecomp <-
  dat$agecomp |>
  # exclude fleets 2 and 3, and only use marginal data for fleet 4
  dplyr::filter(fleet %in% c(1, -4)) |>
  # exclude male observations
  dplyr::select(!dplyr::starts_with("m", ignore.case = FALSE)) |>
  # assign observations to ageing error 1
  dplyr::mutate(Gender = 1, Ageerr = 1) |>
  # remove redundant observations
  # NOTE: this removes some age comp data because there
  # were years with multiple observations from the same fleet
  # due to multiple ageing error matrices
  dplyr::distinct(year, fleet, .keep_all = TRUE)

# Simplify SS3 control file
ctl <- petrale_input$ctl
ctl$Nages <- 17 # previously 40
ctl$Nsexes <- 1 # previously 2
# turn on empirical weight-at-age
# ctl$EmpiricalWAA <- 1 # previously zero
# remove male mortality and growth parameters
ctl$MG_parms <- ctl$MG_parms |>
  dplyr::filter(!grepl("_Mal_", rownames(ctl$MG_parms)))
ctl$max_bias_adj <- -1 # set bias adjust = 1.0 for all years

# filter catchability stuff to only fleet 4
ctl$Q_options <- ctl$Q_options |>
  dplyr::filter(fleet == 4)
ctl$Q_parms <- ctl$Q_parms |>
  dplyr::filter(grepl("WCGBTS", rownames(ctl$Q_parms)))

# add new age-based selex parameters
# age-based logistic (pattern 12) for fleets 1 and 4
# have south fishery (fleet 2) mirror fleet 1
# no parameters for fleet 3 (ignored index left in place to avoid renumbering fleets)
ctl$age_selex_types <- ctl$age_selex_types |>
  dplyr::mutate(
    Pattern = c(12, 15, 0, 12),
    Special = c(0, 1, 0, 0)
  ) # fleet 2 mirror fleet 1

# create 4 parameter rows
ctl$age_selex_parms <- cbind(
  data.frame(
    LO = 1,
    HI = 10,
    INIT = c(5, 2, 5, 2),
    PRIOR = 0,
    PR_SD = 99,
    PR_type = 0,
    PHASE = 2
  ),
  matrix(0, nrow = 4, ncol = 7)
)
names(ctl$age_selex_parms)[8:14] <- names(ctl$size_selex_parms)[8:14]
rownames(ctl$age_selex_parms) <- paste0(
  "AgeSel_", c("P_1_", "P_2_"),
  c("fleet1", "fleet1", "fleet4", "fleet4")
)

# remove length-based selectivity
ctl$size_selex_types <- ctl$size_selex_types |>
  dplyr::mutate(Pattern = 0, Discard = 0, Male = 0)
ctl$size_selex_parms <- NULL
ctl$size_selex_parms_tv <- NULL

# turn off variance adjustments
ctl$DoVar_adjust <- 0
ctl$Variance_adjustment_list <- NULL
# use wtatage file
ctl$EmpiricalWAA <- 1

# Create wtatage.ss file
# wtatage <- SS_readwtatage(
#   "c:/ss/Petrale/Petrale2023/petrale/models//2023.a034.001/wtatage.ss_new"
# )
# wtatage for fleets -1:0 (population), and fleets 1:4 (real fleets)
wtatage <- matrix(
  rep(
    c(
      0.0010, # age 0
      0.0148, 0.0617, 0.1449, 0.2570, 0.3876, 0.5260, 0.6640, 0.7957, 0.9175,
      1.0273, 1.1247, 1.2097, 1.2831, 1.3460, 1.3994, 1.4446, 1.4821
    ),
    6
  ),
  nrow = 6, ncol = 18, byrow = TRUE
)
# wtatage for fleet = -2 (fecundity * maturity)
wtatage <- rbind(
  c(
    0, 0, 0, 2.34337e-07, 6.7895e-06, 5.41665e-05, 0.000173396, 0.000338953,
    0.000514454, 0.000681685, 0.000834535, 0.000971472, 0.00109252, 0.00119837,
    0.00129006, 0.00136882, 0.00143602, 0.00149302
  ),
  wtatage
)
names(wtatage) <- 0:17
wtatage <- data.frame(
  Yr = -1800,
  Seas = 1,
  Sex = 1,
  Bio_Pattern = 1,
  BirthSeas = 1,
  Fleet = -2:4,
  wtatage
)

if (TRUE) {
  # write modified data and control files
  new_input <- petrale_input
  new_input$dat <- dat
  new_input$ctl <- ctl
  new_input$wtatage <- wtatage

  # new_input$dir <- "c:/ss/Petrale/Petrale2023/petrale/models/2023.a050.001_FIMS_case-study"
  new_input$dir <- "c:/ss/Petrale/Petrale2023/petrale/models/2023.a050.002_FIMS_case-study_wtatage"
  # create directory if it doesn't exist
  if (!dir.exists(new_input$dir)) {
    dir.create(new_input$dir)
  }

  r4ss::SS_write(new_input, dir = new_input$dir, overwrite = TRUE)

  r4ss::run(
    new_input$dir,
    show_in_console = TRUE,
    skipfinished = FALSE
  )
  p2 <- r4ss::SS_output(new_input$dir)
}
```

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false
  
# define the dimensions based on the range in the data
# for petrale, the north fleet catch begins in 1896, 
# not the start year of the model 1876 (which is when the southern fleet catch begins,
# but for simplicity this petrale example only uses the northern catch)

# fleet 4 used conditional age-at-length data with marginal observations
# entered as fleet == -4 (to exclude from likelihood due to redundancy)
# using only marginals for FIMS and exclude CAAL data by filtering out
# the fleet 4 age data
# similarly, some early ages were excluded from fleet 1 in the original model
# by assigning to fleet -1, these get filtered by get_ss3_data()

# only include age comps with fleet = -4 or 1:
petrale_input$dat$agecomp <- petrale_input$dat$agecomp |>
  dplyr::filter(fleet %in% c(-4, 1)) |>
  dplyr::mutate(fleet = abs(fleet)) |>
  dplyr::arrange(fleet, year, dplyr::desc(Nsamp)) |>
  dplyr::distinct(fleet, year, .keep_all = TRUE)

petrale_input$dat$lencomp <- petrale_input$dat$lencomp |>
    dplyr::filter(
    # filter out discard length-composition data
    !(part == 1 & fleet == 1)
  )

# convert SS3 data into FIMS format using function defined in the R directory
mydat <- get_ss3_data(
  ss3_inputs = petrale_input,
  fleets = c(1, 4),
  ages = ages,
  lengths = petrale_input[["dat"]][["lbin_vector"]]
) |>
  # rename fleet4 as fleet2 (fleets 2 and 3 already removed above)
  dplyr::mutate(name = dplyr::case_when(
    name == "fleet1" ~ name,
    name == "fleet4" ~ "survey1" # change fleet4 to survey1 to match yellowtail case-study
  )) |>
  dplyr::filter(
    # filter out forecast years
    timing %in% years
  )


# set up FIMS data objects
data_4_model <- FIMS::FIMSFrame(mydat |>
  # Remove the length-composition data
  # TODO: remove this as we create an age-to-length-conversion matrixt
  dplyr::filter(type != "length_comp") |>
  dplyr::select(-length)
)
```

## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

# lots of code below copied PIFSC-opakapaka.qmd
# rename SS3 output to avoid modifying lines copied from opakapaka model
rep <- petrale_output

# Create default parameters
default_parameters <- FIMS::create_default_configurations(
  data = data_4_model
) |>
  FIMS::create_default_parameters(
    data = data_4_model
  ) |>
  tidyr::unnest(cols = data)

recdevs <- rep$recruitpars |> # $recruitpars is a subset of $parameters with Yr as an additional column
  dplyr::filter(Yr %in% years[-1]) |> # filter out initial numbers parameters
  dplyr::select(Yr, Value)

init_naa <- rep$natage |>
  dplyr::filter(Time == min(years), Sex == 1) |>  # numbers at age in FIMS start year, not SS3 model
  dplyr::select(paste(ages)) |> 
  as.numeric() # TODO: figure out if *1000 is needed
init_naa <- exp(rep$parameters |> dplyr::filter(grepl("R0", Label)) |> dplyr::pull(Init)) * 1000 * exp(-(ages - 1) * rep$parameters |> dplyr::filter(grepl("NatM.*Fem", Label)) |> dplyr::pull(Value))
init_naa[n_ages] <- init_naa[n_ages] / rep$parameters |> dplyr::filter(grepl("NatM.*Fem", Label)) |> dplyr::pull(Value) # sum of infinite series

parameters <- default_parameters |>
  # SS3 model had length-based selectivity which leads to sex-specific
  # age-based selectivity due to sexually-dimorphic growth.
  # I didn't bother to calculate an age-based inflection point averaged over sexes
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(20.5, 1.8), # arbitrary guess of slope
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "fleet1",
      label = c("inflection_point", "slope"),
      value = c(10, 2),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "survey1",
      label = c("inflection_point", "slope"),
      value = c(6, 2),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  )  |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_rzero", "logit_steep", "log_sd"),
      value = c(
        petrale_input$ctl$SR_parms["SR_LN(R0)", "INIT"],
        FIMS::logit(0.2, 1.0, petrale_input$ctl$SR_parms["SR_BH_steep", "INIT"]),
        0.5
      )
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_devs",
      time = (FIMS::get_start_year(data_4_model) + 1):FIMS::get_end_year(data_4_model),
      # The last value of the initial numbers at age is the first
      # recruitment deviation
      value = recdevs$Value,
    ),
    by = c("module_name", "label", "time")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = FIMS::get_ages(data_4_model),
      value = log(init_naa),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_M",
      value = log(rep$parameters["NatM_uniform_Fem_GP_1", "Value"]),
      estimation_type = "constant"
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet1",
      time = years,
      label = "log_Fmort",
      estimation_type = "constant",
      value = rep[["exploitation"]] |> 
        dplyr::filter(Yr %in% years) |>
        dplyr::mutate(North = ifelse(North == 0, -999, log(North))) |>
        dplyr::pull(North)
    ),
    by = c("module_name", "fleet_name", "label", "time")
  )

# Run the model without optimization to help ensure a viable model
test_fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = FALSE)

# Run the  model with optimization
fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)
```

## Plotting results

```{r}
#| label: comparison-plots
# gather index fit info
index_results <- data.frame(
  observed = FIMS::m_index(data_4_model, "survey1"),
  expected = FIMS::get_report(fit)[["index_expected"]][[2]]
) |>
  dplyr::mutate(year = years) |>
  dplyr::filter(observed > 0) # filter out -999 rows

# plot index fit
ggplot2::ggplot(index_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Index (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected), color = "blue") +
  ggplot2::theme_bw()


# gather catch info 
catch_results <- data.frame(
  year = years,
  observed = FIMS::m_landings(data_4_model, fleet = "fleet1"),
  expected = FIMS::get_report(fit)[["landings_expected"]][[1]]
)

# plot catch fit
ggplot2::ggplot(catch_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  #ggplot2::aes(color = fleet)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected)) + #, color = fleet)) +
  ggplot2::theme_bw()

# gather biomass info from FIMS and SS3
biomass <- rep$timeseries |>
  dplyr::select(Yr, SpawnBio, Bio_all) |> 
  dplyr::filter(Yr %in% years) |> 
  dplyr::rename(
    "SS3_SpawnBio" = "SpawnBio",
    "SS3_Bio" = "Bio_all"
  ) |>
  dplyr::mutate(
    FIMS_SpawnBio = FIMS::get_report(fit)[["spawning_biomass"]][[1]][-1] , 
    FIMS_Bio = FIMS::get_report(fit)[["biomass"]][[1]][-1]
  ) |> ##CHECK: Is FIMS ssb reporting n_years+1 or initial year-1?
  tidyr::pivot_longer(cols = -Yr) |>
  tidyr::separate_wider_delim(
    cols = "name",
    delim = "_", names = c("Model", "Type")
  )

# plot comparison of spawning biomass and total biomass time series
ggplot2::ggplot(biomass, ggplot2::aes(x = Yr, y = value)) +
  ggplot2::geom_line(ggplot2::aes(color = Model)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("") +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw()

# gather info on recruitment
recruits <- rep$recruit |>
  dplyr::filter(Yr %in% years) |>
  dplyr::select(Yr, exp_recr, raw_dev) |>
  dplyr::rename("SS3_recruit" = "exp_recr",
                "SS3_recdev" = "raw_dev",
                "Year" = "Yr") |> 
  dplyr::mutate(FIMS_recruit = c(FIMS::get_report(fit)[["expected_recruitment"]][[1]][-1]),
                FIMS_recdev = c(
                  NA,
                  FIMS::get_estimates(fit) |>
                    dplyr::filter(
                      module_name == "Recruitment",
                      label == "log_devs"
                    ) |>
                    dplyr::pull(estimated)
                ),
                SS3_recruit = SS3_recruit * 1000) |>
  tidyr::pivot_longer(cols = -Year) |>
  tidyr::separate_wider_delim(cols = "name", delim = "_", names = c("Model", "Type"))             

# plot recruit time series
ggplot2::ggplot(recruits, ggplot2::aes(x = Year, y = value, color = Model)) +
  ggplot2::geom_line() + 
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw()

```

```{r}
#| label: cleanup
#| output: false
#| warning: false
FIMS::clear()
```
