---
title: NWFSC Case Study Petrale Sole
stock: "NWFSC petrale"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Petrale sole"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model a simpler SS3 model was developed using code in https://github.com/NOAA-FIMS/case-studies/blob/petrale-with-new-r4ss-converter/content/R/NWFSC-petrale_simplify_assessment.R

These changes included 
* Remove data
  * Remove male lengths and ages
  * Remove discard fractions and discard comps
* Simplify selectivity
  * Remove length-based retention functions
  * Convert to age-based logistic (from asymptotic length-based double-normal with time blocks)
* Use time-invariant population-level female mean weight at age from the SS3 weight-at-age file for all fleets
* Use time-invariant maturity x fecundity at age from the SS3 weight-at-age file

## Setting up the data

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# convert SS3 data into FIMS format using function in r4ss and convert to FIMSFrame object
data_4_model <- r4ss::ss3_data_to_fims(
    ss3_dir = "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage/",
    ss_new = TRUE,
    fleets = c(-4, 1, 2, 4)
) |>
    FIMS::FIMSFrame()


# read SS3 output files from modified petrale sole assessment on GitHub
rep <- r4ss::SS_output(
    "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

# # commented out pending fix to https://github.com/nmfs-ost/stockplotr/pull/317
# # alternative output processing using stockplotr
# output_petrale <- stockplotr::convert_output(
#     "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage/Report.sso",
#     model = "ss3"
# )

```

## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

# lots of code below copied PIFSC-opakapaka.qmd
years <- (data_4_model |> FIMS::get_start_year()):(data_4_model |>
    FIMS::get_end_year())
ages <- data_4_model |> FIMS::get_ages()

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
  data = data_4_model,
  growth_module_type = "VonBertalanffy"
)

# Note: the last value of the initial numbers at age is the first
# recruitment deviation
recdev_years <- default_parameters |>
    dplyr::filter(module_name == "Recruitment" & label == "log_devs") |>
    dplyr::pull(timing)
recdevs <- rep$recruitpars |> # $recruitpars is a subset of $parameters with Yr as an additional column
    dplyr::filter(Yr %in% recdev_years) |>
    dplyr::pull(Value)

# initial numbers-at-age from SS3 output
init_naa_ss3 <- rep$natage |>
        dplyr::filter(Time == min(years), Sex == 1) |> # numbers at age in FIMS start year, not SS3 model
        dplyr::select(paste(ages)) |>
        as.numeric() 

# alternative initial numbers-at-age based on exponential decay using female M
init_naa <- exp(
    rep$parameters |> dplyr::filter(grepl("R0", Label)) |> dplyr::pull(Value)
) *
    1000 *
    exp(
        -(ages - 1) *
            rep$parameters |>
                dplyr::filter(grepl("NatM.*Fem", Label)) |>
                dplyr::pull(Value)
    )
# init_naa calculation for plus group is based on sum of infinite series,
# which is equivalent to dividing the initial number at the last age by M
init_naa[length(ages)] <- init_naa[length(ages)] /
    rep$parameters |>
        dplyr::filter(grepl("NatM.*Fem", Label)) |>
        dplyr::pull(Value) # sum of infinite series

# ratio is close to 1, suggesting that approximation is reasonable
# range(init_naa / init_naa_ss3)
# # [1] 1.037574 1.228785

# function to get estimated parameter value from SS3 model
get_ss3_estimate <- function(string) {
    rep$parameters |>
        dplyr::filter(grepl(
            pattern = string,
            Label,
            fixed = TRUE
        )) |>
        dplyr::pull(Value)
}

parameters <- default_parameters |>
  # update the growth parameters with initial values from SS3
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Growth",
      label = c(
        "length_at_ref_age_1",
        "length_at_ref_age_2",
        "growth_coefficient_K",
        "reference_age_for_length_1",
        "reference_age_for_length_2",
        "length_weight_a", # added length-weight relationship parameters
        "length_weight_b"
      ),
      value = c(8.87713, 47.4413, .197191, 1, 17, 2.035e-06, 3.478),
      estimation_type = "constant"
    ),
    by = c("module_name", "label")
  ) |>
  # change first reference age for sd values from 0 to 1
  dplyr::mutate(
    age = dplyr::if_else(
      module_name == "Growth" & label == "length_at_age_sd_at_ref_ages" & age == 0,
      1,
      age
    )
  ) |> 
  dplyr::rows_update( #Note: this could probably be combined with Growth block above and put NAs for age for the other parameters
    tibble::tibble(
      module_name = "Growth",
      label = "length_at_age_sd_at_ref_ages",
      age = c(1, 17),
      value = c(1.2682, 4.8766)
    ),
    by=c("module_name", "label", "age")
  ) |>

    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Maturity",
    #         label = c("inflection_point", "slope"),
    #         value = c(7.0, 1.0), # maturity based on the derived age-based pattern from original length-based SS3 model, arbitrary guess of slope
    #     ),
    #     by = c("module_name", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "North",
    #         label = c("inflection_point", "slope"),
    #         value = c(10, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "South",
    #         label = c("inflection_point", "slope"),
    #         value = c(10, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "WCGBTS",
    #         label = c("inflection_point", "slope"),
    #         value = c(6, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>

    # set initial values for stock-recruitment parameters based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                #log(1000) + # adding log(1000) to account for R0 units in 1000s in SS3
                    get_ss3_estimate("SR_LN(R0)"),
                FIMS::logit(
                    0.2,
                    1.0,
                    get_ss3_estimate("SR_BH_steep")
                ),
                get_ss3_estimate("SR_sigmaR")
            )
        ),
        by = c("module_name", "label")
    ) |>
    # set initial values for recruitment deviations based on SS3 output,
    # starting with the second year of the SS3 model since the first recruitment deviation
    # corresponds to the initial numbers at age
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = "log_devs",
            timing = recdev_years,
            value = recdevs,
        ),
        by = c("module_name", "label", "timing")
    ) |>
    # set initial value for log_q based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            fleet = "WCGBTS",
            label = "log_q",
            value = get_ss3_estimate("LnQ_base_WCGBTS(4)")
        ),
        by = c("fleet", "label")
    ) |>
    # set initial numbers at age based on SS3 output calculated above
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = FIMS::get_ages(data_4_model),
            value = log(init_naa),
            estimation_type = "fixed_effects"
        ),
        by = c("module_name", "label", "age")
    ) |>
    # set initial M based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log(get_ss3_estimate("NatM_uniform_Fem_GP_1")),
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )

# confirm that all columns except "value" and "estimation_type" are identical between parameters and default_parameters
if (!identical(
    parameters |>
        dplyr::select(-value, -estimation_type),
    default_parameters |>
        dplyr::select(-value, -estimation_type)
)) {
    cli::cli_alert_danger("Columns other than 'value' and 'estimation_type' are not identical between parameters and default_parameters")
}
# figure out if any rows have different estimate_type
cli::cli_alert_info("Rows with different estimation_type between default_parameters parameters:")
diff_estimation_type <- parameters |>
    dplyr::filter(estimation_type != default_parameters$estimation_type) |>
    dplyr::select(module_name, fleet, label, estimation_type)

if (nrow(diff_estimation_type) == 0) {
    cli::cli_alert_success("No rows with different estimation_type.")
} else {
    print(diff_estimation_type, n = Inf, width = Inf)
}

# parameters which have been updated from defaults
cli::cli_alert_info("Rows with updated initial values from default_parameters to parameters:")
updated_parameter_counts <- parameters |>
    dplyr::filter(value != default_parameters$value) |>
    dplyr::select(module_name, fleet, label, value) |>
    dplyr::count(module_name, label)

if (nrow(updated_parameter_counts) == 0) {
    cli::cli_alert_success("No rows with updated initial values.")
} else {
    print(updated_parameter_counts, n = Inf, width = Inf)
}
# Run the model without optimization to help ensure a viable model
test_fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = FALSE)

# Run the  model with optimization
fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)
#fit_est <- fit # backup in case we want to set fit <- test_fit
```

## Plotting results

TODO: Add plotting code for the FIMS model results

```{r}
#| eval: true
#| label: growth-curve-plots

## Plot Growth Curve

# Pull the report and estimates from the fitted model.
report_fit <- FIMS::get_report(fit)
est_fit <- FIMS::get_estimates(fit)

# Pull model ages.
ages_fit <- FIMS::get_ages(data_4_model)

# Pull mean length-at-age from the report.
growth_laa_values <- unlist(report_fit[["growth_mean_LAA"]])

# Convert the report vector into an age-by-year matrix.
n_ages_fit <- length(ages_fit)
n_growth_years_fit <- length(growth_laa_values) / n_ages_fit

growth_laa_matrix <- matrix(
  growth_laa_values,
  nrow = n_ages_fit,
  ncol = n_growth_years_fit
)

# Put the Growth curve into a plotting table.
growth_curve_fit <- tibble::tibble(
  age = rep(ages_fit, times = n_growth_years_fit),
  year_i = rep(seq_len(n_growth_years_fit), each = n_ages_fit),
  mean_length = as.vector(growth_laa_matrix)
) |>
  dplyr::group_by(age) |>
  dplyr::summarise(
    mean_length = mean(mean_length, na.rm = TRUE),
    min_length = min(mean_length, na.rm = TRUE),
    max_length = max(mean_length, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(platform = "FIMS")

# SS3 mean length at age for comparison
growth_curve_ss3 <- rep$endgrowth |> 
  dplyr::select(Age_Beg, Len_Beg, SD_Beg) |> # beginning of year values, could also use "Mid" instead of "Beg"
  dplyr::rename(age = Age_Beg, mean_length = Len_Beg, sd_length = SD_Beg) |> 
  dplyr::mutate(min_length = NA, max_length = NA, platform = "SS3")

# combine FIMS and SS3 growth curves for plotting
growth_curve_both <- dplyr::bind_rows(growth_curve_fit, growth_curve_ss3)

# Plot mean length-at-age for both models.
ggplot2::ggplot(
  growth_curve_both,
  ggplot2::aes(x = age, y = mean_length, color = platform, fill = platform, linetype = platform)
) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::geom_point(size = 2) +
  ggplot2::labs(
    title = "fit Growth curve",
    subtitle = "Mean length-at-age from FIMS vs SS3",
    x = "Age",
    y = "Mean length-at-age"
  ) +
  ggplot2::theme_bw()
```

```{r}
#| label: cleanup
#| output: false
#| warning: false
FIMS::clear()
```