---
title: NWFSC Case Study Petrale Sole
stock: "NWFSC petrale"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Petrale sole"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model a simpler SS3 model was developed using code in https://github.com/NOAA-FIMS/case-studies/blob/petrale-with-new-r4ss-converter/content/R/NWFSC-petrale_simplify_assessment.R

These changes included 
* Remove data
  * Remove male lengths and ages
  * Remove discard fractions and discard comps
  * Remove ageing error
  * Remove fecundity relationship
* Simplify selectivity
  * Remove length-based retention functions
  * Convert to age-based logistic (from asymptotic length-based double-normal with time blocks)

## Setting up the data

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# convert SS3 data into FIMS format using function in r4ss and convert to FIMSFrame object
data_4_model <- r4ss::ss3_data_to_fims(
  ss3_dir = "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.005_FIMS_case-study_1sex_with_growth_no_fecundity/",
  ss_new = TRUE,
  fleets = c(-4, 1, 2, 4)
) |>
  dplyr::filter(!type %in% c("age_to_length_conversion", "weight_at_age")) |>
  FIMS::FIMSFrame()


# read SS3 output files from modified petrale sole assessment on GitHub
rep <- r4ss::SS_output(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.005_FIMS_case-study_1sex_with_growth_no_fecundity/",
  printstats = FALSE,
  verbose = FALSE,
  covar = FALSE
)

# alternative output processing using stockplotr
output_petrale <- stockplotr::convert_output(
  "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.005_FIMS_case-study_1sex_with_growth_no_fecundity/Report.sso",
  model = "ss3"
)
output_petrale_modified <- output_petrale |>
  dplyr::mutate(
    module_name = ifelse(
      module_name == "TIME_SERIES",
      "Population",
      module_name
    )
  )
```

## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

# lots of code below copied PIFSC-opakapaka.qmd
years <- (data_4_model |> FIMS::get_start_year()):(data_4_model |>
  FIMS::get_end_year())
ages <- data_4_model |> FIMS::get_ages()

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
  data = data_4_model # ,
  # growth_module_type = "VonBertalanffy"
)

# Note: the last value of the initial numbers at age is the first
# recruitment deviation
recdev_years <- default_parameters |>
  dplyr::filter(module_name == "Recruitment" & label == "log_devs") |>
  dplyr::pull(timing)
recdevs <- rep$recruitpars |> # $recruitpars is a subset of $parameters with Yr as an additional column
  dplyr::filter(Yr %in% recdev_years) |>
  dplyr::pull(Value)

# initial numbers-at-age from SS3 output
init_naa_ss3 <- rep$natage |>
  dplyr::filter(Time == min(years), Sex == 1) |> # numbers at age in FIMS start year, not SS3 model
  dplyr::select(paste(ages)) |>
  as.numeric()

# alternative initial numbers-at-age based on exponential decay using female M
init_naa <- exp(
  rep$parameters |> dplyr::filter(grepl("R0", Label)) |> dplyr::pull(Value)
) *
  exp(
    -(ages - 1) *
      rep$parameters |>
        dplyr::filter(grepl("NatM.*Fem", Label)) |>
        dplyr::pull(Value)
  )
# init_naa calculation for plus group is based on sum of infinite series,
# which is equivalent to dividing the initial number at the last age by M
init_naa[length(ages)] <- init_naa[length(ages)] /
  rep$parameters |>
    dplyr::filter(grepl("NatM.*Fem", Label)) |>
    dplyr::pull(Value) # sum of infinite series

# ratio is close to 1, suggesting that approximation is reasonable
# range(init_naa / init_naa_ss3)

# initial fishing mortality for North fleet based on SS3 output
Fmort_SS3 <- rep$catch |>
  dplyr::select(Yr, Fleet_Name, F) |>
  dplyr::filter(Yr %in% years)


# function to get estimated parameter value from SS3 model
get_ss3_estimate <- function(string) {
  rep$parameters |>
    dplyr::filter(grepl(
      pattern = string,
      Label,
      fixed = TRUE
    )) |>
    dplyr::pull(Value)
}

# function to convert SS3 logistic selectivity parameters to FIMS logistic selectivity parameters
# SS3 logistic selectivity is defined as:
# S(a) = 1 / (1 + exp(-ln(19) * (a - p1) / p2))}
# FIMS logistic selectivity is defined as:
# S(a) = 1 / (1 + exp(-1 * slope * (a - inflection)))
convert_ss3_logistic_to_fims <- function(ss3_parameters) {
  fims_inflection <- ss3_parameters[1]
  # -ln(19) / p2 = -1 * slope
  fims_slope <- log(19) / ss3_parameters[2]
  return(c(fims_inflection, fims_slope))
}

parameters <- default_parameters |>
  # update the growth parameters with initial values from SS3
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Growth",
      label = c(
        "mean_length_young",
        "mean_length_old",
        "von_bertalanffy_coefficient_K",
        "reference_age_for_length_1",
        "reference_age_for_length_2",
        "length_weight_a", # added length-weight relationship parameters
        "length_weight_b"
      ),
      # values below are SS3 estimates, code won't work for 2-sex model
      # or model with time-varying growth parameters
      value = c(
        rep$Growth_Parameters$L_a_A1,
        rep$Growth_Parameters$L_a_A2,
        rep$Growth_Parameters$K,
        rep$Growth_Parameters$A1,
        rep$Growth_Parameters$A2,
        rep$Growth_Parameters$WtLen1,
        rep$Growth_Parameters$WtLen2
      ),
      estimation_type = "constant"
    ),
    by = c("module_name", "label")
  ) |>
  # change first reference age for sd values from 0 to 1
  dplyr::mutate(
    age = dplyr::if_else(
      module_name == "Growth" &
        label == "length_at_age_sd_at_ref_ages" &
        age == 0,
      1,
      age
    )
  ) |>
  dplyr::rows_update(
    # Note: this could probably be combined with Growth block above and put NAs for age for the other parameters
    tibble::tibble(
      module_name = "Growth",
      label = "length_at_age_sd_at_ref_ages",
      age = c(1, 17),
      value = c(rep$Growth_Parameters$CVmin, rep$Growth_Parameters$CVmax)
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(7.0, 1.0), # maturity based on the derived age-based pattern from original length-based SS3 model, arbitrary guess of slope
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "North",
      label = c("inflection_point", "slope"),
      value = c(
        get_ss3_estimate("Age_inflection_North(1)"),
        log(19) / get_ss3_estimate("Age_95%width_North(1)")
      ),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "South",
      label = c("inflection_point", "slope"),
      # south fleet set to mirror north fleet in SS3 model, so use same values here
      value = c(
        get_ss3_estimate("Age_inflection_North(1)"),
        log(19) / get_ss3_estimate("Age_95%width_North(1)")
      ),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "WCGBTS",
      label = c("inflection_point", "slope"),
      value = c(
        get_ss3_estimate("Age_inflection_WCGBTS(4)"),
        log(19) / get_ss3_estimate("Age_95%width_WCGBTS(4)")
      ),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  # set initial values for stock-recruitment parameters based on SS3 output
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_rzero", "logit_steep", "log_sd"),
      value = c(
        # log(1000) + # adding log(1000) to account for R0 units in 1000s in SS3
        get_ss3_estimate("SR_LN(R0)"),
        FIMS::logit(
          0.2,
          1.0,
          get_ss3_estimate("SR_BH_steep")
        ),
        get_ss3_estimate("SR_sigmaR")
      )
    ),
    by = c("module_name", "label")
  ) |>
  # set initial values for recruitment deviations based on SS3 output,
  # starting with the second year of the SS3 model since the first recruitment deviation
  # corresponds to the initial numbers at age

  # TODO: how to make constant?
  #   Error in `FIMS::initialize_fims()`:
  # x Missing required inputs for recruitment process random or fixed effects.
  # i There is a distribution specified for the Recruitment `module_name` in the configurations tibble, but no parameters are specified for the recruitment   process in the
  #   parameters tibble.
  # i Implement either one of the following options to resolve this error:
  # i 1. Add parameter, `log_devs` or `log_r`, for the recruitment process in the parameters tibble with an estimation_type of random_effects or fixed_effects.
  # i 2. Set the distribution for the Recruitment distribution and distribution_type to `NA` in the configurations tibble.
  # Run `rlang::last_trace()` to see where the error occurred.

  # dplyr::rows_update(
  #   tibble::tibble(
  #     module_name = "Recruitment",
  #     label = "log_devs",
  #     timing = recdev_years,
  #     value = recdevs,
  #     estimation_type = "constant",
  #     distribution = NA,
  #     distribution_type = NA
  #   ),
  #   by = c("module_name", "label", "timing")
  # ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_devs",
      timing = recdev_years,
      value = recdevs,
      # estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "timing")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      label = "log_Fmort",
      fleet = Fmort_SS3$Fleet_Name,
      timing = Fmort_SS3$Yr,
      value = log(Fmort_SS3$F + 0.0001) # add small value to avoid log(0),
    ),
    by = c("module_name", "label", "fleet", "timing")
  ) |>
  # set initial value for log_q based on SS3 output
  dplyr::rows_update(
    tibble::tibble(
      fleet = "WCGBTS",
      label = "log_q",
      value = get_ss3_estimate("LnQ_base_WCGBTS(4)")
    ),
    by = c("fleet", "label")
  ) |>
  # set initial numbers at age based on SS3 output calculated above
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = FIMS::get_ages(data_4_model),
      value = log(init_naa),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  # set initial M based on SS3 output
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_M",
      value = log(get_ss3_estimate("NatM_uniform_Fem_GP_1")),
      estimation_type = "constant"
    ),
    by = c("module_name", "label")
  )

# confirm that all columns except "value" and "estimation_type" are identical between parameters and default_parameters
if (
  !identical(
    parameters |>
      dplyr::select(-value, -estimation_type),
    default_parameters |>
      dplyr::select(-value, -estimation_type)
  )
) {
  cli::cli_alert_danger(
    "Columns other than 'value' and 'estimation_type' are not identical between parameters and default_parameters"
  )
}
# figure out if any rows have different estimate_type
cli::cli_alert_info(
  "Rows with different estimation_type between default_parameters parameters:"
)
diff_estimation_type <- parameters |>
  dplyr::filter(estimation_type != default_parameters$estimation_type) |>
  dplyr::select(module_name, fleet, label, estimation_type)

if (nrow(diff_estimation_type) == 0) {
  cli::cli_alert_success("No rows with different estimation_type.")
} else {
  print(diff_estimation_type, n = Inf, width = Inf)
}

# parameters which have been updated from defaults
cli::cli_alert_info(
  "Rows with updated initial values from default_parameters to parameters:"
)
updated_parameter_counts <- parameters |>
  dplyr::filter(value != default_parameters$value) |>
  dplyr::select(module_name, fleet, label, value) |>
  dplyr::count(module_name, label)

if (nrow(updated_parameter_counts) == 0) {
  cli::cli_alert_success("No rows with updated initial values.")
} else {
  print(updated_parameter_counts, n = Inf, width = Inf)
}
```

```{r}
#| label: run-model

input <- parameters |>
  FIMS::initialize_fims(data = data_4_model)
input$model$ReportGrowthDerivedALKTensor(TRUE)

# Run the model without optimization to help ensure a viable model
test_fit <- input |>
  FIMS::fit_fims(optimize = FALSE)

fit <- test_fit

# Run the  model with optimization
fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)

fit_est <- fit
```
## Plotting inputs

```{r plotting-inputs}
data_4_model@data |>
  dplyr::filter(type == "age_comp" & observed != -999) |>
  ggplot2::ggplot(
    ggplot2::aes(x = age, y = observed, col = fleet, fill = fleet)
  ) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::xlab("Age") +
  ggplot2::ylab("Number") +
  ggplot2::facet_wrap(~timing) +
  ggtitle("Age compositions by fleet and time step")


data_4_model@data |>
  dplyr::filter(type == "length_comp" & observed != -999) |>
  ggplot2::ggplot(
    ggplot2::aes(x = length, y = observed, col = fleet, fill = fleet)
  ) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::xlab("Length") +
  ggplot2::ylab("Number") +
  ggplot2::facet_wrap(~timing) +
  ggtitle("Length compositions by fleet and time step")
```
## Plotting results

TODO: Add plotting code for the FIMS model results

```{r}
output <- FIMS::get_estimates(fit) |>
  # mutate commands below copied from other case studies, but not sure what all it is doing or why
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + FIMS::get_start_year(data_4_model) - 1,
    estimate = estimated,
    era = "time", # not sure if this is needed
    fleet = dplyr::recode_values(
      module_id,
      1 ~ data_4_model@fleets[1],
      2 ~ data_4_model@fleets[2],
      3 ~ data_4_model@fleets[3],
      default = NA_character_
    )
  )

# # attempt at stockplotr plot didn't turn out well, so commented out for now
# stockplotr::plot_spawning_biomass(
#   list(
#     "FIMS" = output,
#     "SS3" = output_petrale_modified
#   )
# )

# SS3 timeseries includes one extra initial biomass year before the initial year
ss3_biomass <- rep$timeseries |>
  dplyr::filter(Era != "VIRG") |>
  dplyr::select(Yr, SpawnBio, Bio_all) |>
  dplyr::rename(
    year = Yr,
    Bio = Bio_all
  ) |>
  tidyr::pivot_longer(
    cols = c(SpawnBio, Bio),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "SS3")

# FIMS biomass and spawning biomass include one extra initial-year value.
# That first value corresponds to start_year - 1, so plot FIMS on
# (start_year - 1):end_year rather than start_year:end_year.
fims_biomass <- tibble::tibble(
  year = (FIMS::get_start_year(data_4_model) - 1):FIMS::get_end_year(
    data_4_model
  ),
  SpawnBio = FIMS::get_report(fit)[["spawning_biomass"]][[1]],
  Bio = FIMS::get_report(fit)[["biomass"]][[1]]
) |>
  tidyr::pivot_longer(
    cols = c(SpawnBio, Bio),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "FIMS")

biomass <- dplyr::bind_rows(ss3_biomass, fims_biomass)

ggplot2::ggplot(biomass, ggplot2::aes(x = year, y = value)) +
  ggplot2::geom_line(ggplot2::aes(color = Model)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("") +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::ylim(0, NA)

# SS3 recruitment output is already indexed by report year in `Yr`, so use
# those years directly for the recruitment comparison plot.
ss3_recruits <- rep$recruit |>
  dplyr::select(Yr, pred_recr, raw_dev) |>
  dplyr::rename(
    Year = Yr,
    recruit = pred_recr,
    recdev = raw_dev
  ) |>
  dplyr::mutate(
    recruit = recruit
  ) |>
  tidyr::pivot_longer(
    cols = c(recruit, recdev),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "SS3")

# FIMS expected recruitment is reported on modeled years start_year:end_year.
# Recruitment deviations are not defined for the first modeled year, so pad
# that first year with NA before plotting.
fims_recruit_years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(
  data_4_model
)

fims_recruits <- tibble::tibble(
  Year = fims_recruit_years,
  recruit = FIMS::get_report(fit)[["expected_recruitment"]][[1]][seq_along(
    fims_recruit_years
  )],
  recdev = c(
    NA,
    FIMS::get_estimates(fit) |>
      dplyr::filter(label == "log_devs", module_name == "Recruitment") |>
      # dplyr::mutate(estimated = exp(estimated)) |>
      dplyr::pull(estimated)
  )
) |>
  tidyr::pivot_longer(
    cols = c(recruit, recdev),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "FIMS") |>
  dplyr::filter(!is.na(value))

# Drop undefined recruitment deviations before plotting time series.
recruits <- dplyr::bind_rows(ss3_recruits, fims_recruits) |>
  dplyr::filter(!is.na(value))

ggplot2::ggplot(recruits, ggplot2::aes(x = Year, y = value, color = Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::ggtitle("Recruitment and Recruitment Deviations") +
  ggplot2::theme_bw()


# plot catch
catch_years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(
  data_4_model
)

observed_catch <- data_4_model@data |>
  dplyr::filter(type == "catch") |>
  dplyr::select(fleet, year = timing, observed)

expected_catch <- output |>
  dplyr::filter(label == "catch_expected") |>
  dplyr::select(fleet, year, expected)

catch_results <- observed_catch |>
  dplyr::left_join(expected_catch, by = c("fleet", "year")) |>
  dplyr::filter(observed > -999)

# print(catch_results)

ggplot2::ggplot(catch_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point(ggplot2::aes(color = fleet)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected, color = fleet)) +
  ggplot2::theme_bw()
```


```{r}
#| eval: true
#| label: growth-curve-plots

## Plot Growth Curve

# Pull the report and estimates from the fitted model.
report_fit <- FIMS::get_report(fit)
est_fit <- FIMS::get_estimates(fit)

# Pull model ages.
ages_fit <- FIMS::get_ages(data_4_model)

# Pull mean length-at-age from the report.
growth_laa_values <- unlist(report_fit[["growth_mean_LAA"]])

# Convert the report vector into an age-by-year matrix.
n_ages_fit <- length(ages_fit)
n_growth_years_fit <- length(growth_laa_values) / n_ages_fit

growth_laa_matrix <- matrix(
  growth_laa_values,
  nrow = n_ages_fit,
  ncol = n_growth_years_fit
)

# Put the Growth curve into a plotting table.
growth_curve_fit <- tibble::tibble(
  age = rep(ages_fit, times = n_growth_years_fit),
  year_i = rep(seq_len(n_growth_years_fit), each = n_ages_fit),
  mean_length = as.vector(growth_laa_matrix)
) |>
  dplyr::group_by(age) |>
  dplyr::summarise(
    mean_length = mean(mean_length, na.rm = TRUE),
    min_length = min(mean_length, na.rm = TRUE),
    max_length = max(mean_length, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(platform = "FIMS")

# SS3 mean length at age for comparison
growth_curve_ss3 <- rep$endgrowth |>
  dplyr::select(Age_Beg, Len_Beg, SD_Beg) |> # beginning of year values, could also use "Mid" instead of "Beg"
  dplyr::rename(age = Age_Beg, mean_length = Len_Beg, sd_length = SD_Beg) |>
  dplyr::mutate(min_length = NA, max_length = NA, platform = "SS3")

# combine FIMS and SS3 growth curves for plotting
growth_curve_both <- dplyr::bind_rows(growth_curve_fit, growth_curve_ss3)

# Plot mean length-at-age for both models.
ggplot2::ggplot(
  growth_curve_both,
  ggplot2::aes(
    x = age,
    y = mean_length,
    color = platform,
    fill = platform,
    linetype = platform
  )
) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::geom_point(size = 2) +
  ggplot2::labs(
    title = "fit Growth curve",
    subtitle = "Mean length-at-age from FIMS vs SS3",
    x = "Age",
    y = "Mean length-at-age"
  ) +
  ggplot2::theme_bw()

ggsave(
  filename = "growth_curve_plot.png",
  width = 6,
  height = 4.5,
  units = "in",
  dpi = 300
)

realized_age_to_length_conversion_dataframe <- get_realized_age_to_length_conversion_dataframe(
  fit = fit,
  data = data_4_model
)

plot_age_to_length_conversion(realized_age_to_length_conversion_dataframe)
ggsave(
  filename = "age_to_length_conversion_plot.png",
  width = 6,
  height = 4.5,
  units = "in",
  dpi = 300
)
```

```{r}
#| label: cleanup
#| output: false
#| warning: false
FIMS::clear()
```