---
title: NWFSC Case Study Petrale Sole
stock: "NWFSC petrale"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Petrale sole"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: NWFSC
* Analyst: Ian G. Taylor

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model a simpler SS3 model was developed using code in https://github.com/NOAA-FIMS/case-studies/blob/petrale-with-new-r4ss-converter/content/R/NWFSC-petrale_simplify_assessment.R

These changes included 
* Remove data
  * Remove male lengths and ages
  * Remove discard fractions and discard comps
* Simplify selectivity
  * Remove length-based retention functions
  * Convert to age-based logistic (from asymptotic length-based double-normal with time blocks)
* Use time-invariant population-level female mean weight at age from the SS3 weight-at-age file for all fleets
* Use time-invariant maturity x fecundity at age from the SS3 weight-at-age file

## Setting up the data

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false

# convert SS3 data into FIMS format using function in r4ss and convert to FIMSFrame object
data_4_model <- r4ss::ss3_data_to_fims(
    ss3_dir = "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage/",
    ss_new = TRUE,
    fleets = c(-4, 1, 2, 4)
) |>
    FIMS::FIMSFrame()


# read SS3 output files from modified petrale sole assessment on GitHub
rep <- r4ss::SS_output(
    "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage",
    printstats = FALSE,
    verbose = FALSE,
    covar = FALSE
)

# # commented out pending fix to https://github.com/nmfs-ost/stockplotr/pull/317
# # alternative output processing using stockplotr
# output_petrale <- stockplotr::convert_output(
#     "https://raw.githubusercontent.com/pfmc-assessments/petrale/main/models/2023.a050.003_FIMS_case-study_wtatage/Report.sso",
#     model = "ss3"
# )

```

## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

# lots of code below copied PIFSC-opakapaka.qmd
years <- (data_4_model |> FIMS::get_start_year()):(data_4_model |>
    FIMS::get_end_year())
ages <- data_4_model |> FIMS::get_ages()

# Create default parameters
default_parameters <- FIMS::create_default_configurations(
    data = data_4_model
) |>
    FIMS::create_default_parameters(
        data = data_4_model
    ) |>
    tidyr::unnest(cols = data)
# Note: the last value of the initial numbers at age is the first
# recruitment deviation
recdev_years <- default_parameters |>
    dplyr::filter(module_name == "Recruitment" & label == "log_devs") |>
    dplyr::pull(time)
recdevs <- rep$recruitpars |> # $recruitpars is a subset of $parameters with Yr as an additional column
    dplyr::filter(Yr %in% recdev_years) |>
    dplyr::pull(Value)

# initial numbers-at-age from SS3 output
init_naa_ss3 <- 1000 *
    rep$natage |>
        dplyr::filter(Time == min(years), Sex == 1) |> # numbers at age in FIMS start year, not SS3 model
        dplyr::select(paste(ages)) |>
        as.numeric() # TODO: figure out if *1000 is needed

# alternative initial numbers-at-age based on exponential decay using female M
init_naa <- exp(
    rep$parameters |> dplyr::filter(grepl("R0", Label)) |> dplyr::pull(Value)
) *
    1000 *
    exp(
        -(ages - 1) *
            rep$parameters |>
                dplyr::filter(grepl("NatM.*Fem", Label)) |>
                dplyr::pull(Value)
    )
# init_naa calculation for plus group is based on sum of infinite series,
# which is equivalent to dividing the initial number at the last age by M
init_naa[length(ages)] <- init_naa[length(ages)] /
    rep$parameters |>
        dplyr::filter(grepl("NatM.*Fem", Label)) |>
        dplyr::pull(Value) # sum of infinite series

# ratio is close to 1, suggesting that approximation is reasonable
# range(init_naa / init_naa_ss3)
# # [1] 1.037574 1.228785

# function to get estimated parameter value from SS3 model
get_ss3_estimate <- function(string) {
    rep$parameters |>
        dplyr::filter(grepl(
            pattern = string,
            Label,
            fixed = TRUE
        )) |>
        dplyr::pull(Value)
}

parameters <- default_parameters |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Maturity",
    #         label = c("inflection_point", "slope"),
    #         value = c(7.0, 1.0), # maturity based on the derived age-based pattern from original length-based SS3 model, arbitrary guess of slope
    #     ),
    #     by = c("module_name", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "North",
    #         label = c("inflection_point", "slope"),
    #         value = c(10, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "South",
    #         label = c("inflection_point", "slope"),
    #         value = c(10, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>
    # dplyr::rows_update(
    #     tibble::tibble(
    #         module_name = "Selectivity",
    #         fleet = "WCGBTS",
    #         label = c("inflection_point", "slope"),
    #         value = c(6, 2),
    #         estimation_type = "constant"
    #     ),
    #     by = c("module_name", "fleet", "label")
    # ) |>

    # set initial values for stock-recruitment parameters based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = c("log_rzero", "logit_steep", "log_sd"),
            value = c(
                log(1000) + # adding log(1000) to account for R0 units in 1000s in SS3
                    get_ss3_estimate("SR_LN(R0)"),
                FIMS::logit(
                    0.2,
                    1.0,
                    get_ss3_estimate("SR_BH_steep")
                ),
                get_ss3_estimate("SR_sigmaR")
            )
        ),
        by = c("module_name", "label")
    ) |>
    # set initial values for recruitment deviations based on SS3 output,
    # starting with the second year of the SS3 model since the first recruitment deviation
    # corresponds to the initial numbers at age
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Recruitment",
            label = "log_devs",
            time = recdev_years,
            value = recdevs,
        ),
        by = c("module_name", "label", "time")
    ) |>
    # set initial value for log_q based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            fleet = "WCGBTS",
            label = "log_q",
            value = get_ss3_estimate("LnQ_base_WCGBTS(4)")
        ),
        by = c("fleet", "label")
    ) |>
    # set initial numbers at age based on SS3 output calculated above
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_init_naa",
            age = FIMS::get_ages(data_4_model),
            value = log(init_naa),
            estimation_type = "fixed_effects"
        ),
        by = c("module_name", "label", "age")
    ) |>
    # set initial M based on SS3 output
    dplyr::rows_update(
        tibble::tibble(
            module_name = "Population",
            label = "log_M",
            value = log(get_ss3_estimate("NatM_uniform_Fem_GP_1")),
            estimation_type = "constant"
        ),
        by = c("module_name", "label")
    )

# confirm that all columns except "value" and "estimation_type" are identical between parameters and default_parameters
if (!identical(
    parameters |>
        dplyr::select(-value, -estimation_type),
    default_parameters |>
        dplyr::select(-value, -estimation_type)
)) {
    cli::cli_alert_danger("Columns other than 'value' and 'estimation_type' are not identical between parameters and default_parameters")
}
# figure out if any rows have different estimate_type
cli::cli_alert_info("Rows with different estimation_type between default_parameters parameters:")
diff_estimation_type <- parameters |>
    dplyr::filter(estimation_type != default_parameters$estimation_type) |>
    dplyr::select(module_name, fleet, label, estimation_type)

if (nrow(diff_estimation_type) == 0) {
    cli::cli_alert_success("No rows with different estimation_type.")
} else {
    print(diff_estimation_type, n = Inf, width = Inf)
}

# parameters which have been updated from defaults
cli::cli_alert_info("Rows with updated initial values from default_parameters to parameters:")
updated_parameter_counts <- parameters |>
    dplyr::filter(value != default_parameters$value) |>
    dplyr::select(module_name, fleet, label, value) |>
    dplyr::count(module_name, label)

if (nrow(updated_parameter_counts) == 0) {
    cli::cli_alert_success("No rows with updated initial values.")
} else {
    print(updated_parameter_counts, n = Inf, width = Inf)
}
# Run the model without optimization to help ensure a viable model
test_fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = FALSE)

# Run the  model with optimization
fit <- parameters |>
    FIMS::initialize_fims(data = data_4_model) |>
    FIMS::fit_fims(optimize = TRUE, get_sd = FALSE)
fit_est <- fit # backup in case we want to set fit <- test_fit
```

## Plotting results

TODO: Add plotting code for the FIMS model results

```{r}
#| label: cleanup
#| output: false
#| warning: false
FIMS::clear()
```
