---
title: U.S. West Coast Case Study of Pacific Hake
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Pacific Hake"

terminal_year <- 2024

hake_maturity_inflection_point <- 2.2
hake_maturity_slope <- 1.8
sample_size_multiplier <- 10
log_rzero_prior_mean <- 15
log_rzero_prior_sd <- 1.155
init_naa_prior_mean <- 18
init_naa_prior_sd <- 0.155
natural_mortality <- 0.235
steepness <- 0.8118
# Taken from CEATTLE estimates
natural_mortality_by_age <- dplyr::tibble(
  age = 1:5,
  cannibalism = c(
    # 0.619078417, 0.038591177, 0.012472784, 0.010002492, 0.018692867
    0.0, 0.0, 0.0, 0.0, 0.0
  )
)

# The model can be found within the following folder on the hake server 
# base_dir <- "/srv/hake/models/2025/01-version/01-base-models/01-base"
base_dir <- "https://raw.githubusercontent.com/pacific-hake/hake-assessment/refs/heads/model-files/models/2025"
maturity_data <- utils::read.csv("https://raw.githubusercontent.com/pacific-hake/hake-assessment/refs/heads/master/data-tables/maturity-ogives.csv")

# Hake results are in million metric tons, i.e., 1,000,000 t
# Read in file that was made smaller from full results
# make_hake_rds_smaller(
#   "c:/users/Kelli.Johnson/Documents/hake/01-base.rds",
#   file.path("c:/github/NOAA-FIMS/case-studies/content", data_directory, "pacific_hake.RDS")
# )
ss3_results <- readRDS(file.path(data_directory, "pacific_hake.RDS"))
ss3_sb <- ss3_results[["spawning_biomass"]]
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: U.S. West Coast
* Analyst: Kelli F. Johnson

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

The operational model was not altered in an attempt to more closely match a FIMS model.

## Data

The data for the FIMS model was created using `get_ss3_data()`, written by Drs. Ian G. Taylor and Megumi C. Oshima. The function was provided the 2025 Stock Synthesis model files using `r4ss::SS_read()`. All data for a FIMS model is contained in one data frame, using a long format, i.e., one row per observation instead of a matrix format. Here, we call this object `data_4_model`.

```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false

inputs <- r4ss::SS_read(base_dir)
inputs[["wtatage"]] <- r4ss::SS_readwtatage(file.path(base_dir, "wtatage.ss_new"))

stock_synthesis_data <- r4ss::ss3_data_to_fims(
  ss_new = TRUE,
  ss3_inputs = inputs,
  # TODO: Allow for a sensitivity with age-1 index but use Numbers
  fleets = seq(inputs[["dat"]][["Nfleets"]])[-3]
) |>
  dplyr::mutate(
    uncertainty = ifelse(
      test = type == "age" & name == "Acoustic_Survey",
      uncertainty * 30,
      uncertainty
    )
  )
start_year <- dplyr::filter(
  stock_synthesis_data,
  type == "age_comp",
  value != -999
) |>
  dplyr::pull(timing) |>
  min()

data_4_model <- dplyr::bind_rows(
    stock_synthesis_data |>
      dplyr::filter(timing <= terminal_year, timing >= start_year),
    add_additional_year(
      stock_synthesis_data,
      timing_to_average = terminal_year:(terminal_year - 4),
      type_to_filter = "weight_at_age",
      age
    )
  ) |>
  FIMS::FIMSFrame()
```

```{r}
#| label: prepare-year-conversion-matrix
#| echo: false
year_conversion_matrix <- dplyr::tibble(
  year = get_start_year(data_4_model):(get_end_year(data_4_model) + 1)
) |>
  dplyr::mutate(
    index = dplyr::row_number()
  )
```

```{r}
#| warning: false
#| label: show-fims-data
#| output: true
#| echo: true

# Default print method of a FIMSFrame object
data_4_model

# Summary of the data types available
FIMS::get_data(data_4_model) |>
  dplyr::filter(value != -999) |>
  dplyr::group_by(type, name) |>
  dplyr::count()
```

### Weight-at-age data

Currently, FIMS uses a single weight-at-age matrix as there is no seasonality or timing in FIMS and weight-at-age is assumed to be the same across the population and fleets. Thus, no simplifications were needed for the FIMS model regarding weight-at-age data because all of the weight-at-age matrices in the Stock Synthesis model are replicates of each other, e.g., weight at age for fleet two is the same as beginning and middle of the year population weight at age in this model. 

The last five years of weight-at-age data were averaged to provide weight-at-age information for the beginning of the year after the last year of landings were removed from the population.

### Maturity

For the 2025 base model, year-specific maturity at age was derived from a temperature-specific relationship. FIMS uses a single set of logistic maturity parameters for all years, and thus, the year-specific maturity ogives had to be translated to a single logistic curve.

First, annual estimates of maturity at age for day `r max(maturity_data[["doy"]])` were averaged across years. Subsequently, this averaged ogive was fit using a logistic curve external to the assessment model. This lead to a curve with an inflection of `r hake_maturity_inflection_point` and a slope of `r hake_maturity_slope` (Fig. @fig:maturity-plot). These values were specified as fixed inputs to the FIMS model.

Changes to the maturity model in FIMS can easily be accommodated but there are few populations that require changes, and thus, efforts have not been placed here. If desired, we could easily make the logistic parameters vary with time.

```{r}
#| label: maturity-plot
#| fig-cap: "Averaged maturity at age from the 2025 base model (empirical) compared to logistic maturity (logistic)."
#| fig-alt: "The two maturity ogives are quite similar with only small differences near the origin of lower empirical maturity for age-1 fish."
#| fig:align: center
#| echo: false

maturity_data |>
  dplyr::filter(doy == max(doy)) |>
  dplyr::group_by(age) |>
  dplyr::summarize(empirical = mean(p_mature)) |>
  dplyr::full_join(
    execute_logistic(
      unique(c(0, get_ages(data_4_model))),
      hake_maturity_slope,
      hake_maturity_inflection_point
    ) |>
      dplyr::rename(age = x, logistic = value),
    by = "age"
  ) |>
  tidyr::pivot_longer(cols = c("empirical", "logistic")) |>
  ggplot2::ggplot(ggplot2::aes(age, value, color = name)) +
  ggplot2::geom_line(linewidth = 1.5) +
  ggplot2::xlab("Age") +
  ggplot2::ylab("Probability of being mature") +
  ggplot2::scale_color_brewer(palette = "Dark2", name = "Method")
```

### Fisheries data

Any number of fleets, their landings, compositions, and catch-per-unit effort data can be accommodated in FIMS, and thus, regardless if the U.S. and Canadian fishery data are grouped into a single fleet, the fishery data can be accommodated in a FIMS model.

The Dirichlet-multinomial distribution is not yet available in FIMS. Compositional data were fit using a multinomial distribution. The input sample sizes for the survey age-composition data were multiplied by `r sample_size_multiplier` to up-weight the survey data.

A separate analysis was conducted to investigate the fit of the age-composition data to each of the distributions available in Stock Synthesis. The analysis concluded that, if the true sample size is known, that the hake ages can be fit using a multinomial distribution. More research is being conducted regarding how true these conclusions are when the input sample size is not properly specified.

### Survey data

Any number of surveys can be accommodated in a FIMS model. Additionally, the units of the data can be in metric tons or numbers.

<!-- TODO: run a sensitivity with the age-1 index -->
Regardless, the age-1 index is not fit to in this case study because it is not in the 2025 base model.

No extra uncertainty was estimated for the index of abundance because this feature is not available in FIMS. 

## Model configurations

### Fishing mortality

Annual fishing mortality was estimated using fixed effects.

### Recruitment

The natural log of initial recruitment was estimated as a fixed effect, initialized at the same value as the base model (`r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "INIT"]`). In the base model, this parameter is bounded between `r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "LO"]` and `r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "HI"]`. Parameter bounds do not exist in FIMS but priors do, and thus, a normal prior with a mean of `r log_rzero_prior_mean` and a standard deviation of `r log_rzero_prior_sd` was used.

Steepness was fixed at the median estimate from the 2025 base model of `r steepness`. The beta prior used in the base model is not yet available in FIMS.

FIMS models are currently powered by Template Model Builder (TMB), which allows for the estimation of random effects, similar to models built using the Woods Hole Assessment Model (WHAM). We harnessed this capability and estimated recruitment deviations as a random effect with a normal distribution and an estimated standard deviation.

```{r}
#| warning: false
#| label: prepare-initial-naa
init_naa <- exp(init_naa_prior_mean) *
  1000 *
  exp(-(get_ages(data_4_model) - 1) * natural_mortality)
init_naa[get_n_ages(data_4_model)] <- init_naa[get_n_ages(data_4_model)] /
  natural_mortality # sum of infinite series
```

The time series of recruitment deviations is one element shorter than the time series in a Stock Synthesis model because the initial numbers at age object includes initial numbers for age-zero fish. The vector of the natural log of initial numbers at age was estimated using fixed effects and initialized at values in line with initial values for unfished recruitment and natural mortality. Additionally, a normal prior with a mean of `r init_naa_prior_mean` and a standard deviation of `r init_naa_prior_sd` was placed on the natural log of initial numbers at age in an attempt to aid model convergence. The input data to the model was also truncated to exclude the early years before age-composition data is available. This aided in the estimation of these parameters.

The priors on the natural log of unfished recruitment and the natural log of initial numbers at age should be transformed using a Jacobian transformation because the parameters are in log space but that transformation is not currently available in FIMS. The Jacobian allows for adjustment of the scale of integration so the total measure is preserved.

### Selectivity

The logistic and double-logistic selectivity curves are currently the only selectivity forms that are available in FIMS. Both forms were investigated for Pacific Hake. We also investigated the use of a time-varying inflection point and/or slope for the fishing fleet. Additionally, we investigated allowing just a portion of the parameter vector to vary with time starting at several different time points, where the base model starts in 1991.
In the end, time-invariant logistic selectivity was used for both the fleet and the survey.

Work is in progress to implement an age-specific non-parametric selectivity curve in FIMS as there is a need for this option from both coasts. Consequently, we have put a high priority on it within the FIMS development team, and it will be available before the 2027 Pacific Hake assessment.

### Natural mortality

```{r}
#| label: setup-natural-mortality
#| echo: false
natural_mortality_matrix <- natural_mortality_by_age |>
  tidyr::complete(
    age = get_ages(data_4_model),
    fill = list(cannibalism = 0)
  ) |>
  tidyr::crossing(
    year = get_start_year(data_4_model):get_end_year(data_4_model)
  ) |>
  dplyr::arrange(year, age) |>
  dplyr::mutate(m = cannibalism + natural_mortality)
```

<!-- TODO: Get the median of the base model -->
<!-- TODO: Fix the parameter naming when parameters are mapped -->
<!-- TODO: Do a sensitivity where M is fixed at the values estimated from the RCEATTLE model -->
<!-- TODO: Do a sensitivity where M is estimated using a single value with the lognormal prior -->
Natural mortality in FIMS is age and year specific. To estimate a single parameter, the entire parameter vector could be mapped to a single value. That was not investigated here. Instead, natural mortality for all years was fixed at the sum of the median estimate of natural mortality from the 2025 base model (`r natural_mortality`) and the mean estimate of cannibalism across all years for a given age group. The use of age-specific natural mortality was to help stabilize the estimates of initial numbers at age, though the endeavour was not entirely successful.

```{r}
#| label: natural-mortality-table
#| tbl-cap: "Estimates of cannibalism from the CEATTLE model for ages 1--5."
#| output: true
#| echo: false

natural_mortality_matrix |>
  dplyr::filter(age %in% 1:5, year == get_start_year(data_4_model)) |>
  dplyr::select(-year)
```

The lognormal prior is available in FIMS, and thus, the age- and year-specific parameter vector could be mapped to a single value and estimated using a log-normal prior but that was not investigated here.

### Parameterization

FIMS models are configured with a data frame or R objects rather than an input file.

```{r}
#| label: setup-model
#| output: false

time_varying_selectivity <- TRUE
selectivity_used <- "Logistic"
FIMS::clear()
parameters <- FIMS::create_default_configurations(data = data_4_model) |>
  tidyr::unnest(data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = c("Fishery", "Acoustic_Survey"),
      module_type = c(selectivity_used, "Logistic")
    ),
    by = c("module_name", "fleet_name")
  ) |>
  FIMS::create_default_parameters(data = data_4_model) |>
  tidyr::unnest(data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = c("Maturity", "Maturity"),
      label = c("inflection_point", "slope"),
      value = c(
        hake_maturity_inflection_point,
        hake_maturity_slope
      )
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = c("Population"),
      label = c("log_M"),
      # estimation_type = "fixed_effects",
      value = log(dplyr::pull(natural_mortality_matrix, m)),
      time = dplyr::pull(natural_mortality_matrix, year),
      age = dplyr::pull(natural_mortality_matrix, age)
    ),
    by = c("module_name", "label", "time", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = FIMS::get_ages(data_4_model),
      value = log(init_naa),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = rep(c("Fishery", "Acoustic_Survey"), each = 2),
      label = rep(c("inflection_point", "slope"), 2),
      # TODO: play with these start values
      value = c(2.05, 3.17, 4.4, 3.3),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "Acoustic_Survey",
      label = c("log_q"),
      value = log(0.832),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  )
  
fleet_names <- get_fleets(data_4_model)

# Initialize lists to store fleet-related objects
fleet <- fleet_selectivity <-
  fleet_landings <- fleet_landings_distribution <-
  fleet_index <- fleet_index_distribution <-
  fleet_age_comp <- fleet_agecomp_distribution <-
  vector("list", length(fleet_names))

for (i in seq_along(fleet_names)) {
  # Selectivity
  fleet_selectivity[[i]] <- FIMS:::initialize_selectivity(
    parameters = parameters,
    data = data_4_model,
    fleet_name = fleet_names[i]
  )
  # Time-varying selectivity for the fishery
  if (i == 1 & time_varying_selectivity) {
    selectivity_type <- parameters |>
      dplyr::filter(module_name == "Selectivity", fleet_name == "Fishery") |>
      dplyr::slice(1) |>
      dplyr::pull(module_type)
    if (selectivity_type == "Logistic"){
      fleet_selectivity[[i]]$inflection_point$resize(get_n_years(data_4_model))
      for (ii in seq(get_n_years(data_4_model))) {
        fleet_selectivity[[i]]$inflection_point[ii]$value <- 1
        fleet_selectivity[[i]]$inflection_point[ii]$estimation_type$set("fixed_effects")
      }
      fleet_selectivity[[i]]$slope$resize(get_n_years(data_4_model))
      for (ii in seq(get_n_years(data_4_model))) {
        fleet_selectivity[[i]]$slope[ii]$value <- 1
        fleet_selectivity[[i]]$slope[ii]$estimation_type$set("fixed_effects")
      }
    }
    if (selectivity_type == "DoubleLogistic") {
      fleet_selectivity[[i]]$inflection_point_asc$resize(get_n_years(data_4_model))
      for (ii in seq(get_n_years(data_4_model))) {
        fleet_selectivity[[i]]$inflection_point_asc[ii]$value <- 1
        fleet_selectivity[[i]]$inflection_point_asc[ii]$estimation_type$set("fixed_effects")
      }
      # fleet_selectivity[[i]]$inflection_point_desc$resize(get_n_years(data_4_model))
      # for (ii in seq(get_n_years(data_4_model))) {
      #   fleet_selectivity[[i]]$inflection_point_desc[ii]$value <- 1
      #   fleet_selectivity[[i]]$inflection_point_desc[ii]$estimation_type$set("fixed_effects")
      # }
    }
  }

  # Setup the fleets
  fleet_module_ids <- c(
    selectivity = fleet_selectivity[[i]]$get_id()
  )
  fleet_types <- get_data(data_4_model) |>
    dplyr::filter(name == fleet_names[i]) |>
    dplyr::pull(type) |>
    unique()
  data_distribution_names_for_fleet_i <- parameters |>
    dplyr::filter(fleet_name == fleet_names[i] & distribution_type == "Data") |>
    dplyr::pull(module_type)

  # Landings
  if ("landings" %in% fleet_types &&
    "Landings" %in% data_distribution_names_for_fleet_i) {
    # Initialize landings module for the current fleet
    fleet_landings[[i]] <- FIMS:::initialize_landings(
      data = data_4_model,
      fleet_name = fleet_names[i]
    )
    # Add the module ID for the initialized landings to the list of fleet module IDs
    fleet_module_ids <- c(
      fleet_module_ids,
      c(landings = fleet_landings[[i]]$get_id())
    )
  }

  # Index
  if ("index" %in% fleet_types &&
    "Index" %in% data_distribution_names_for_fleet_i) {
    # Initialize index module for the current fleet
    fleet_index[[i]] <- FIMS:::initialize_index(
      data = data_4_model,
      fleet_name = fleet_names[i]
    )
    fleet_module_ids <- c(
      fleet_module_ids,
      c(index = fleet_index[[i]]$get_id())
    )
  }

  # Age composition
  if ("age_comp" %in% fleet_types &&
    "AgeComp" %in% data_distribution_names_for_fleet_i) {
    # Initialize age composition module for the current fleet
    fleet_age_comp[[i]] <- FIMS:::initialize_comp(
      data = data_4_model,
      fleet_name = fleet_names[i],
      type = "AgeComp"
    )
    fleet_module_ids <- c(
      fleet_module_ids,
      c(age_comp = fleet_age_comp[[i]]$get_id())
    )
  }

  fleet[[i]] <- FIMS:::initialize_fleet(
    parameters = parameters,
    data = data_4_model,
    fleet_name = fleet_names[i],
    linked_ids = fleet_module_ids
  )

  # Fleet uncertainty
  fleet_sd_input <- parameters |>
    dplyr::filter(fleet_name == fleet_names[i] & label == "log_sd") |>
    dplyr::mutate(
      label = "sd",
      value = exp(value)
    )

  if ("index" %in% fleet_types &&
    "Index" %in% data_distribution_names_for_fleet_i) {
    fleet_index_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = lognormal(link = "log"),
      sd = fleet_sd_input,
      data_type = "index"
    )
  }

  if ("landings" %in% fleet_types &&
    "Landings" %in% data_distribution_names_for_fleet_i) {
    fleet_landings_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = lognormal(link = "log"),
      sd = fleet_sd_input,
      data_type = "landings"
    )
  }

  if ("age_comp" %in% fleet_types &&
    "AgeComp" %in% data_distribution_names_for_fleet_i) {
    fleet_agecomp_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = multinomial(link = "logit"),
      data_type = "agecomp"
    )
  }
}

# Recruitment
recruitment <- methods::new(BevertonHoltRecruitment)
recruitment_process <- methods::new(LogDevsRecruitmentProcess)
recruitment$SetRecruitmentProcessID(recruitment_process$get_id())
# set up log_rzero (equilibrium recruitment)
recruitment$log_rzero[1]$value <- inputs[["ctl"]][["SR_parms"]]["SR_LN(R0)", "INIT"]
recruitment$log_rzero[1]$estimation_type$set("fixed_effects")
# set up logit_steep
recruitment$logit_steep[1]$value <- FIMS::logit(0.2, 1.0, steepness)
recruitment$logit_steep[1]$estimation_type$set("constant")
# recruit deviations should enter the model in normal space.
# The log is taken in the likelihood calculations
# Added an element to the log_devs vector because we need recruitment in the
# terminal year + 1
recruitment$log_devs$resize(get_n_years(data_4_model))
for (y in seq(recruitment$log_devs$size())) {
  recruitment$log_devs[y]$value <- 0
}
recruitment$log_devs$set_all_estimable(TRUE)
recruitment$n_years$set(get_n_years(data_4_model))
recruitment_distribution <- methods::new(DnormDistribution)
# TODO: check with Andrea about log space here
# logR_sd needs to enter the model logged b/c the exp() is
# taken before the likelihood calculation
recruitment_distribution$log_sd[1]$value <- 
  inputs[["ctl"]][["SR_parms"]]["SR_sigmaR", "INIT"]

# TODO: check with Andrea about this not being in the parameter vector
recruitment_distribution$log_sd[1]$estimation_type$set("fixed_effects")
recruitment_distribution$observed_values$resize(recruitment$log_devs$size())
recruitment_distribution$expected_values$resize(recruitment$log_devs$size())
for (i in seq(recruitment$log_devs$size())) {
  recruitment_distribution$observed_values[i]$value <- 0
  recruitment_distribution$expected_values[i]$value <- 0
}
recruitment_distribution$set_distribution_links(
  "random_effects",
  recruitment$log_devs$get_id()
)
recruitment$log_devs$set_all_random(TRUE)
# recruitment$log_devs[46]$estimation_type$set("constant")
# recruitment$log_devs[47]$estimation_type$set("constant")
# recruitment$log_devs[48]$estimation_type$set("constant")

# Put a prior on log_rzero
log_rzero_prior <- methods::new(DnormDistribution)
log_rzero_prior$expected_values$resize(1)
log_rzero_prior$expected_values[1]$value <- log_rzero_prior_mean
log_rzero_prior$log_sd$resize(1)
log_rzero_prior$log_sd[1]$value <- log(log_rzero_prior_sd)
log_rzero_prior$set_distribution_links(
  "prior",
  recruitment$log_rzero$get_id()
)

# Growth
growth <- FIMS:::initialize_growth(
  parameters = parameters,
  data = data_4_model
)

# Maturity
maturity <- FIMS:::initialize_maturity(
  parameters = parameters,
  data = data_4_model
)

# Population
population_module_ids <- c(
  recruitment = recruitment$get_id(),
  growth = growth$get_id(),
  maturity = maturity$get_id(),
  fleets = purrr::map(fleet, \(x) x$get_id())
)
population <- FIMS:::initialize_population(
  parameters = parameters,
  data = data_4_model,
  # TODO: need to remove linked_ids from the function and add module_id to the
  # parameters tibble
  linked_ids = population_module_ids
)
# # Put a prior on natural mortality
# M_prior <- methods::new(DlnormDistribution)
# M_prior$expected_values$resize(1)
# M_prior$expected_values[1]$value <- inputs[["ctl"]][["MG_parms"]][
#   grep("NatM_", row.names(inputs[["ctl"]][["MG_parms"]])),
#   "PRIOR"
# ]
# M_prior$log_sd$resize(1)
# M_prior$log_sd[1]$value <- inputs[["ctl"]][["MG_parms"]][
#   grep("NatM_", row.names(inputs[["ctl"]][["MG_parms"]])),
#   "PR_SD"
# ]
# M_prior$set_distribution_links(
#   "prior",
#   population$log_M$get_id()
# )
# Put a prior on init_naa
init_naa_prior <- methods::new(DnormDistribution)
init_naa_prior$expected_values$resize(1)
init_naa_prior$expected_values[1]$value <- init_naa_prior_mean
init_naa_prior$log_sd$resize(1)
init_naa_prior$log_sd[1]$value <- init_naa_prior_sd
init_naa_prior$set_distribution_links(
  "prior",
  population$log_init_naa$get_id()
)
fims_model <- methods::new(CatchAtAge)
fims_model$AddPopulation(population$get_id())

# Initialize the model
CreateTMBModel()
```

### Fit

#### Set up the fit

```{r}
#| label: setup-fit
#| output: false

# Create parameter list from Rcpp modules
parameter_list <- list(
  parameters = list(
    p = get_fixed(),
    re = get_random()
  ),
  model = fims_model
)
obj <- TMB::MakeADFun(
  data = list(),
  parameters = parameter_list$parameters,
  map = parameter_list$map,
  random = "re",
  DLL = "FIMS",
  silent = TRUE
)
parameter_names <- names(FIMS:::get_parameter_names(obj[["par"]]))
random_names <- names(FIMS:::get_random_names(obj[["env"]]$parList()$re))

map <- list(
  p = parameter_names,
  re = parameter_list[["parameters"]][["re"]]
)
# map[["p"]] <- map_time_varying(
#   map[["p"]],
#   "Population.1.log_M",
#   "single",
#   1:(get_n_ages(data_4_model) * get_n_years(data_4_model))
# )

if (time_varying_selectivity & TRUE) {
  # Make selectivity constant until 1991
  # selectivity_year <- 1981
  selectivity_year <- 2010

  year_vector <- get_start_year(data_4_model):get_end_year(data_4_model)
  year_sequence <- seq(year_vector)

  # # Change inflection_point early years to be one parameter
  map[["p"]] <- map_time_varying(
    map[["p"]],
    "Selectivity.1.inflection_point",
    "early",
    which(year_vector < selectivity_year)
  )
  # Change inflection_point late years to be one parameter
  map[["p"]] <- map_time_varying(
    map[["p"]],
    "Selectivity.1.inflection_point",
    "late",
    which(year_vector >= get_end_year(data_4_model) - 5)
  )
  # Change slope early years to be one parameter
  map[["p"]] <- map_time_varying(
    map[["p"]],
    "Selectivity.1.slope",
    "early",
    which(year_vector < selectivity_year)
  )
  # Change slope late years to be one parameter
  map[["p"]] <- map_time_varying(
    map[["p"]],
    "Selectivity.1.slope",
    "late",
    which(year_vector >= get_end_year(data_4_model) - 2)
  )
  map[["p"]] <- as.factor(map[["p"]])
  # map[["re"]] <- as.factor(gsub("\\d+$", "", random_names))
  # Bayesian estimation
  obj <- TMB::MakeADFun(
    data = list(),
    parameters = parameter_list$parameters,
    map = list(p = map[["p"]]),
    random = "re",
    DLL = "FIMS",
    silent = TRUE
  )
}
```

#### Maximum likelihood

```{r}
#| label: fit-mle
#| output: false

# MLE estimation using code from FIMS::fit_fims()
for (ii in 1:3) {
  # control$trace is reset to zero regardless of verbosity because the
  # differences in values printed out using control$trace will be
  # negligible between these different runs and is not worth printing
  opt <- with(
    obj,
    nlminb(
      start = if (ii == 1) {
        par
      } else {
        opt[["par"]]
      },
      objective = fn,
      gradient = gr,
      control = list(eval.max = 10000, iter.max = 10000, trace = 0)
    )
  )
}
sd_report <- TMB::sdreport(obj)
# Check that the standard errors are not NA
any(is.na(summary(
  TMB::sdreport(obj, skip.delta.method = TRUE),
  "fixed"
)[, 2]))

# ggplot2::ggplot(
#   data = data.frame(
#     year = get_start_year(data_4_model):(get_end_year(data_4_model) + 1),
#     spawning_biomass = sd_report$value[grepl("^spawning_biomass$", names(sd_report$value))]
#   ),
#   ggplot2::aes(x = year, y = spawning_biomass)
# ) + 
#   ggplot2::geom_line() +
#   ggplot2::geom_line(
#     data = data.frame(
#       year = as.numeric(names(ss3_sb)),
#       spawning_biomass = ss3_sb
#     ),
#     color = "red"
#   )
```

#### Bayesian

```{r}
#| label: fit-bayesian

# TODO: Implement this in parallel
fit <- SparseNUTS::sample_snuts(
  obj,
  num_samples = 8000,
  num_warmup = 1000,
  chains = 1,
  print = FALSE,
  refresh = 0,
  quiet = TRUE,
  control = list(adapt_delta = 0.99)
)

posterior <- as.data.frame(fit)
colnames(posterior) <- c(
  levels(map[["p"]]),
  paste0("Recruitment.1.recruitment_deviation.", obj$env$random)
)
dimnames(fit[["samples"]])[[3]] <- c(colnames(posterior), "lp__")

posteriors_labeled <- tidyr::pivot_longer(
  posterior[, -NCOL(posterior)],
  cols = dplyr::everything(),
  names_to = "name"
) |>
  dplyr::left_join(
    colnames(posterior) |>
      dplyr::as_tibble() |>
      tidyr::separate_wider_delim(
        delim = ".",
        cols = value,
        names = c("module_name", "fleet_number", "label", "module_id"),
        cols_remove = FALSE,
        too_few = "align_start"
      ) |>
      dplyr::group_by(fleet_number, label) |>
      dplyr::mutate(
        index = if (dplyr::n() == 1) {
          NA_integer_
        } else {
          dplyr::row_number(label)
        },
        fleet_name = dplyr::if_else(fleet_number == 1, "Fishery", "Acoustic_Survey"),
        words = gsub("_", " ", label)
      ) |>
      dplyr::ungroup(),
    by = c("name" = "value")
  ) |>
  dplyr::left_join(
    y = year_conversion_matrix,
    by = c("index")
  )

# Get derived quantities using report()
post <- apply(as.data.frame(posterior), 1, \(x) obj$report(x))
SB_long <- purrr::map(post, "spawning_biomass") |>
  purrr::map(
    \(x) data.frame(value = unlist(x), year_i = 1:length(unlist(x)))
  ) |>
  dplyr::bind_rows(.id = "simulation") |>
  dplyr::left_join(
    y = year_conversion_matrix,
    by = c("year_i" = "index")
  )
```

## Results

### Parameter plots

```{r}
#| label: plot-selectivity-survey
#| fig-cap: "Posterior estimates of the logistic selectivity curves for the survey."
#| fig-alt: "The slope of the survey curve was poorly estimated."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(module_name == "Selectivity", fleet_number == 2),
  ggplot2::aes(x = value)
) +
  tidybayes::stat_slabinterval(
    fill = "skyblue", 
    alpha = 0.6,
    .width = c(.66, .95) # Shows both 66% and 95% intervals
  ) +
  ggplot2::facet_wrap(fleet_name ~ words, scales = "free_x") +
  ggplot2::labs(title = "Posterior Distribution", x = "Parameter Value", y = "Density") +
  ggplot2::theme_minimal()
```

```{r}
#| label: plot-selectivity-fishery-inflection-point
#| fig-cap: "Posterior estimates of the inflection point (x-axis; age) of the logistic curve for time-varying fishery selectivity."
#| fig-alt: "Over the last decade we have been selecting younger fish."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      module_name == "Selectivity",
      label == "inflection_point",
      fleet_number == 1,
      year < get_end_year(data_4_model)
    ),
  ggplot2::aes(x = value, y = as.factor(year), fill = ggplot2::after_stat(x))
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Year", y = "Year", x = "Inflection point") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-selectivity-fishery-slope
#| fig-cap: "Posterior estimates of the slope (x-axis) of the logistic curve for time-varying fishery selectivity."
#| fig-alt: "There are a handful of years where the estimate is highly uncertain."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      module_name == "Selectivity",
      label == "slope",
      fleet_number == 1,
      year < get_end_year(data_4_model) +1
    ),
  ggplot2::aes(x = value, y = as.factor(year), fill = ggplot2::after_stat(x))
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Year", y = "Year", x = "Slope") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-log-Fmort
#| fig-cap: "Posterior estimates of the natural log of fishing mortality for each year (y axis). Black dots are median estimates from the Stock Synthesis model."
#| fig-alt: "Fishing levels have been increasing over time."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      label == "log_Fmort",
      fleet_number == 1
    ),
  ggplot2::aes(x = value, y = as.factor(year), fill = ggplot2::after_stat(x))
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::geom_point(
    data = dplyr::tibble(
      value = ss3_results[["log_Fmort"]],
      year = as.numeric(names(ss3_results[["log_Fmort"]]))
    ) |>
      dplyr::filter(
        year <= get_end_year(data_4_model),
        year >= get_start_year(data_4_model)
        
      ),
    mapping = ggplot2::aes(y = as.factor(year), x = value),
    color = "black",
    inherit.aes = FALSE,
    size = 2.2
  ) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Year", y = "Year", x = "Ln(Fishing Mortality)") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-init-naa
#| fig-cap: "Posterior estimates of the natural log of initial numbers at age."
#| fig-alt: "Older ages were poorly estimated."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      label == "log_init_naa"
    ),
  ggplot2::aes(
    x = value,
    y = factor(module_id, labels = get_ages(data_4_model)),
    fill = ggplot2::after_stat(x)
  )
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Age", y = "Age", x = "Ln(Initial Numbers at Age)") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-recruitment-deviations
#| fig-cap: "Posterior estimates of recruitment deviations (red) and median estimates from Stock Synthesis."
#| fig-alt: "The terminal year estimates were poorly estimated."
#| echo: false
#| warning: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      words == "recruitment deviation"
    ),
  ggplot2::aes(y = value, x = as.factor(year))
) +
  tidybayes::stat_pointinterval(
    # .width = c(0.66, 0.95),
    color = "darkred"
  ) +
  ggplot2::geom_point(
    data = dplyr::tibble(
      value = ss3_results[["rec_dev"]],
      year = as.numeric(names(ss3_results[["rec_dev"]])) - 1
    ) |>
      dplyr::filter(
        year <= get_end_year(data_4_model),
        year >= get_start_year(data_4_model)
        
      ),
    mapping = ggplot2::aes(x = as.factor(year), y = value),
    color = "black",
    inherit.aes = FALSE,
    size = 2.2
  ) +
  ggplot2::labs(x = "Year", y = "Recruitment Deviation") +
  ggplot2::theme_minimal() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

### Derived quantity plots

```{r}
#| label: plot-compare-spawning-biomass
#| fig-cap: "Comparisons of spawning biomass between the 2025 base model (points are median estimates) and the FIMS model (line with 95 percent credibility interval)."
#| fig-alt: "The FIMS model overestimates values from the 1980s compared to the SS3 model and underestimates more recent values."
#| echo: false
#| warning: false

ggplot2::ggplot(
  data = SB_long |>
    dplyr::group_by(year) |>
    dplyr::summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE)
    ),
  ggplot2::aes(x = year, y = median)
) +
  ggplot2::geom_line() +
  ggplot2::geom_ribbon(
    ggplot2::aes(ymin = lower, ymax = upper),
    alpha = 0.2
  ) +
  ggplot2::geom_point(
    data = dplyr::tibble(
      ss3_sb,
      year = as.numeric(names(ss3_sb))
    ),
    mapping = ggplot2::aes(x = year, y = ss3_sb)
  )
```

### Diagnostic plots

```{r}
#| label: plot-diagnostic-slow
#| fig-cap: "Pairs plots of the six slowest mixing parameters in the FIMS model. Most of the parameters shown relate to the estimation of initial numbers at age."
#| fig-alt: "Correlated parameters indicate large amounts of parameter spaces were covered by the chains."
#| echo: false
#| warning: false
SparseNUTS:::pairs.tmbfit(fit, pars = 1:6, order = "slow")
```

```{r}
#| label: plot-diagnostic-marginals
#| fig-cap: "Posterior distributions (bins) and marginal estimates (red lines) for the natural log of initial numbers at age for each age, where smaller numbered parameters pertain to younger ages."
#| fig-alt: "Older ages were poorly estimated."
#| echo: false
#| warning: false
SparseNUTS::plot_marginals(fit, pars = grep("log_init", dimnames(fit[["samples"]])[[3]]))
```

```{r}
#| label: plot-diagnostics-uncertainties
#| fig-cap: "Estimates of parameter uncertainty from the Bayesian estimation (x axis) compared to the same model fit using maximum likelihood estimation (MLE; y axis)."
#| fig-alt: "Parameters that were poorly estimated in an MLE context had even larger estimates of uncertainty in a Bayesian context."
#| echo: false
#| warning: false
SparseNUTS::plot_uncertainties(fit)
```

```{r}
#| label: cleanup
#| echo: false
FIMS::clear()
```
