---
title: U.S. West Coast Case Study of Pacific Hake
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Pacific Hake"

terminal_year <- 2024

hake_maturity_inflection_point <- 2.2
hake_maturity_slope <- 1.8
sample_size_multiplier <- 10
log_rzero_prior_mean <- 15
log_rzero_prior_sd <- 1.155
init_naa_prior_mean <- 18
init_naa_prior_sd <- 1.155
natural_mortality <- 0.235
steepness <- 0.8118
natural_mortality_by_age <- dplyr::tibble(
  age = 1:5,
  cannibalism = c(
    0.619078417, 0.038591177, 0.012472784, 0.010002492, 0.018692867
  )
)

# The model can be found within the following folder on the hake server 
# base_dir <- "/srv/hake/models/2025/01-version/01-base-models/01-base"
base_dir <- "https://raw.githubusercontent.com/pacific-hake/hake-assessment/refs/heads/model-files/models/2025"
maturity_data <- utils::read.csv("https://raw.githubusercontent.com/pacific-hake/hake-assessment/refs/heads/master/data-tables/maturity-ogives.csv")
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: U.S. West Coast
* Analyst: Kelli F. Johnson

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

The data for the FIMS model was created using `get_ss3_data()`, written by Drs. Ian G. Taylor and Megumi C. Oshima, on the input list generated by `r4ss::SS_read()` using model files from the 2025 assessment that can be found on GitHub.

```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false

inputs <- r4ss::SS_read(base_dir)
inputs[["wtatage"]] <- r4ss::SS_readwtatage(file.path(base_dir, "wtatage.ss_new"))

data_4_model <- get_ss3_data(
  ss3_inputs = inputs,
  fleets = seq(inputs[["dat"]][["Nfleets"]]),
  ages = c(0, inputs[["dat"]][["agebin_vector"]])
) |>
  dplyr::filter(
    timing <= inputs[["dat"]][["endyr"]],
    value != -999,
    # TODO: allow fleet3 but specify units as numbers
    name != "fleet3"
  ) |>
  dplyr::mutate(
    name = dplyr::case_when(
      name == "fleet1" ~ "fishery",
      name == "fleet2" ~ "survey"
    ),
    uncertainty = ifelse(
      test = type == "age" & name == "survey",
      uncertainty * 10,
      uncertainty
    )
  ) |>
  dplyr::filter(timing <= terminal_year, timing > 1976) |>
  FIMS::FIMSFrame()
```

```{r}
#| label: prepare-year-conversion-matrix
#| echo: false
year_conversion_matrix <- dplyr::tibble(
  year = get_start_year(data_4_model):(get_end_year(data_4_model) + 1)
) |>
  dplyr::mutate(
    index = dplyr::row_number()
  )
```

To get the operational model to more closely match a FIMS model the following changes were made:

### Weight-at-age data

Weight-at-age data was taken from the .ss_new file rather than the input file to Stock Synthesis just for ease of formatting. No simplifications were needed for the FIMS model regarding weight-at-age data because all of the weight-at-age matrices in the Stock Synthesis model are replicates of each other, e.g., weight at age for fleet two is the same as beginning and middle of the year population weight at age in this model. Currently, FIMS only uses a single weight-at-age matrix as there is no seasonality or timing in FIMS and weight-at-age is assumed to be the same across the population and fleets.

The last five years of weight-at-age data were averaged to provide weight-at-age information for the beginning of the year after the last year of landings were removed from the population.

### Maturity

For the 2025 base model, year-specific maturity at age was derived from a temperature-specific relationship. FIMS uses a single set of logistic maturity parameters for all years, and thus, the year-specific maturity ogives had to be translated to a single logistic curve. First, annual estimates of maturity at age for day `r max(maturity_data[["doy"]])` were averaged across years. Subsequently, this averaged ogive was fit using a logistic curve. This lead to a curve with an inflection of `r hake_maturity_inflection_point` and a slope of `r hake_maturity_slope` (Fig. @fig:maturity-plot). These values were specified as fixed inputs to the FIMS model.

Changes to the maturity model in FIMS can easily be accommodated but there are few populations that require changes, and thus, efforts have not been placed here. If desired, we could easily make the logistic parameters be time varying.

```{r}
#| label: maturity-plot
#| fig-cap: "Averaged maturity at age from the 2025 base model (empirical) compared to logistic maturity (logistic)."
#| fig-alt: "The two maturity ogives are quite similar with only small differences near the origin of lower empirical maturity for age-1 fish."
#| fig:align: center
#| echo: false

maturity_data |>
  dplyr::filter(doy == max(doy)) |>
  dplyr::group_by(age) |>
  dplyr::summarize(empirical = mean(p_mature)) |>
  dplyr::full_join(
    execute_logistic(
      get_ages(data_4_model),
      hake_maturity_slope,
      hake_maturity_inflection_point
    ) |>
      dplyr::rename(age = x, logistic = value),
    by = "age"
  ) |>
  tidyr::pivot_longer(cols = c("empirical", "logistic")) |>
  ggplot2::ggplot(ggplot2::aes(age, value, color = name)) +
  ggplot2::geom_line(linewidth = 1.5) +
  ggplot2::xlab("Age") +
  ggplot2::ylab("Probability of being mature") +
  ggplot2::scale_color_brewer(palette = "Dark2", name = "Method")
```

### Fisheries data

Any number of fleets, their landings, compositions, and catch-per-unit effort data can be accommodated in FIMS, and thus, regardless if the U.S. and Canadian fishery data are grouped into a single fleet, the fishery data can be accommodated in a FIMS model.

The Dirichlet-multinomial distribution is not yet available in FIMS, just the multinomial distribution is available to fit compositional data. The input sample sizes for the survey age-composition data were arbitrarily multiplied by `r sample_size_multiplier` to up-weight the data that are thought to be more informative of the population.

A separate analysis was conducted to investigate the fit of the data to each of the distributions available in Stock Synthesis. The results of this are still preliminary but the take-home message was that the compositional data can be fit using a multinomial distribution but this requires the use of iterative weighting rather than the internal estimation of the scaling parameter, which is conveniently available when the Dirichlet-multinomial distribution is used.

### Survey data

Same as the fishery data, the standard survey data for any number of surveys can be accommodated in a FIMS model. When this case study first started, FIMS could not accommodate the age-1 index of abundance because indices had to be in metric tons and not numbers. This is no longer a restriction in FIMS and data from any type of fleet, i.e., fishery or survey, data can be in numbers or biomass.

<!-- TODO: run a sensitivity with the age-1 index -->
Regardless, the age-1 index is not fit to in this case study because it is not in the 2025 base model.

No extra uncertainty was estimated for the index of abundance. 

### Fishing mortality

Annual fishing mortality was estimated using fixed effects.

### Recruitment

The natural log of initial recruitment was estimated as a fixed effect, initialized at the same value as the base model (`r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "INIT"]`). In the base model, this parameter is bounded between `r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "LO"]` and `r inputs$ctl$SR_parms[grepl("LN", row.names(inputs$ctl$SR_parms)), "HI"]`. Parameter bounds do not exist in FIMS but priors do, and thus, a normal prior with a mean of `r log_rzero_prior_mean` and a standard deviation of `r log_rzero_prior_sd`.

Steepness was fixed at the median estimate from the 2025 base model of `r steepness`. The beta prior used in the base model is not yet available in Stock Synthesis.

FIMS models are currently powered by Template Model Builder (TMB), which allows for the estimation of random effects, similar to models built using the Woods Hole Assessment Model (WHAM). We harnessed this capability and estimated recruitment deviations as a random effect with a normal distribution and the same $\sigma_{r}$ that is fixed in the Stock Synthesis model. It is recommended that $\sigma_{r}$ be estimated for random effects models but we did not investigate that here.
<!-- TOOD: investigate estimating sigma_r -->
<!-- TOOD: ask Andrea for the paper that suggests sigma_r can be estimated in stock assessment models when deviations are fit using random effects -->

```{r}
#| warning: false
#| label: prepare-initial-naa
#| output: false
init_naa <- exp(init_naa_prior_mean) *
  1000 *
  exp(-(get_ages(data_4_model) - 1) * natural_mortality)
init_naa[get_n_ages(data_4_model)] <- init_naa[get_n_ages(data_4_model)] /
  natural_mortality # sum of infinite series
```

The time series of recruitment deviations is one element shorter than the time series in Stock Synthesis because the initial numbers at age object include initial numbers for age-zero fish. The vector of the natural log of initial numbers at age was estimated using fixed effects and initialized at values in line with initial values for unfished recruitment and natural mortality. Additionally, a normal prior with a mean of `r init_naa_prior_mean` and a standard deviation of `r init_naa_prior_sd` was placed on the natural log of initial numbers at age in an attempt to aid model convergence.

The priors on the natural log of unfished recruitment and the natural log of initial numbers at age should be transformed using a Jacobian transformation because the parameters are in log space but that transformation is not currently available in FIMS. The Jacobian allows for adjustment of the scale of integration so the total measure is preserved.

### Selectivity

The logistic and double-logistic selectivity curves are currently the only selectivity forms that are currently available in FIMS. Both forms were investigated for Pacific Hake. We also investigated the use of a time-varying inflection point and/or slope for the fishing fleet. Additionally, we investigated allowing just a portion of the parameter vector to vary with time starting at several different time points, where the base model starts in 1991. In the end, time-invariant logistic selectivity was used for both the fleet and the survey. The data past `r terminal_year` were not well fit using the time-invariant logistic curve, and rather than spend time fitting to a curve that does not fit very well, we choose to truncate the data to `r terminal_year` and older. The remainder of the data will be fit to once age-specific selectivity is available in FIMS.

Work is in progress to implement an age-specific non-parametric selectivity curve in FIMS as there is a need for this option from both coasts. Consequently, we have put a high priority on it within the FIMS development team and it will be available before the 2027 Pacific Hake assessment.

### Natural mortality

<!-- TODO: Ensure that M is fixed at the median from the base model -->
<!-- TODO: Fix the parameter naming when parameters are mapped -->
<!-- TODO: Do a sensitivity where M is fixed at the values estimated from the RCEATTLE model -->
<!-- TODO: Do a sensitivity where M is estimated using a single value with the lognormal prior -->
Natural mortality in FIMS is age and year specific. To estimate a single parameter, the entire parameter vector could be mapped to a single value. That was not investigated here. Instead, natural mortality for all years was fixed at the sum of the median estimate of natural mortality from the 2025 base model (`r natural_mortality`) plus the mean estimate of cannibalism across all years for a given age group. The use of age-specific natural mortality was done to help stabilize the estimates of initial numbers at age, though the endeavour was not entirely successful. 

The lognormal prior is available in FIMS, and thus, the age- and year-specific parameter vector could be mapped to a single value and estimated using a log-normal prior but that was not investigated here.

```{r}
#| label: setup-natural-mortality
#| echo: false
natural_mortality_matrix <- natural_mortality_by_age |>
  tidyr::complete(
    age = get_ages(data_4_model),
    fill = list(cannibalism = 0)
  ) |>
  tidyr::crossing(
    year = get_start_year(data_4_model):get_end_year(data_4_model)
  ) |>
  dplyr::arrange(year, age) |>
  dplyr::mutate(m = cannibalism + natural_mortality)
```

## Set up the data

All data for a FIMS model is contained in one data frame, using a long format, i.e., one row per observation instead of a matrix format. Here, we call this object `data_4_model`. The creation of this data set is assisted by the R function called `get_ss3_data()`.

```{r}
#| warning: false
#| label: show-fims-data
#| output: true
#| echo: true

# Default print method of a FIMSFrame object
data_4_model

# Summary of the data types available
FIMS::get_data(data_4_model) |>
  dplyr::filter(value != -999) |>
  dplyr::group_by(type, name) |>
  dplyr::count()
```

## Run FIMS model

### Parameterization

```{r}
#| label: setup-model

time_varying_selectivity <- FALSE
FIMS::clear()
parameters <- FIMS::create_default_configurations(data = data_4_model) |>
  tidyr::unnest(data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = c("fishery", "survey"),
      module_type = c("Logistic", "Logistic")
    ),
    by = c("module_name", "fleet_name")
  ) |>
  FIMS::create_default_parameters(data = data_4_model) |>
  tidyr::unnest(data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = c("Maturity", "Maturity"),
      label = c("inflection_point", "slope"),
      value = c(
        hake_maturity_inflection_point,
        hake_maturity_slope
      )
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = c("Population"),
      label = c("log_M"),
      value = log(dplyr::pull(natural_mortality_matrix, m)),
      time = dplyr::pull(natural_mortality_matrix, year),
      age = dplyr::pull(natural_mortality_matrix, age)
    ),
    by = c("module_name", "label", "time", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = FIMS::get_ages(data_4_model),
      value = log(init_naa),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = rep(c("fishery", "survey"), each = 2),
      label = rep(c("inflection_point", "slope"), 2),
      # TODO: play with these start values
      value = c(2.05, 3.17, 4.4, 3.3),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "survey",
      label = c("log_q"),
      value = log(0.832),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  )
  
fleet_names <- get_fleets(data_4_model)

# Initialize lists to store fleet-related objects
fleet <- fleet_selectivity <-
  fleet_landings <- fleet_landings_distribution <-
  fleet_index <- fleet_index_distribution <-
  fleet_age_comp <- fleet_agecomp_distribution <-
  vector("list", length(fleet_names))

for (i in seq_along(fleet_names)) {
  # Selectivity
  fleet_selectivity[[i]] <- FIMS:::initialize_selectivity(
    parameters = parameters,
    data = data_4_model,
    fleet_name = fleet_names[i]
  )
  # Time-varying selectivity for the fishery
  if (i == 1 & time_varying_selectivity) {
    selectivity_type <- parameters |>
      dplyr::filter(module_name == "Selectivity", fleet_name == "fishery") |>
      dplyr::slice(1) |>
      dplyr::pull(module_type)
    inflection_point_prior <- methods::new(DlnormDistribution)
    inflection_point_prior$expected_values$resize(1)
    inflection_point_prior$expected_values[1]$value <- 1.354
    inflection_point_prior$log_sd$resize(1)
    inflection_point_prior$log_sd[1]$value <- 0.691
    if (selectivity_type == "Logistic"){
      fleet_selectivity[[i]]$inflection_point$resize(get_n_years(data_4_model))
      for (ii in seq(get_n_years(data_4_model))) {
        fleet_selectivity[[i]]$inflection_point[ii]$value <- 1
        fleet_selectivity[[i]]$inflection_point[ii]$estimation_type$set("fixed_effects")
      }
      # Put a prior on inflection_point
      # TODO: ensure that the prior is being implemented on the entire vector
      inflection_point_prior$set_distribution_links(
        "prior",
        fleet_selectivity[[i]]$inflection_point$get_id()
      )
    }
    if (selectivity_type == "DoubleLogistic") {
      fleet_selectivity[[i]]$inflection_point_asc$resize(get_n_years(data_4_model))
      for (ii in seq(get_n_years(data_4_model))) {
        fleet_selectivity[[i]]$inflection_point_asc[ii]$value <- 1
        fleet_selectivity[[i]]$inflection_point_asc[ii]$estimation_type$set("fixed_effects")
      }
      # fleet_selectivity[[i]]$inflection_point_desc$resize(get_n_years(data_4_model))
      # for (ii in seq(get_n_years(data_4_model))) {
      #   fleet_selectivity[[i]]$inflection_point_desc[ii]$value <- 1
      #   fleet_selectivity[[i]]$inflection_point_desc[ii]$estimation_type$set("fixed_effects")
      # }
      # Put a prior on inflection_point
      inflection_point_prior$set_distribution_links(
        "prior",
        fleet_selectivity[[i]]$inflection_point_asc$get_id()
      )
    }
  }

  # Setup the fleets
  fleet_module_ids <- c(
    selectivity = fleet_selectivity[[i]]$get_id()
  )
  fleet_types <- get_data(data_4_model) |>
    dplyr::filter(name == fleet_names[i]) |>
    dplyr::pull(type) |>
    unique()
  data_distribution_names_for_fleet_i <- parameters |>
    dplyr::filter(fleet_name == fleet_names[i] & distribution_type == "Data") |>
    dplyr::pull(module_type)

  # Landings
  if ("landings" %in% fleet_types &&
    "Landings" %in% data_distribution_names_for_fleet_i) {
    # Initialize landings module for the current fleet
    fleet_landings[[i]] <- FIMS:::initialize_landings(
      data = data_4_model,
      fleet_name = fleet_names[i]
    )
    # Add the module ID for the initialized landings to the list of fleet module IDs
    fleet_module_ids <- c(
      fleet_module_ids,
      c(landings = fleet_landings[[i]]$get_id())
    )
  }

  # Index
  if ("index" %in% fleet_types &&
    "Index" %in% data_distribution_names_for_fleet_i) {
    # Initialize index module for the current fleet
    fleet_index[[i]] <- FIMS:::initialize_index(
      data = data_4_model,
      fleet_name = fleet_names[i]
    )
    fleet_module_ids <- c(
      fleet_module_ids,
      c(index = fleet_index[[i]]$get_id())
    )
  }

  # Age composition
  if ("age_comp" %in% fleet_types &&
    "AgeComp" %in% data_distribution_names_for_fleet_i) {
    # Initialize age composition module for the current fleet
    fleet_age_comp[[i]] <- FIMS:::initialize_comp(
      data = data_4_model,
      fleet_name = fleet_names[i],
      type = "AgeComp"
    )
    fleet_module_ids <- c(
      fleet_module_ids,
      c(age_comp = fleet_age_comp[[i]]$get_id())
    )
  }

  fleet[[i]] <- FIMS:::initialize_fleet(
    parameters = parameters,
    data = data_4_model,
    fleet_name = fleet_names[i],
    linked_ids = fleet_module_ids
  )

  # Fleet uncertainty
  fleet_sd_input <- parameters |>
    dplyr::filter(fleet_name == fleet_names[i] & label == "log_sd") |>
    dplyr::mutate(
      label = "sd",
      value = exp(value)
    )

  if ("index" %in% fleet_types &&
    "Index" %in% data_distribution_names_for_fleet_i) {
    fleet_index_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = lognormal(link = "log"),
      sd = fleet_sd_input,
      data_type = "index"
    )
  }

  if ("landings" %in% fleet_types &&
    "Landings" %in% data_distribution_names_for_fleet_i) {
    fleet_landings_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = lognormal(link = "log"),
      sd = fleet_sd_input,
      data_type = "landings"
    )
  }

  if ("age_comp" %in% fleet_types &&
    "AgeComp" %in% data_distribution_names_for_fleet_i) {
    fleet_agecomp_distribution[[i]] <- FIMS:::initialize_data_distribution(
      module = fleet[[i]],
      family = multinomial(link = "logit"),
      data_type = "agecomp"
    )
  }
}

# Recruitment
recruitment <- methods::new(BevertonHoltRecruitment)
recruitment_process <- methods::new(LogDevsRecruitmentProcess)
recruitment$SetRecruitmentProcessID(recruitment_process$get_id())
# set up log_rzero (equilibrium recruitment)
recruitment$log_rzero[1]$value <- inputs[["ctl"]][["SR_parms"]]["SR_LN(R0)", "INIT"]
recruitment$log_rzero[1]$estimation_type$set("fixed_effects")
# set up logit_steep
recruitment$logit_steep[1]$value <- FIMS::logit(0.2, 1.0, steepness)
recruitment$logit_steep[1]$estimation_type$set("constant")
# recruit deviations should enter the model in normal space.
# The log is taken in the likelihood calculations
# Added an element to the log_devs vector because we need recruitment in the
# terminal year + 1
recruitment$log_devs$resize(get_n_years(data_4_model))
for (y in seq(recruitment$log_devs$size())) {
  recruitment$log_devs[y]$value <- 0
}
recruitment$log_devs$set_all_estimable(TRUE)
recruitment$n_years$set(get_n_years(data_4_model))
recruitment_distribution <- methods::new(DnormDistribution)
# TODO: check with Andrea about log space here
# logR_sd needs to enter the model logged b/c the exp() is
# taken before the likelihood calculation
recruitment_distribution$log_sd <- methods::new(ParameterVector, 1)
recruitment_distribution$log_sd[1]$value <- log(
  inputs[["ctl"]][["SR_parms"]]["SR_sigmaR", "INIT"]
)
# TODO: check with Andrea about this not being in the parameter vector
recruitment_distribution$log_sd[1]$estimation_type$set("constant")
recruitment_distribution$x$resize(recruitment$log_devs$size())
recruitment_distribution$expected_values$resize(recruitment$log_devs$size())
for (i in seq(recruitment$log_devs$size())) {
  recruitment_distribution$x[i]$value <- 0
  recruitment_distribution$expected_values[i]$value <- 0
}
recruitment_distribution$set_distribution_links(
  "random_effects",
  recruitment$log_devs$get_id()
)
recruitment$log_devs$set_all_random(TRUE)

# Put a prior on log_rzero
log_rzero_prior <- methods::new(DnormDistribution)
log_rzero_prior$expected_values$resize(1)
log_rzero_prior$expected_values[1]$value <- log_rzero_prior_mean
log_rzero_prior$log_sd$resize(1)
log_rzero_prior$log_sd[1]$value <- log(log_rzero_prior_sd)
log_rzero_prior$set_distribution_links(
  "prior",
  recruitment$log_rzero$get_id()
)

# Growth
growth <- FIMS:::initialize_growth(
  parameters = parameters,
  data = data_4_model
)
# Add weight-at-age data for the terminal year + 1 b/c beginning of the year
# calculations
growth$n_years$set(get_n_years(data_4_model) + 1)
growth$weights$resize(growth$weights$size() + get_n_ages(data_4_model))
ii <- 0
for(i in (growth$weights$size() - get_n_ages(data_4_model)):(growth$weights$size() - 1)) {
  growth$weights$set(
    i,
    get_data(data_4_model) |>
      dplyr::filter(
        type == "weight-at-age",
        timing %in% (get_end_year(data_4_model) - 4):get_end_year(data_4_model),
        age == ii
      ) |>
      dplyr::pull(value) |>
      mean()
  )
  ii <- ii + 1
}

# Maturity
maturity <- FIMS:::initialize_maturity(
  parameters = parameters,
  data = data_4_model
)

# Population
population_module_ids <- c(
  recruitment = recruitment$get_id(),
  growth = growth$get_id(),
  maturity = maturity$get_id(),
  fleets = purrr::map(fleet, \(x) x$get_id())
)
population <- FIMS:::initialize_population(
  parameters = parameters,
  data = data_4_model,
  # TODO: need to remove linked_ids from the function and add module_id to the
  # parameters tibble
  linked_ids = population_module_ids
)
# Put a prior on init_naa
init_naa_prior <- methods::new(DnormDistribution)
init_naa_prior$expected_values$resize(1)
init_naa_prior$expected_values[1]$value <- init_naa_prior_mean
init_naa_prior$log_sd$resize(1)
init_naa_prior$log_sd[1]$value <- log(init_naa_prior_sd)
init_naa_prior$set_distribution_links(
  "prior",
  population$log_init_naa$get_id()
)
fims_model <- methods::new(CatchAtAge)
fims_model$AddPopulation(population$get_id())

# Initialize the model
CreateTMBModel()
```

### Fit

#### Set up the fit

```{r}
#| label: setup-fit

# Create parameter list from Rcpp modules
parameter_list <- list(
  parameters = list(
    p = get_fixed(),
    re = get_random()
  ),
  model = fims_model
)
# fit <- FIMS::fit_fims(input = parameter_list, optimize = TRUE)
obj <- TMB::MakeADFun(
  data = list(),
  parameters = parameter_list$parameters,
  map = parameter_list$map,
  random = "re",
  DLL = "FIMS",
  silent = TRUE
)
parameter_names <- names(FIMS:::get_parameter_names(obj[["par"]]))
random_names <- names(FIMS:::get_random_names(obj[["env"]]$parList()$re))
```

#### Maximum likelihood

```{r}
#| label: fit-mle

# MLE estimation using code from FIMS::fit_fims()
for (ii in 1:3) {
  # control$trace is reset to zero regardless of verbosity because the
  # differences in values printed out using control$trace will be
  # negligible between these different runs and is not worth printing
  opt <- with(
    obj,
    nlminb(
      start = if (ii == 1) {
        par
      } else {
        opt[["par"]]
      },
      objective = fn,
      gradient = gr,
      control = list(eval.max = 10000, iter.max = 10000, trace = 0)
    )
  )
}
FIMS::set_fixed(opt$par)
sd_report <- TMB::sdreport(obj)
# Check that the standard errors are not NA
any(is.na(summary(
  TMB::sdreport(obj, skip.delta.method = TRUE),
  "fixed"
)[, 2]))

fit <- FIMSFit(
  input = parameter_list,
  obj = obj,
  opt = opt,
  sdreport = sd_report
)
output <- FIMS::get_estimates(fit) |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + FIMS::get_start_year(data_4_model) - 1,
    estimate = estimated
  )
```

#### Bayesian

```{r}
#| label: fit-bayesian
#| eval: false

# TODO: Implement this in parallel
mcmc <- adnuts::sample_snuts(
  obj,
  num_samples = 5000,
  num_warmup = 200,
  chains = 4,
  cores = 1,
  seed = 1,
  # init = "random",
  control = list(adapt_delta = .95)
)

posterior <- adnuts::extract_samples(
  mcmc,
  inc_lp = TRUE,
  inc_warmup = FALSE,
  unbounded = FALSE
)
colnames(posterior) <- c(
  parameter_names,
  paste0("Recruitment.1.recruitment_deviation.", obj$env$random),
  "likelihood"
)

posteriors_labeled <- tidyr::pivot_longer(
  posterior[, -NCOL(posterior)],
  cols = dplyr::everything(),
  names_to = "name"
) |>
  dplyr::left_join(
    colnames(posterior) |>
      dplyr::as_tibble() |>
      tidyr::separate_wider_delim(
        delim = ".",
        cols = value,
        names = c("module_name", "fleet_number", "label", "module_id"),
        cols_remove = FALSE,
        too_few = "align_start"
      ) |>
      dplyr::group_by(fleet_number, label) |>
      dplyr::mutate(
        index = dplyr::case_when(
          dplyr::n() == 1 ~ NA_integer_,
          TRUE ~ dplyr::row_number(label)
        ),
        fleet_name = dplyr::if_else(fleet_number == 1, "fishery", "survey"),
        words = gsub("_", " ", label)
      ) |>
      dplyr::ungroup(),
    by = c("name" = "value")
  ) |>
  dplyr::left_join(
    y = year_conversion_matrix,
    by = c("index")
  )

post <- apply(as.data.frame(mcmc), 1, \(x) obj$report(x))
SB_long <- purrr::map(post, "spawning_biomass") |>
  purrr::map(
    \(x) data.frame(value = unlist(x), year_i = 1:length(unlist(x)))
  ) |>
  dplyr::bind_rows(.id = "simulation") |>
  dplyr::left_join(
    y = year_conversion_matrix,
    by = c("year_i" = "index")
  )
```

#### Mapped parameters

Code to map off parameters to another parameter is the same as "mirroring" parameters in Stock Synthesis.

```{r}
#| label: fit-mapped
#| echo: false
#| eval: false

if (time_varying_selectivity & FALSE) {
  # Make selectivity constant until 1991
  selectivity_year <- 1991
  map <- parameter_list$parameters
  map$p[grepl(parameter_names, pattern = ".1.inflection_point")] <- c(
  # map$p[grepl(parameter_names, pattern = "inflection_point_asc")] <- c(
    rep(2, length(get_start_year(data_4_model):(selectivity_year - 1))),
    rnorm(
      length(selectivity_year:get_end_year(data_4_model)),
      mean = 2,
      sd = 0.1
    ) + 0.001
  )
  map$p[grepl(pattern = "log_Fmort", parameter_names)] <- rnorm(
    length(grep(pattern = "log_Fmort", parameter_names)),
    mean = -1,
    sd = 0.05
  )
  map$p <- as.factor(map$p)
  map$re <- as.factor(map$re)
  # Bayesian estimation
  obj <- TMB::MakeADFun(
    data = list(),
    parameters = parameter_list$parameters,
    map = map,
    random = "re",
    DLL = "FIMS",
    silent = TRUE
  )
}
# MLE estimation
for (ii in 1:3) {
  # control$trace is reset to zero regardless of verbosity because the
  # differences in values printed out using control$trace will be
  # negligible between these different runs and is not worth printing
  opt <- with(
    obj,
    nlminb(
      start = if (ii == 1) {
        par
      } else {
        opt[["par"]]
      },
      objective = fn,
      gradient = gr,
      control = list(
        eval.max = 10000,
        iter.max = 10000,
        trace = 0
      )
    )
  )
}
```

## Summarize results

### Plotting using {stockplotr}

```{r}
#| label: comparison-plots
#| echo: false

stockplotr::plot_landings(
  dplyr::filter(output, module_id == 1)
) +
  ggplot2::geom_point(
    data = dplyr::filter(output, module_id == 1, label == "landings_expected"),
    ggplot2::aes(x = year, y = observed)
  )

stockplotr::plot_spawning_biomass(
  dplyr::filter(output, label == "spawning_biomass")
)

stockplotr::plot_recruitment(
  output
)

# index
stockplotr::plot_timeseries(
  stockplotr::filter_data(
    output |> dplyr::filter(module_id == 2),
    label_name = "^index_expected$",
    geom = "line"
  ),
  x = "year",
  y = "estimated",
  ylab = "Relative Index of Abundance"
) +
  ggplot2::geom_point(
    data = dplyr::filter(
      FIMS::get_estimates(fit),
      module_id == 2,
      label == "index_expected",
      observed > -999
    ) |>
      dplyr::left_join(
        year_conversion_matrix,
        by = c("year_i" = "index")
      ),
    ggplot2::aes(x = year, y = observed)
  ) +
  stockplotr::theme_noaa()

# Fishing mortality
stockplotr::plot_timeseries(
  stockplotr::filter_data(
    output |> dplyr::filter(module_id == 1),
    label_name = "log_Fmort$",
    geom = "line"
  ),
  x = "year",
  y = "estimate",
  ylab = "natural log of Fishing Mortality"
) +
  stockplotr::theme_noaa()
```

### Plotting Bayesian results

#### Parameter plots

```{r}
#| label: plot-selectivity
#| fig-cap: "Posterior estimates of the logistic selectivity curves for the fishery and the survey."
#| fig-alt: "The slope of the survey curve was poorly estimated."
#| echo: false

# Posterior estimates of selectivity
ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(module_name == "Selectivity"),
  ggplot2::aes(x = value)
) +
  tidybayes::stat_slabinterval(
    fill = "skyblue", 
    alpha = 0.6,
    .width = c(.66, .95) # Shows both 66% and 95% intervals
  ) +
  ggplot2::facet_wrap(fleet_name ~ words, scales = "free_x") +
  ggplot2::labs(title = "Posterior Distribution", x = "Parameter Value", y = "Density") +
  ggplot2::theme_minimal()
```

```{r}
#| label: plot-selectivity-curves
#| fig-cap: "Estimated selectivity curves for the fishery and the survey using logistic selectivity."
#| fig-alt: "The survey selects younger fish than the fishery but they both select older fish equally."
#| echo: false

temp <- purrr::map2(
  .x = dplyr::filter(
    output,
    module_name == "Selectivity",
    label == "slope"
  ) |>
    dplyr::pull(estimated),
  .y = dplyr::filter(
    output,
    module_name == "Selectivity",
    label == "inflection_point"
  ) |>
    dplyr::pull(estimated),
  .f = execute_logistic,
  x = get_ages(data_4_model)
)
names(temp) <- get_fleets(data_4_model)

g <- ggplot2::ggplot(
  data = dplyr::bind_rows(temp, .id = "fleet"),
  ggplot2::aes(x = x, y = value, group = fleet, color = fleet)
) +
  ggplot2::geom_line()
g
```

```{r}
#| label: plot-log-Fmort
#| fig-cap: "Posterior estimates of the natural log of fishing mortality for each year (y axis)."
#| fig-alt: "The terminal year was poorly estimated."
#| echo: false

# Posterior estimates of fishing mortality
ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      label == "log_Fmort"
    ),
  ggplot2::aes(x = value, y = as.factor(year), fill = ggplot2::after_stat(x))
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Year", y = "Year", x = "Ln(Fishing Mortality)") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-init-naa
#| fig-cap: "Posterior estimates of the natural log of initial numbers at age."
#| fig-alt: "Older ages were poorly estimated."
#| echo: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      label == "log_init_naa"
    ),
  ggplot2::aes(
    x = value,
    y = factor(module_id, labels = get_ages(data_4_model)),
    fill = ggplot2::after_stat(x)
  )
) +
  ggridges::geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  ggplot2::scale_fill_viridis_c(name = "Value", option = "C") +
  ggplot2::labs(title = "Posterior Shift by Age", y = "Age", x = "Ln(Initial Numbers at Age)") +
  ggridges::theme_ridges()
```

```{r}
#| label: plot-recruitment-deviations
#| fig-cap: "Posterior estimates of recruitment deviations."
#| fig-alt: "The terminal year estimates were poorly estimated."
#| echo: false

ggplot2::ggplot(
  posteriors_labeled |>
    dplyr::filter(
      words == "recruitment deviation"
    ),
  ggplot2::aes(y = value, x = as.factor(year))
) +
  tidybayes::stat_pointinterval(
    # .width = c(0.66, 0.95),
    color = "darkred"
  ) +
  ggplot2::labs(x = "Year", y = "Recruitment Deviation") +
  ggplot2::theme_minimal() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

#### Derived quantity plots

```{r}
#| label: plot-compare-spawning-biomass
#| fig-cap: "Comparisons of spawning biomass between the 2025 base model (points are median estimates) and the FIMS model (line with 95 percent credibility interval)."
#| fig-alt: "The FIMS model overestimates values from the 1980s compared to the SS3 model and underestimates more recent values."
#| echo: false

# Hake results are in million metric tons, i.e., 1,000,000 t
ss3_sb_median <- c(
  "1966" = 0.994896, "1967" = 1.00664, "1968" = 1.01708, "1969" = 1.162515,
  "1970" = 1.33501, "1971" = 1.36887, "1972" = 1.560225, "1973" = 1.853525,
  "1974" = 1.79134, "1975" = 1.548275, "1976" = 2.237185, "1977" = 2.15888,
  "1978" = 1.784635, "1979" = 1.92232, "1980" = 1.813545, "1981" = 1.736695,
  "1982" = 1.875015, "1983" = 2.393425, "1984" = 2.408205, "1985" = 2.40912,
  "1986" = 2.41643, "1987" = 2.4756, "1988" = 2.445505, "1989" = 2.11734,
  "1990" = 2.092725, "1991" = 1.89941, "1992" = 1.726115, "1993" = 1.34966,
  "1994" = 1.30146, "1995" = 1.159205, "1996" = 1.12109, "1997" = 1.113265,
  "1998" = 0.9282575, "1999" = 0.798846, "2000" = 0.925014, "2001" = 1.36914,
  "2002" = 1.955155, "2003" = 1.825095, "2004" = 1.48332, "2005" = 1.17802,
  "2006" = 0.9373315, "2007" = 0.7403355, "2008" = 0.6686465, "2009" = 0.6280765,
  "2010" = 0.7547615, "2011" = 0.7524495, "2012" = 0.9411375, "2013" = 1.728305,
  "2014" = 2.0092, "2015" = 1.46967, "2016" = 1.19547, "2017" = 1.640085,
  "2018" = 1.752535, "2019" = 1.439125, "2020" = 1.35688, "2021" = 1.120755,
  "2022" = 1.159235, "2023" = 1.47779, "2024" = 1.638705, "2025" = 1.736495,
  "2026" = 1.245515, "2027" = 0.865635, "2028" = 0.5916625
) * 1000000

ggplot2::ggplot(
  data = SB_long |>
    dplyr::group_by(year) |>
    dplyr::summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE)
    ),
  ggplot2::aes(x = year, y = median)
) +
  ggplot2::geom_line() +
  ggplot2::geom_ribbon(
    ggplot2::aes(ymin = lower, ymax = upper),
    alpha = 0.2
  ) +
  ggplot2::geom_point(
    data = dplyr::tibble(
      ss3_sb_median,
      year = as.numeric(names(ss3_sb_median))
    ),
    mapping = ggplot2::aes(x = year, y = ss3_sb_median)
  )
```

#### Diagnostic plots

```{r}
#| label: plot-diagnostic-slow
#| fig-cap: "Pairs plots of the six slowest mixing parameters in the FIMS model. Most of the parameters shown relate to the estimation of initial numbers at age."
#| fig-alt: "Correlated parameters indicate large amounts of parameter spaces were covered by the chains."
#| echo: false
adnuts::pairs_admb(mcmc, pars = 1:6, order = "slow")
```

```{r}
#| label: plot-diagnostic-marginals
#| fig-cap: "Posterior distributions (bins) and marginal estimates (red lines) for the natural log of initial numbers at age for each age, where smaller numbered parameters pertain to younger ages."
#| fig-alt: "Older ages were poorly estimated."
#| echo: false
adnuts::plot_marginals(mcmc, pars = grep("log_init", parameter_names))
```

```{r}
#| label: plot-diagnostics-uncertainties
#| fig-cap: "Estimates of parameter uncertainty from the Bayesian estimation (x axis) compared to the same model fit using maximum likelihood estimation (MLE; y axis)."
#| fig-alt: "Parameters that were poorly estimated in an MLE context had even larger estimates of uncertainty in a Bayesian context."
adnuts::plot_uncertainties(mcmc)
```

## Clean up

Memory is allocated when FIMS modules are initialized and must be cleared using `FIMS::clear()`. Though, closing your R session will also clear the memory.
```{r}
#| label: cleanup
FIMS::clear()
```
