---
title: "Untitled"
format: html
---

```{r fims1, warning=FALSE, message=FALSE}
# automatically loads fims Rcpp module
remotes::install_github("NOAA-FIMS/FIMS", ref = "dev-length-wrapper-and-test")
library(FIMS)
library(TMB)
library(dplyr)
library(tidyr)
library(r4ss)
# clear memory
clear()

```
## Setting up Data
```{r fims-dims}
load(file.path(getwd(), "content", "data_files", "opaka_length.RDS"))
include_age_comps <- FALSE
include_length_comps <- TRUE
years <- seq(opaka_dat$styr, opaka_dat$endyr)
nyears <- 75 # the number of years which we have data for.
nseasons <- 1 # the number of seasons in each year. FIMS currently defaults to 1
ages <- seq(1, 21) # age vector.
nages <- length(ages) # the number of age groups.
if(include_length_comps){
  comp_lengths <- opaka_dat$lbin_vector # length vector.
  nlengths <- length(comp_lengths) # the number of length bins.
}
```

### Preparing Data using FIMSFrame

 We will be reading data into the model using the FIMSFrame S4 R class set up in [R/fimsframe.R](https://github.com/NOAA-FIMS/FIMS/blob/main/R/fimsframe.R)

```{r fimsframe}
# use FIMS data frame

source("./content/R/get_ss3_data.R")
opaka_dat_fims <- get_ss3_data(opaka_dat, fleets = c(1,2,3), ages = ages, lengths = comp_lengths)
## age to length conversion matrix 
# Growth function values to create age to length conversion matrix from model
#comparison project
mg_pars <- rep$parameters  |> 
dplyr::filter(stringr::str_detect(Label, "_GP_"))
Linf <- mg_pars$Value[3]
K <- mg_pars$Value[4]
a0 <- -0.29
amax <- 21
cv <- mg_pars$Value[5]

L2Wa <- mg_pars$Value[7]
L2Wb <- mg_pars$Value[8]

AtoL <- function(a,Linf,K,a_0){
  L <- Linf*(1-exp(-K*(a-a_0)))
}

ages <- 1:amax
len_bins <- comp_lengths

#Create length at age conversion matrix and fill proportions using above 
#growth parameters
length_age_conversion <- matrix(NA,nrow=length(ages),ncol=length(len_bins))
for(i in seq_along(ages)){
  #Calculate mean length at age to spread lengths around
  mean_length <- AtoL(ages[i],Linf,K,a0)
  #mean_length <- AtoLSchnute(ages[i],L1,L2,a1,a2,Ks)
  #Calculate the cumulative proportion shorter than each composition length
  temp_len_probs<-pnorm(q=len_bins,mean=mean_length,sd=mean_length*cv)
  #Reset the first length proportion to zero so the first bin includes all
  #density smaller than that bin
  temp_len_probs[1]<-0
  #subtract the offset length probabilities to calculate the proportion in each
  #bin. For each length bin the proportion is how many fish are larger than this
  #length but shorter than the next bin length.
  temp_len_probs <- c(temp_len_probs[-1],1)-temp_len_probs
  length_age_conversion[i,] <- temp_len_probs
}
colnames(length_age_conversion) <- len_bins
rownames(length_age_conversion) <- ages

#Extract years and fleets from milestone 1 data
start_date <- unique(opaka_dat_fims$datestart[opaka_dat_fims$type=="landings"])
end_date <- unique(opaka_dat_fims$dateend[opaka_dat_fims$type=="landings"])
observers <- unique(opaka_dat_fims$name[opaka_dat_fims$type=="length"])

#Create data frame for new fleet and year specific length at age conversion proportions
#These are identical across years and fleets in this default example
length_age_data <- opaka_dat_fims[rep(1,length(ages)*
                                    length(len_bins)*
                                    length(start_date)*
                                    length(observers)),]

row_index <- 1
for(k in seq_along(start_date)){
  for(l in seq_along(observers)){
    for(i in seq_along(ages)){
      for(j in seq_along(len_bins)){
        length_age_data[row_index,] <- data.frame(
                                        type = "age-to-length-conversion",
                                        name = observers[l],
                                        age = ages[i],
                                        length = len_bins[j],
                                        datestart = start_date[k],
                                        dateend = end_date[k],
                                        value = length_age_conversion[i,j],
                                        unit = "proportion",
                                        uncertainty = 30) #is this neff? can it vary by year?
        row_index <- row_index + 1
      }
    }
  }
}
opaka_dat_fims <- rbind(opaka_dat_fims, length_age_data)
#save(list = "opaka_dat_fims", file = file.path(getwd(), "content", "opaka_dat_fims.RData"))
load(file = file.path(getwd(), "content", "opaka_dat_fims.RData"))
fims_frame <- FIMSFrame(opaka_dat_fims)
```

The `fims_frame` object contains a `@data` slot that holds a long data frame with:
* 3 fleets: commercial fishery (fleet1), survey (fleet2), non-commercial fishery (fleet3)
* landings for fleets 1 and 3
* cpue for fleets 1 and 2
* length composition data for fleet 2


### Prepare Parameters using `create_default_parameters()`

```{r, max.height='100px', attr.output='.numberLines'}
# Define fleet specifications for fleet1 and survey1
fleet1 <- fleet2 <- fleet3 <- list(
  selectivity = list(form = "LogisticSelectivity"),
  data_distribution = c(
    Index = "DlnormDistribution",
    # Remove this next line once dev has the appropriate if statement
    AgeComp = "DmultinomDistribution", 
    LengthComp = "DmultinomDistribution"  ##CHECK
  )
)

# Create default parameters
default_parameters <- fims_frame |>
  create_default_parameters(
    fleets = list(fleet1 = fleet1, fleet2 = fleet2, fleet3 = fleet3)
  )
```

```{r modify-default-parameters}

recdevs <- rep$parameters |> filter(stringr::str_detect(Label, "RecrDev")) |> select(Label, Value)
init_naa <- exp(opaka_ctl$SR_parms["SR_LN(R0)", "INIT"]) * exp(-(ages - 1) * 0.135)
init_naa[nages] <- init_naa[nages] / 0.135

parameters <- default_parameters |>
  update_parameters(
    modified_parameters = list(
      fleet1 = list(
        Fleet.log_Fmort.value = log(rep$exploitation$FRS),
        LogisticSelectivity.inflection_point.value = 36.01,
        LogisticSelectivity.slope.value = 3.46
      )
    )
  ) |>
    update_parameters(
    modified_parameters = list(
      fleet2 = list(
        LogisticSelectivity.inflection_point.value = 2.897,
        LogisticSelectivity.slope.value = 3.254,
        Fleet.log_q.value = -4.12772
      )
    )
  ) |>
  update_parameters(
    modified_parameters = list(
      fleet3 = list(
        Fleet.log_Fmort.value = log(rep$exploitation$Non_comm),
        LogisticSelectivity.inflection_point.value = 40,
        LogisticSelectivity.slope.value = 11
      )
    )
  ) |>
  update_parameters(
    modified_parameters = list(
      recruitment = list(
        BevertonHoltRecruitment.log_rzero.value = opaka_ctl$SR_parms["SR_LN(R0)", "INIT"],
        BevertonHoltRecruitment.log_devs.value = recdevs$Value,
        BevertonHoltRecruitment.log_devs.estimated = FALSE
      )
    )
  ) |>
  update_parameters(
    modified_parameters = list(
      maturity = list(
        LogisticMaturity.inflection_point.value = 3.5,
        LogisticMaturity.inflection_point.estimated = FALSE,
        LogisticMaturity.slope.value = 3,
        LogisticMaturity.slope.estimated = FALSE
      )
    )
  ) |>
  update_parameters(
    modified_parameters = list(
      population = list(
        Population.log_init_naa.value = log(init_naa),
        Population.log_M.value = rep(0.135, get_n_years(fims_frame)*get_n_ages(fims_frame)) 
      )
    )
  )


```

## Initialize modules and fit the model

With data and parameters in place, we can now initialize modules using `initialize_fims()` and fit the model using `fit_fims()`.
```{r, max.height='100px', attr.output='.numberLines', fit-fims, eval = TRUE}
# Run the model without optimization to help ensure a viable model
test_fit <- parameters |>
  initialize_fims(data = fims_frame) |>
  fit_fims(optimize = FALSE)

# Run the  model with optimization
fit <- parameters |>
  initialize_fims(data = fims_frame) |>
  fit_fims(optimize = TRUE)

# Clear memory post-run
clear()
```
