---
title: PIFSC Opakapaka Case Study
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "main Hawaiian Islands Opakapaka"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`  
* Stock name: `r common_name`
* Region: Pacific Islands
* Analyst: Meg Oshima

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* In our model maturity and selectivity are length-based processes so I converted (aka eye-balled logistic curves that look like they match for ages) to values that I think would be roughly equivalent in ages. Eventually, having length-based processes would be necessary for this model.   
* The scale of the fitted CPUE is lower than the observed values.  
* If I don't fix the initial numbers-at-age, the model puts a really large number of individuals in age-0 bin and then 0 fish in the rest of the ages. Not sure what is happening here.  

## Setting up the data

```{r}
#| eval: false
#| label: local-production-model
#| echo: false

# read in data and control files of original model
# Locally this is
# C:/Users/Megumi.Oshima/Documents/Opaka-FIMS-Case-Study/Model/01_original_model
opaka_mod_dir <- file.path(
  getwd(),
  "..",
  "Opaka-FIMS-Case-Study",
  "Model",
  "01_original_model"
)
opaka_dat <- r4ss::SS_readdat_3.30(file.path(opaka_mod_dir, "data.ss"))
opaka_ctl <- r4ss::SS_readctl_3.30(
  file.path(opaka_mod_dir, "control.ss"),
  datlist = file.path(opaka_mod_dir, "data.ss")
)

# create directory for new simplified model
opaka_length_dir <- file.path(opaka_mod_dir, "..", "09_case_study_lengths")
dir.create(opaka_length_dir)

# remove size freq data
opaka_dat$N_sizefreq_methods_rd <- 0
opaka_dat$N_sizefreq_methods <- NULL
opaka_dat$nbins_per_method <- NULL
opaka_dat$units_per_method <- NULL
opaka_dat$scale_per_method <- NULL
opaka_dat$mincomp_per_method <- NULL
opaka_dat$Nobs_per_method <- NULL
opaka_dat$Comp_Error_per_method <- NULL
opaka_dat$ParmSelect_per_method <- NULL
opaka_dat$sizefreq_bins_list <- NULL
opaka_dat$sizefreq_data_list <- NULL

# remove super periods for length comp data
len_dat_original <- read.csv(
  file.path(opaka_mod_dir, "..", "..", "Data", "Opaka_len_data.csv")
)
opaka_dat$lencomp <- len_dat_original

# remove dirichlet weighting for length comps
opaka_dat$len_info$CompError <- 0
opaka_dat$len_info$ParmSelect <- 0

# remove initial F estimation
opaka_dat$catch[1,'catch'] <- 0
# add agecomp dummy data
opaka_dat$N_agebins <- 21
opaka_dat$agebin_vector <- seq(1, 21)
opaka_dat$N_ageerror_definitions <- 1
opaka_dat$ageerror <- rbind(
    seq(0.5, 43.5),
    rep(.01, 44)
)
opaka_dat$age_info <- data.frame(
  mintailcomp <- rep(0, 4),
  addtocomp = 1e-7,
  combine_M_F = 1,
  CompressBins = 0,
  CompError = 0,
  ParmSelect = 0,
  minsamplesize = 1
)
opaka_dat$Lbin_method <- 1
agecomp_info <- data.frame(
  year = rep(seq(2017, 2023), 2),
  month = 1,
  fleet = rep(c(2, 4), each = 7),
  sex = 0,
  part = 0,
  ageerr = 1,
  Lbin_lo = -1,
  Lbin_hi = -1,
  Nsamp = 10
)
dummy_agecomp <- as.data.frame(matrix(
  data = 1,
  nrow = nrow(agecomp_info),
  ncol = length(opaka_dat$agebin_vector)
))
colnames(dummy_agecomp) <- paste0("a", opaka_dat$agebin_vector)
opaka_dat$agecomp <- cbind(agecomp_info, dummy_agecomp)

r4ss::SS_writedat_3.30(
  datlist = opaka_dat,
  outfile = file.path(opaka_length_dir, "data.ss"),
  overwrite = TRUE
)

# remove growth platoon
opaka_ctl$N_platoon <- 1
opaka_ctl$sd_ratio <- NULL
opaka_ctl$submorphdist <- NULL

# remove intial F estimation
opaka_ctl$init_F <- NULL

# remove extra SE parameter
opaka_ctl$Q_options[1, "extra_se"] <- 0
opaka_ctl$Q_parms <- opaka_ctl$Q_parms[-2, ]
opaka_ctl$Variance_adjustment_list <- NULL
opaka_ctl$DoVar_adjust <- 0
opaka_ctl$sd_offset <- 0

# remove dirichlet weighting parameter lines
opaka_ctl$dirichlet_parms <- NULL

# fix commercial selectivity
opaka_ctl$size_selex_parms$PHASE[1:2] <- -2

# add age selectivity
opaka_ctl$age_selex_types <- data.frame(
  Pattern = rep(12, 4),
  Discard = 0,
  Male = 0,
  Special = 0
)

# control file wouldn't write when age_selex_params are manually added
# opaka_ctl$age_selex_parms <-  data.frame(
#   "LO" = c(0, -5, 0, 0, 0, -10, 0, -20),
#   "HI" = c(40, 50, 40, 40, 60, 60, 10, 50),
#   "INIT" = c(1.81975, 0.0093046, 1, 3, 1.97182, 0.00040, 1.29111, 0.00115),
#   "PRIOR" = c(5, 6, 5, 6, 5, 6, 2, .5),
#   "PR_SD" = c(99, 99, 99, 99, 99, 99, 5, 2),
#   "PR_type" = 0,
#   "PHASE" = c(-2, -2, -2, -2, -99, -99, -2, -2),
#   "env-var" = 0,
#   "use_dev" = 0,
#   "dev_mnyr" = 0,
#   "dev_mxyr" = 0,
#   "dev_PH" = 0,
#   "Block" = 0,
#   "Block_Fxn" = 0
# )

age_ctl <- r4ss::SS_readctl_3.30(
  file.path(opaka_mod_dir, "..", "03_age_comps", "control.ss_new"),
  datlist = file.path(opaka_mod_dir, "..", "03_age_comps", "data.ss")
)
age_selex_params <- age_ctl$age_selex_parms
opaka_ctl$age_selex_parms <- age_selex_params
opaka_ctl$age_selex_parms$PHASE <- -2

r4ss::SS_writectl_3.30(
  opaka_ctl,
  outfile = file.path(opaka_length_dir, "control.ss"),
  overwrite = TRUE
)

ss_files <- c("forecast.ss", "starter.ss", "ss_opt_win.exe")
file.copy(file.path(opaka_mod_dir, ss_files), opaka_length_dir)

# create a bootstrap data file to get age comp data
start <- r4ss::SS_readstarter(file.path(opaka_length_dir, "starter.ss"))
start$N_bootstraps <- 3
r4ss::SS_writestarter(start, dir = opaka_length_dir, overwrite = TRUE)

# run SS3
r4ss::run(dir = opaka_length_dir, exe = "ss_opt_win.exe", skipfinished = FALSE)

file.copy(
  file.path(opaka_length_dir, "data_boot_001.ss"),
  file.path(opaka_length_dir, "data.ss"),
  overwrite = TRUE
)
start <- r4ss::SS_readstarter(file.path(opaka_length_dir, "starter.ss"))
start$N_bootstraps <- 1
r4ss::SS_writestarter(start, dir = opaka_length_dir, overwrite = T)

# run SS3
r4ss::run(dir = opaka_length_dir, exe = "ss_opt_win.exe", skipfinished = F)

# check model
rep <- r4ss::SS_output(dir = opaka_length_dir)
SS_plots(rep)

# package up data, control and rep file for using in FIMS
rm("opaka_dat")
rm("opaka_ctl")
opaka_dat <- r4ss::SS_readdat_3.30(file.path(opaka_length_dir, "data.ss"))
opaka_ctl <- r4ss::SS_readctl_3.30(
  file.path(opaka_length_dir, "control.ss"),
  datlist = file.path(opaka_length_dir, "data.ss")
)
save(
  list = c("opaka_dat", "opaka_ctl", "rep"),
  file = file.path(opaka_length_dir, "opaka_length.RDS")
)
```

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false
load(file.path(data_directory, "opaka_model.RDS"))
include_age_comps <- FALSE
years <- seq(opaka_dat$styr, opaka_dat$endyr)
n_years <- length(years) # the number of years which we have data for.
ages <- seq(1, 21) # age vector.
n_ages <- length(ages) # the number of age groups.
comp_lengths <- opaka_dat$lbin_vector # length vector.
nlengths <- length(comp_lengths) # the number of length bins.

opaka_dat_fims <- get_ss3_data(
  list(
    dat = opaka_dat,
    ctl = opaka_ctl,
    start = list(),
    fore = list(),
    wtatage = rep[["wtatage"]]
  ),
  fleets = c(1,2,3),
  ages = ages,
  lengths = comp_lengths
) |>
  dplyr::filter(type != "age_comp")
## age to length conversion matrix 
# Growth function values to create age to length conversion matrix from model
#comparison project
mg_pars <- rep$parameters  |> 
dplyr::filter(stringr::str_detect(Label, "_GP_"))
Linf <- mg_pars$Value[3]
K <- mg_pars$Value[4]
a0 <- -0.29
amax <- 21
cv <- mg_pars$Value[5]

L2Wa <- mg_pars$Value[7]
L2Wb <- mg_pars$Value[8]

AtoL <- function(a,Linf,K,a_0){
  L <- Linf*(1-exp(-K*(a-a_0)))
  }

ages <- 1:amax
len_bins <- comp_lengths

#Create length at age conversion matrix and fill proportions using above 
#growth parameters
length_age_conversion <- matrix(NA,nrow=length(ages),ncol=length(len_bins))
for(i in seq_along(ages)){
  #Calculate mean length at age to spread lengths around
  mean_length <- AtoL(ages[i],Linf,K,a0)
  #mean_length <- AtoLSchnute(ages[i],L1,L2,a1,a2,Ks)
  #Calculate the cumulative proportion shorter than each composition length
  temp_len_probs<-pnorm(q=len_bins,mean=mean_length,sd=mean_length*cv)
  #Reset the first length proportion to zero so the first bin includes all
  #density smaller than that bin
  temp_len_probs[1]<-0
  #subtract the offset length probabilities to calculate the proportion in each
  #bin. For each length bin the proportion is how many fish are larger than this
  #length but shorter than the next bin length.
  temp_len_probs <- c(temp_len_probs[-1],1)-temp_len_probs
  length_age_conversion[i,] <- temp_len_probs
}
colnames(length_age_conversion) <- len_bins
rownames(length_age_conversion) <- ages

#Extract years and fleets from milestone 1 data
start_date <- unique(opaka_dat_fims$timing[opaka_dat_fims$type=="landings"])
observers <- unique(opaka_dat_fims$name[opaka_dat_fims$type=="length_comp"])

#Create data frame for new fleet and year specific length at age conversion proportions
length_age_data <- data.frame(
  type = rep("age-to-length-conversion",length(len_bins)*length(ages)*length(observers)*length(start_date)),
  name = rep(sort(rep(observers,length(len_bins)*length(ages))),length(start_date)),
  age = rep(sort(rep(ages,length(len_bins))),length(observers)*length(start_date)),
  length = rep(len_bins,length(ages)*length(observers)*length(start_date)),
  timing = rep(start_date,each=length(len_bins)*length(ages)*length(observers)),
  value = rep(c(t(length_age_conversion)),length(observers)*length(start_date)),
  unit = rep("proportion",length(len_bins)*length(ages)*length(observers)*length(start_date)),
  uncertainty = rep(30,length(len_bins)*length(ages)*length(observers)*length(start_date)))

# Changing the CPUE indices for fleet1 to be a new fleet, fleet4. This helps with model convergence and fitting as it is the longest time series of data available along with the landings. 
opaka_dat_fims <- opaka_dat_fims |>
  dplyr::mutate(name = ifelse(name == "fleet1" & type == "index", "fleet4", name))

opaka_dat_fims <- type.convert(
  rbind(opaka_dat_fims, length_age_data),
  as.is = TRUE
)

opaka_dat_fims <- opaka_dat_fims |> 
  dplyr::mutate(uncertainty = ifelse(type == "length_comp" & name == "fleet2" & value != -999, 1, uncertainty)) |>
  dplyr::filter(!(type == "index" & name == "fleet3"))

data_4_model <- FIMS::FIMSFrame(opaka_dat_fims)

```

The `data_4_model` object contains a `@data` slot that holds a long data frame with:
* 2 fleets: commercial fishery (fleet1) and survey (fleet2)
* landings for fleet 1 
* cpue for fleet 2
* length composition data for fleet 2


## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model

recdevs <- rep$parameters |>
           dplyr::filter(stringr::str_detect(Label, "RecrDev")) |> 
           dplyr::select(Label, Value)

init_naa <- (exp(opaka_ctl$SR_parms["SR_LN(R0)", "INIT"]) * 1000) * exp(-(ages - 1) * 0.135)
init_naa[n_ages] <- init_naa[n_ages] / 0.135
# Create default parameters
default_parameters <- FIMS::create_default_configurations(
    data = data_4_model
  ) |>
  FIMS::create_default_parameters(
    data = data_4_model
  ) |>
  tidyr::unnest(cols = data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(7, 0.5)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "fleet1",
      label = c("slope", "inflection_point"),
      # Used age selectivity values
      value = c(4.5, 1.81),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "fleet2",
      label = c("slope", "inflection_point"),
      value = c(3, 1),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "fleet3",
      label = c("slope", "inflection_point"),
      value = c(4.5, 1.97),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet_name = "fleet4",
      label = c("slope", "inflection_point"),
      value = c(4.5, 1.81),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_M"),
      value = log(0.135)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = ages,
      value = log(init_naa),
      estimation_type = "constant"
    ),
    by = c("module_name", "label", "age")
  ) |>
  # dplyr::rows_update(
  #   tibble::tibble(
  #     module_name = "Recruitment",
  #     # Transformed 0.999 to logit where a previous version just used 0.999
  #     # Wondering if we should use logit(0.75) as noted previously for scamp
  #     # as the null recruitment model?
  #     label = c("log_rzero", "logit_steep", "log_sd"),
  #     value = c(opaka_ctl$SR_parms["SR_LN(R0)", "INIT"], -log(1.0 - 0.76) + log(0.76 - 0.2), sca$parm.cons$rec_sigma[8])
  #   ),
  #   by = c("module_name", "label")
  # ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_devs",
      time = years[-1],
      # The last value of the initial numbers at age is the first
      # recruitment deviation
      value = recdevs$Value,
    ),
    by = c("module_name", "label", "time")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet1",
      time = years,
      label = "log_Fmort",
      value = log(rep$exploitation$FRS)
    ),
    by = c("module_name", "fleet_name", "label", "time")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet2",
      label = "log_q",
      value = -4.12772
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet3",
      label = "log_Fmort",
      time = years,
      value = log(rep$exploitation$Non_comm)
    ),
    by = c("module_name", "fleet_name", "label", "time")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet4",
      label = "log_q",
      value = -3.90281 #value from SS
    ),
    by = c("module_name", "fleet_name", "label")
  )

# Run the  model with optimization
fit <- default_parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE)
```

## Plotting Results

```{r}
#| label: comparison-plots

index_results <- data.frame(
  observed = FIMS::m_index(data_4_model, "fleet2"),
  expected = FIMS::get_report(fit)[["index_expected"]][[2]]
) |>
dplyr::mutate(year = years) |>
dplyr::filter(year > 2016)
#print(index_results)

ggplot2::ggplot(index_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Index (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected), color = "blue") +
  ggplot2::theme_bw()

cpue_results <- data.frame(
  observed = FIMS::m_index(data_4_model, "fleet4"),
  expected = FIMS::get_report(fit)[["index_expected"]][[2]]
) |>
dplyr::mutate(year = years) |>
dplyr::filter(observed >0)
#print(cpue_results)
ggplot2::ggplot(cpue_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point() +
  ggplot2::xlab("Year") +
  ggplot2::ylab("CPUE") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected), color = "blue") +
  ggplot2::theme_bw()


catch_results <- data.frame(
  observed = c(FIMS::m_landings(data_4_model, fleet = "fleet1"), FIMS::m_landings(data_4_model, fleet = "fleet3")),
  expected = c(FIMS::get_report(fit)[["landings_expected"]][[1]], FIMS::get_report(fit)[["landings_expected"]][[3]]),
  fleet = rep(c("fleet1", "fleet3"), each = 75)
) |>
dplyr::mutate(year = rep(years, 2))
#print(catch_results)

ggplot2::ggplot(catch_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point(ggplot2::aes(color = fleet)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected, color = fleet)) +
  ggplot2::theme_bw()

biomass <- rep$timeseries |> 
dplyr::select(Yr, SpawnBio, Bio_all) |> 
dplyr::filter(Yr > 1947) |> ##CHECK: including "initial year" to match length with FIMS but need to check on FIMS
dplyr::rename("SS_SpawnBio" = "SpawnBio",
              "SS_Bio" = "Bio_all") |>
dplyr::mutate(FIMS_SpawnBio = FIMS::get_report(fit)[["spawning_biomass"]][[1]] , 
              FIMS_Bio = FIMS::get_report(fit)[["biomass"]][[1]]) |> ##CHECK: Is FIMS ssb reporting n_years+1 or initial year-1?
tidyr::pivot_longer(cols = -Yr) |>
tidyr::separate_wider_delim(cols = "name", delim = "_", names = c("Model", "Type"))             

ggplot2::ggplot(biomass, ggplot2::aes(x = Yr, y = value)) +
  ggplot2::geom_line(ggplot2::aes(color = Model)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("") +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw()

recruits <- rep$recruit |>
  dplyr::select(Yr, exp_recr, raw_dev) |>
  dplyr::rename("SS_recruit" = "exp_recr",
                "SS_recdev" = "raw_dev",
                "Year" = "Yr") |> 
  dplyr::mutate(FIMS_recruit = FIMS::get_report(fit)[["expected_recruitment"]][[1]][1:75],
                FIMS_recdev = c(
                  NA,
                  FIMS::get_estimates(fit) |>
                    dplyr::filter(label == "log_devs", module_name == "Recruitment") |>
                    dplyr::pull(estimated)
                ),
                SS_recruit = SS_recruit * 1000) |>
  tidyr::pivot_longer(cols = -Year) |>
  tidyr::separate_wider_delim(cols = "name", delim = "_", names = c("Model", "Type"))             

ggplot2::ggplot(recruits, ggplot2::aes(x = Year, y = value, color = Model)) +
  ggplot2::geom_line() + 
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw()
```


```{r}
#| eval: false
#| label: proportions-plots-not-run
## Checking fit to proportion catch number at length
pcnal <- matrix(data = fit@report$pcnal[[2]], nrow = nlengths)

prop.dat <- opaka_dat_fims |> 
filter(type == "length" & name == "fleet2") |> 
group_by(datestart) |>
reframe(prop = value/sum(value)) 
pcnal.obs <- matrix(data = prop.dat$prop, nrow = nlengths)
head(pcnal.obs)

plot(x = 1:nlengths, y = pcnal.obs[,73], pch = 16,  ylim = c(0,1))
lines(x = 1:nlengths, y = pcnal[,73])
pcnal[,69]

## checking estimated numbers at age
head(fit@report$naa[[1]])
naa_mat <- matrix(data = fit@report$naa[[1]], nrow = n_ages)
head(naa_mat)

```

```{r}
#| label: cleanup
FIMS::clear()
```
