---
title: PIFSC Opakapaka Case Study with Length Data
format:
  html:
    code-fold: true
---

## The setup

```{r}
#| echo: false
R_version <- version$version.string
TMB_version <- packageDescription("TMB")$Version
FIMS_commit <- substr(packageDescription("FIMS")$GithubSHA1, 1, 7)
r4ss_version <- packageVersion('r4ss')
```

* R version = `R_version`
* TMB version = `TMB_version`
* FIMS commit = `FIMS_commit`
* `r4ss` version = `r4ss_version`  
* Stock name: Main Hawaiian Islands Opakapaka
* Region: Pacific Islands
* Analyst: Meg Oshima

## Simplifications made to the model

* Removed size-frequency data
* Removed super-periods for length composition data
* Removed data weighting
* Removed initial depletion estimation
* Removed growth platoons
* Fix selectivity for commercial fishery

## Running FIMS

#### Initial setup 
This step involves reading in the SS3 model information (data, control, and report files), converting the data to FIMS format, and defining the dimensions.
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(FIMS)
library(TMB)
library(r4ss)
clear()
load(file.path(getwd(), "content", "data_files", "opaka_length.RDS"))
## Should now have opaka_dat, opaka_ctl, and rep in environment
## Function written by Ian Taylor and Meg Oshima to get SS3 data into FIMSFrame format
source("./content/R/get_ss3_data.R")

# Define the dimensions ----
### years from 1 to 75
years <- seq(opaka_dat$styr, opaka_dat$endyr)
nyears <- length(years)
nseasons <- 1
ages <- seq(1, 21)
nages <- length(ages)
lengths <- opaka_dat$lbin_vector
nlengths <- length(lengths)

# Data prep ----
## Use R/get_ss3_data function to get from data.ss to FIMSFrame
opaka_dat_fims <- get_ss3_data(opaka_dat, fleets = c(1,2,3,4), ages = ages, lengths = lengths)
## length-at-age conversion matrix 
#Growth function values to create length age conversion matrix from model
#comparison project
mg_pars <- rep$parameters  |> 
dplyr::filter(stringr::str_detect(Label, "_GP_"))
Linf <- mg_pars$Value[3]
K <- mg_pars$Value[4]
a0 <- -0.29
amax <- 21
cv <- mg_pars$Value[5]

L2Wa <- mg_pars$Value[7]
L2Wb <- mg_pars$Value[8]

AtoL <- function(a,Linf,K,a_0){
  L <- Linf*(1-exp(-K*(a-a_0)))
}

ages <- 1:amax
len_bins <- lengths

#Create length at age conversion matrix and fill proportions using above 
#growth parameters
length_age_conversion <- matrix(NA,nrow=length(ages),ncol=length(len_bins))
for(i in seq_along(ages)){
  #Calculate mean length at age to spread lengths around
  mean_length <- AtoL(ages[i],Linf,K,a0)
  #mean_length <- AtoLSchnute(ages[i],L1,L2,a1,a2,Ks)
  #Calculate the cumulative proportion shorter than each composition length
  temp_len_probs<-pnorm(q=len_bins,mean=mean_length,sd=mean_length*cv)
  #Reset the first length proportion to zero so the first bin includes all
  #density smaller than that bin
  temp_len_probs[1]<-0
  #subtract the offset length probabilities to calculate the proportion in each
  #bin. For each length bin the proportion is how many fish are larger than this
  #length but shorter than the next bin length.
  temp_len_probs <- c(temp_len_probs[-1],1)-temp_len_probs
  length_age_conversion[i,] <- temp_len_probs
}
colnames(length_age_conversion) <- len_bins
rownames(length_age_conversion) <- ages

#Extract years and fleets from milestone 1 data
start_date <- unique(opaka_dat_fims$datestart[opaka_dat_fims$type=="landings"])
end_date <- unique(opaka_dat_fims$dateend[opaka_dat_fims$type=="landings"])
observers <- unique(opaka_dat_fims$name[opaka_dat_fims$type=="length-comp"])

#Create data frame for new fleet and year specific length at age conversion proportions
#These are identical across years and fleets in this default example
# length_age_data <- opaka_dat_fims[rep(1,length(ages)*
#                                     length(len_bins)*
#                                     length(start_date)*
#                                     length(observers)),]
# 
# row_index <- 1
# for(k in seq_along(start_date)){
#   for(l in seq_along(observers)){
#     for(i in seq_along(ages)){
#       for(j in seq_along(len_bins)){
#         length_age_data[row_index,] <- data.frame(
#                                         type = "length-at-age",
#                                         name = observers[l],
#                                         age = ages[i],
#                                         length = len_bins[j],
#                                         datestart = start_date[k],
#                                         dateend = end_date[k],
#                                         value = length_age_conversion[i,j],
#                                         unit = "proportion",
#                                         uncertainty = 30) #is this neff? can it vary by year?
#         row_index <- row_index + 1
#       }
#     }
#   }
# }

length_age_data <- data.frame(
  type = rep("age-to-length-conversion",length(len_bins)*length(ages)*length(observers)*length(start_date)),
  name = rep(sort(rep(observers,length(len_bins)*length(ages))),length(start_date)),
  age = rep(sort(rep(ages,length(len_bins))),length(observers)*length(start_date)),
  length = rep(len_bins,length(ages)*length(observers)*length(start_date)),
  datestart = rep(start_date,each=length(len_bins)*length(ages)*length(observers)),
  dateend = rep(end_date,each=length(len_bins)*length(ages)*length(observers)),
  value = rep(c(t(length_age_conversion)),length(observers)*length(start_date)),
  unit = rep("proportion",length(len_bins)*length(ages)*length(observers)*length(start_date)),
  uncertainty = rep(30,length(len_bins)*length(ages)*length(observers)*length(start_date)))

opaka_dat_fims <- rbind(opaka_dat_fims, length_age_data)

```

#### Set up FIMS model  

```{r}
length_frame <- FIMS::FIMSFrame(opaka_dat_fims)

# fishery data ----
fims_landings <- FIMS::m_landings(length_frame)
# commercial fishery
commercial_landings <- methods::new(Index, nyears)
commercial_landings$index_data <- fims_landings[1:nyears]
#commercial_cpue <- methods::new(Index, nyears)
#commercial_cpue$index_data <- FIMS::m_index(length_frame, "fleet1")
# non-commercial fishery
non_comm_landings <- methods::new(Index, nyears)
non_comm_landings$index_data <- fims_landings[(nyears+1):length(fims_landings)]

# survey data ----
# camera survey
camera_index <- methods::new(Index, nyears)
camera_index$index_data <- FIMS::m_index(length_frame, "fleet2")

# camera_agecomp <- methods::new(AgeComp, nyears, nages)
# camera_agecomp$age_comp_data <- FIMS::m_agecomp(length_frame, "fleet2")

camera_lencomp <- methods::new(LengthComp, nyears, nlengths)
camera_lencomp$length_comp_data <- FIMS::m_lengthcomp(length_frame, "fleet2")
camera_length_at_age <- FIMS::m_length_at_age(length_frame, "fleet2")

# fishing survey
resfish_index <- methods::new(Index, nyears)
resfish_index$index_data <- as.numeric(FIMS::m_index(length_frame, "fleet4")) #not sure why it is being converted to character so putting as.numeric for now to get it to correct type

# resfish_agecomp <- methods::new(AgeComp, nyears, nages)
# resfish_agecomp$age_comp_data <- FIMS::m_agecomp(length_frame, "fleet4")

resfish_lencomp <- methods::new(LengthComp, nyears, nlengths)
resfish_lencomp$length_comp_data <- as.numeric(FIMS::m_lengthcomp(length_frame, "fleet4"))
resfish_length_at_age <- FIMS::m_length_at_age(length_frame, "fleet4")

# selectivity ----
selex_ss <- rep$parameters |> filter(stringr::str_detect(Label, "Size"))

commercial_selectivity <- methods::new(LogisticSelectivity)
commercial_selectivity$inflection_point[1]$value <- selex_ss$Value[1]
commercial_selectivity$inflection_point[1]$is_random_effect <- FALSE
commercial_selectivity$inflection_point[1]$estimated <- FALSE
commercial_selectivity$slope[1]$value <- selex_ss$Value[2]
commercial_selectivity$slope[1]$is_random_effect <- FALSE
commercial_selectivity$slope[1]$estimated <- FALSE

non_comm_selectivity <- methods::new(LogisticSelectivity)
non_comm_selectivity$inflection_point[1]$value <- selex_ss$Value[5]
non_comm_selectivity$inflection_point[1]$is_random_effect <- FALSE
non_comm_selectivity$inflection_point[1]$estimated <- FALSE
non_comm_selectivity$slope[1]$value <- selex_ss$Value[6]
non_comm_selectivity$slope[1]$is_random_effect <- FALSE
non_comm_selectivity$slope[1]$estimated <- FALSE

camera_selectivity <- methods::new(LogisticSelectivity)
camera_selectivity$inflection_point[1]$value <- selex_ss$Value[3]
camera_selectivity$inflection_point[1]$is_random_effect <- FALSE
camera_selectivity$inflection_point[1]$estimated <- TRUE
camera_selectivity$slope[1]$value <- selex_ss$Value[4]
camera_selectivity$slope[1]$is_random_effect <- FALSE
camera_selectivity$slope[1]$estimated <- TRUE

resfish_selectivity <- methods::new(DoubleLogisticSelectivity)
resfish_selectivity$inflection_point_asc[1]$value <- selex_ss$Value[7]
resfish_selectivity$inflection_point_asc[1]$is_random_effect <- FALSE  
resfish_selectivity$inflection_point_asc[1]$estimated <- TRUE

resfish_selectivity$inflection_point_desc[1]$value <- 40
resfish_selectivity$inflection_point_desc[1]$is_random_effect <- FALSE
resfish_selectivity$inflection_point_desc[1]$estimated <- TRUE

resfish_selectivity$slope_asc[1]$value <- exp(selex_ss$Value[9]) #not sure if this is correct based on SS3 user manual 
resfish_selectivity$slope_asc[1]$is_random_effect <- FALSE 
resfish_selectivity$slope_asc[1]$estimated <- TRUE

resfish_selectivity$slope_desc[1]$value <- .4 #not sure how to translate ss values 
resfish_selectivity$slope_desc[1]$is_random_effect <- FALSE
resfish_selectivity$slope_desc[1]$estimated <- TRUE

# fleet objects ----

# commercial fleet 
commercial_fleet <- methods::new(Fleet)
# Set nyears and nages
commercial_fleet$nages <- nages
commercial_fleet$nyears <- nyears
# Set values for log_Fmort
commercial_fleet$log_Fmort <- new(ParameterVector, log(rep$exploitation$FRS), nyears) # 
# Turn on estimation for F
commercial_fleet$log_Fmort$set_all_estimable(TRUE)
commercial_fleet$log_q[1]$value <- log(1.0)
commercial_fleet$estimate_q <- FALSE
commercial_fleet$random_q <- FALSE
commercial_fleet$SetObservedIndexData(commercial_landings$get_id())
commercial_fleet$SetSelectivity(commercial_selectivity$get_id())
# non_commercial fleet 
non_commercial_fleet <- methods::new(Fleet)
# Set nyears and nages
non_commercial_fleet$nages <- nages
non_commercial_fleet$nyears <- nyears
# Set values for log_Fmort
non_commercial_fleet$log_Fmort <- new(ParameterVector, log(rep$exploitation$Non_comm), nyears) # 
# Turn on estimation for F
non_commercial_fleet$log_Fmort$set_all_estimable(TRUE)
non_commercial_fleet$log_q[1]$value <- log(1.0)
non_commercial_fleet$estimate_q <- FALSE
non_commercial_fleet$random_q <- FALSE
non_commercial_fleet$SetObservedIndexData(non_comm_landings$get_id())
non_commercial_fleet$SetSelectivity(non_comm_selectivity$get_id())

# survey objects ----
# camera survey 
camera_survey <- methods::new(Fleet)
camera_survey$is_survey <- TRUE
camera_survey$nages <- nages
camera_survey$nyears <- nyears
camera_survey$nlengths <- nlengths
camera_survey$log_q[1]$value <- -4.12772 #rep$parameters  |> filter(stringr::str_detect(Label, "LnQ_base_BFISH(2)")) |> pull(Value)
camera_survey$estimate_q <- TRUE
camera_survey$random_q <- FALSE
camera_survey$SetObservedIndexData(camera_index$get_id())
camera_survey$SetSelectivity(camera_selectivity$get_id())
camera_survey$SetObservedLengthCompData(camera_lencomp$get_id())

# research fishing survey 
resfish_survey <- methods::new(Fleet)
resfish_survey$is_survey <- TRUE
resfish_survey$nages <- nages
resfish_survey$nyears <- nyears
resfish_survey$nlengths <- nlengths
resfish_survey$log_q[1]$value <- -5.67861 #rep$parameters  |> filter(stringr::str_detect(Label, "LnQ_base_BFISH_ResFish")) |> pull(Value)
resfish_survey$estimate_q <- TRUE
resfish_survey$random_q <- FALSE
resfish_survey$SetObservedIndexData(resfish_index$get_id())
resfish_survey$SetSelectivity(resfish_selectivity$get_id())
resfish_survey$SetObservedLengthCompData(resfish_lencomp$get_id())

# distributions ---- 
# index
commercial_fleet_index_distribution <- new_data_distribution(
  module = commercial_fleet,
  family = lognormal(link = "log"),
  sd = list(
    value = rep(sqrt(log(0.01^2 + 1)), nyears),
    estimated = FALSE
  ),
  data_type = "index"
)

non_commercial_fleet_index_distribution <- new_data_distribution(
  module = non_commercial_fleet,
  family = lognormal(link = "log"),
  sd = list(
    value = rep(sqrt(log(0.01^2 + 1)), nyears),
    estimated = FALSE
  ),
  data_type = "index"
)

camera_survey_index_distribution <- new_data_distribution(
  module = camera_survey,
  family = lognormal(link = "log"),
  sd = list(
    value = rep(sqrt(log(0.2^2 + 1)), nyears),
    estimated = FALSE
  ),
  data_type = "index"
)

resfish_survey_index_distribution <- new_data_distribution(
  module = resfish_survey,
  family = lognormal(link = "log"),
  sd = list(
    value = rep(sqrt(log(0.2^2 + 1)), nyears),
    estimated = FALSE
  ),
  data_type = "index"
)

# length comps
camera_survey_lengthcomp_distribution <- new_data_distribution(
  module = camera_survey,
  family = multinomial(link = "logit"),
  data_type = "lengthcomp"
)

resfish_survey_lengthcomp_distribution <- new_data_distribution(
  module = resfish_survey,
  family = multinomial(link = "logit"),
  data_type = "lengthcomp"
)
#Set length-at-age conversion matrix
camera_survey$age_length_conversion_matrix <- new(ParameterVector,as.numeric(camera_length_at_age),nages*nlengths)
resfish_survey$age_length_conversion_matrix <- new(ParameterVector,as.numeric(resfish_length_at_age),nages*nlengths)
# Turn off estimation for length-at-age
camera_survey$age_length_conversion_matrix$set_all_estimable(FALSE)
camera_survey$age_length_conversion_matrix$set_all_random(FALSE)
resfish_survey$age_length_conversion_matrix$set_all_estimable(FALSE)
resfish_survey$age_length_conversion_matrix$set_all_random(FALSE)

# population ---- 
#Recruitment
recruitment <- methods::new(BevertonHoltRecruitment)
#There are two parameters we need to set-up: *log_rzero*, and *logit_steep*.
recruitment$log_rzero[1]$value <- opaka_ctl$SR_parms["SR_LN(R0)", "INIT"]
recruitment$log_rzero[1]$is_random_effect <- FALSE
recruitment$log_rzero[1]$estimated <- TRUE
recruitment$logit_steep[1]$value <- -log(1.0 - 0.76) + log(0.76 - 0.2)
recruitment$logit_steep[1]$is_random_effect <- FALSE
recruitment$logit_steep[1]$estimated <- FALSE
recdevs <- rep$parameters |> filter(stringr::str_detect(Label, "RecrDev")) |> select(Label, Value)
recruitment$log_devs <- new(ParameterVector, c(0, as.numeric(recdevs$Value)), nyears)
#distribution to estimate log_devs
recruitment_distribution <- new_process_distribution(
  module = recruitment,
  par = "log_devs",
  family = gaussian(),
  sd = list(value = 0.52, estimated = FALSE),
  is_random_effect = FALSE
)

# growth 
ewaa_growth <- methods::new(EWAAgrowth)
ewaa_growth$ages <- ages
weights <- rep$wtatage |> 
dplyr::filter(sex == 1 & fleet == 1 & year == 1949) |> 
dplyr::select(paste(1:21)) |> 
round(4)
ewaa_growth$weights <- unlist(weights)

# maturity
maturity <- new(LogisticMaturity)
# approximate age-based equivalent to length-based maturity in petrale model
# based on looking at ss3rep$endgrowth |> dplyr::filter(Sex == 1) |> dplyr::select(Age_Beg, Len_Mat)
maturity$inflection_point[1]$value <- 3.5
maturity$inflection_point[1]$is_random_effect <- FALSE
maturity$inflection_point[1]$estimated <- FALSE
maturity$slope[1]$value <- 2 # arbitrary guess, not sure how to relate this to SS value of -2.26
maturity$slope[1]$is_random_effect <- FALSE
maturity$slope[1]$estimated <- FALSE

population <- new(Population)

M_value <- opaka_ctl$MG_parms["NatM_p_1_Fem_GP_1", "INIT"]
population$log_M <- new(ParameterVector, rep(log(M_value), nyears * nages), nyears * nages)
population$log_M$set_all_estimable(FALSE)
# initial numbers at age based on R0 + mortality
init_naa <- exp(recruitment$log_rzero[1]$value) * exp(-(ages - 1) * M_value)
init_naa[nages] <- init_naa[nages] / M_value # sum of infinite series
population$log_init_naa <- new(ParameterVector, log(init_naa), nages)
population$log_init_naa$set_all_estimable(TRUE)
population$nages <- nages
population$ages <- ages
population$nfleets <- 4 # fleet plus surveys
population$nseasons <- 1
population$nyears <- nyears

population$SetMaturity(maturity$get_id())
population$SetGrowth(ewaa_growth$get_id())
population$SetRecruitment(recruitment$get_id())

# create FIMS model
success <- CreateTMBModel()
parameters <- list(p = get_fixed())
obj <- MakeADFun(data = list(), parameters, DLL = "FIMS", silent = TRUE)
opt <- nlminb(obj$par, obj$fn, obj$gr, control = list(eval.max = 10000, iter.max = 10000))
#print(opt)
report <- obj$report()

```

## Add your comparison figures

## Add comparison tables

## What was your experience using FIMS? What could we do to improve usability?

## List any issues that you ran into or found

Please [open an issue](https://github.com/NOAA-FIMS/FIMS/issues/new/choose) if you found something new.

## What features are most important to add based on this case study?

## Add your pages to the project

* Add the files to `_quarto.yml`
