---
title: PIFSC ʻŌpakapaka Case Study
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "main Hawaiian Islands Opakapaka"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`  
* Stock name: `r common_name`
* Region: Pacific Islands
* Analyst: Meg Oshima

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* In our model maturity and selectivity are length-based processes so I converted (aka eye-balled logistic curves that look like they match for ages) to values that I think would be roughly equivalent in ages. Eventually, having length-based processes would be necessary for this model.   
* The scale of the fitted CPUE is lower than the observed values.  
* If I don't fix the initial numbers-at-age, the model puts a really large number of individuals in age-0 bin and then 0 fish in the rest of the ages. Not sure what is happening here.  

## Setting up the data

```{r}
#| eval: false
#| label: local-production-model
#| echo: false
# read in data and control files of original model
# Locally this is
# C:/Users/Megumi.Oshima/Documents/Opaka-FIMS-Case-Study/Model/01_original_model

# code to modify original model and create bootstrap data
# this creates the .RDS file which is included with the case studies but is not run when rendering the case study. The code is included here for transparency and reproducibility but is not intended to be run by users of the case study.
if (FALSE) {
  # local directory to write modified data and control files to and run SS3
  local_dir <- tempdir()

  opaka_inputs_original <- r4ss::SS_read(
    "https://raw.githubusercontent.com/MOshima-PIFSC/Opaka-FIMS-Case-Study/refs/heads/main/Model/01_original_model/",
    ss_new = TRUE
  )

  opaka_dat <- opaka_inputs_original$dat
  opaka_ctl <- opaka_inputs_original$ctl

  # create directory for new simplified model
  opaka_length_dir <- file.path(local_dir, "09_case_study_lengths")
  dir.create(opaka_length_dir)

  # changes to data file

  # remove size freq data
  opaka_dat$N_sizefreq_methods_rd <- 0
  opaka_dat$N_sizefreq_methods <- NULL
  opaka_dat$nbins_per_method <- NULL
  opaka_dat$units_per_method <- NULL
  opaka_dat$scale_per_method <- NULL
  opaka_dat$mincomp_per_method <- NULL
  opaka_dat$Nobs_per_method <- NULL
  opaka_dat$Comp_Error_per_method <- NULL
  opaka_dat$ParmSelect_per_method <- NULL
  opaka_dat$sizefreq_bins_list <- NULL
  opaka_dat$sizefreq_data_list <- NULL

  # remove super periods for length comp data
  len_dat_original <- read.csv(
    "https://raw.githubusercontent.com/MOshima-PIFSC/Opaka-FIMS-Case-Study/refs/heads/main/Data/Opaka_len_data.csv"
  )
  opaka_dat$lencomp <- len_dat_original

  # remove dirichlet weighting for length comps
  opaka_dat$len_info$CompError <- 0
  opaka_dat$len_info$ParmSelect <- 0

  # remove initial F estimation
  opaka_dat$catch[1, "catch"] <- 0
  # add agecomp dummy data
  opaka_dat$N_agebins <- 21
  opaka_dat$agebin_vector <- seq(1, 21)
  opaka_dat$N_ageerror_definitions <- 1
  opaka_dat$ageerror <- rbind(
    seq(0.5, 43.5),
    rep(.01, 44)
  )
  opaka_dat$age_info <- data.frame(
    mintailcomp = rep(0, 4),
    addtocomp = 1e-7,
    combine_M_F = 1,
    CompressBins = 0,
    CompError = 0,
    ParmSelect = 0,
    minsamplesize = 1
  )
  opaka_dat$Lbin_method <- 1
  agecomp_info <- data.frame(
    year = rep(seq(2017, 2023), 2),
    month = 1,
    fleet = rep(c(2, 4), each = 7),
    sex = 0,
    part = 0,
    ageerr = 1,
    Lbin_lo = -1,
    Lbin_hi = -1,
    Nsamp = 10
  )
  dummy_agecomp <- as.data.frame(matrix(
    data = 1,
    nrow = nrow(agecomp_info),
    ncol = length(opaka_dat$agebin_vector)
  ))
  colnames(dummy_agecomp) <- paste0("a", opaka_dat$agebin_vector)
  opaka_dat$agecomp <- cbind(agecomp_info, dummy_agecomp)

  # changes to control file

  # remove growth platoon
  opaka_ctl$N_platoon <- 1
  opaka_ctl$sd_ratio <- NULL
  opaka_ctl$submorphdist <- NULL

  # remove initial F estimation
  opaka_ctl$init_F <- NULL

  # remove extra SE parameter
  opaka_ctl$Q_options[1, "extra_se"] <- 0
  opaka_ctl$Q_parms <- opaka_ctl$Q_parms[-2, ]
  opaka_ctl$Variance_adjustment_list <- NULL
  opaka_ctl$DoVar_adjust <- 0
  opaka_ctl$sd_offset <- 0

  # remove dirichlet weighting parameter lines
  opaka_ctl$dirichlet_parms <- NULL

  # fix commercial selectivity
  opaka_ctl$size_selex_parms$PHASE[1:2] <- -2

  # add age selectivity
  opaka_ctl$age_selex_types <- data.frame(
    Pattern = rep(12, 4),
    Discard = 0,
    Male = 0,
    Special = 0
  )

  # control file wouldn't write when age_selex_params are manually added
  # opaka_ctl$age_selex_parms <-  data.frame(
  #   "LO" = c(0, -5, 0, 0, 0, -10, 0, -20),
  #   "HI" = c(40, 50, 40, 40, 60, 60, 10, 50),
  #   "INIT" = c(1.81975, 0.0093046, 1, 3, 1.97182, 0.00040, 1.29111, 0.00115),
  #   "PRIOR" = c(5, 6, 5, 6, 5, 6, 2, .5),
  #   "PR_SD" = c(99, 99, 99, 99, 99, 99, 5, 2),
  #   "PR_type" = 0,
  #   "PHASE" = c(-2, -2, -2, -2, -99, -99, -2, -2),
  #   "env-var" = 0,
  #   "use_dev" = 0,
  #   "dev_mnyr" = 0,
  #   "dev_mxyr" = 0,
  #   "dev_PH" = 0,
  #   "Block" = 0,
  #   "Block_Fxn" = 0
  # )

  age_ctl <- r4ss::SS_read(
    "https://raw.githubusercontent.com/MOshima-PIFSC/Opaka-FIMS-Case-Study/refs/heads/main/Model/03_age_comps/",
    ss_new = TRUE
  )$ctl

  age_selex_params <- age_ctl$age_selex_parms
  opaka_ctl$age_selex_parms <- age_selex_params
  opaka_ctl$age_selex_parms$PHASE <- -2

  # gather modified elements into new list of inputs
  opaka_inputs_modified <- opaka_inputs_original
  opaka_inputs_modified$dat <- opaka_dat
  opaka_inputs_modified$ctl <- opaka_ctl
  # create a bootstrap data file to get age comp data
  opaka_inputs_modified$start$N_bootstraps <- 3

  # write modified input files to local directory to run SS3
  r4ss::SS_write(
    opaka_inputs_modified,
    dir = opaka_length_dir,
    overwrite = TRUE
  )


  ## specify SS3 exe name and get exe from repo:
  exe <- "ss3"
  r4ss::get_ss3_exe(dir = opaka_length_dir, version = "v3.30.21")
  ## run SS3
  r4ss::run(dir = opaka_length_dir, exe = exe, skipfinished = FALSE)

  file.copy(
    file.path(opaka_length_dir, "data_boot_001.ss"),
    file.path(opaka_length_dir, "data.ss"),
    overwrite = TRUE
  )

  # run SS3 using bootstrap data
  r4ss::run(dir = opaka_length_dir, exe = exe, skipfinished = FALSE)

  # check model
  rep <- r4ss::SS_output(dir = opaka_length_dir)
  r4ss::SS_plots(rep)

  # package up data, control and rep file for using in FIMS
  rm("opaka_dat")
  rm("opaka_ctl")

  # read input files again (to get the bootstrap data and the wtatage.ss_new)
  opaka_inputs_bootstrap <- r4ss::SS_read(opaka_length_dir, read_wtatage = TRUE)
  # save to model directory
  save(
    list = c("opaka_inputs_bootstrap", "rep"),
    file = file.path(opaka_length_dir, "opaka_model.RDS")
  )
  # copy to data directory for case studies repository
  file.copy(
    file.path(opaka_length_dir, "opaka_model.RDS"),
    file.path(data_directory, "opaka_model.RDS"),
    overwrite = TRUE
  )
} # end if (FALSE) for code used to prepare .RDS file
```

```{r}
#| label: prepare-fims-data
#| output: false
#| warning: false
# load opaka_inputs_bootstrap and rep objects with model input and output
load(file.path(data_directory, "opaka_model.RDS"))
include_age_comps <- FALSE
years <- seq(opaka_inputs_bootstrap$dat$styr, opaka_inputs_bootstrap$dat$endyr)
alk_years <- c(years, max(years) + 1)
n_years <- length(years) # the number of years which we have data for.
ages <- seq(1, 21) # age vector.
n_ages <- length(ages) # the number of age groups.
comp_lengths <- opaka_inputs_bootstrap$dat$lbin_vector # length vector.
nlengths <- length(comp_lengths) # the number of length bins.

opaka_dat_fims_raw <- r4ss::ss3_data_to_fims(
  ss3_inputs = opaka_inputs_bootstrap,
  ss3_output = rep,
  fleets = c(1, 2, 3),
  maxage = max(ages),
  lengths = comp_lengths
) |>
  dplyr::filter(type != "age_comp")

opaka_age_to_length <- opaka_dat_fims_raw |>
  dplyr::filter(type == "age_to_length_conversion") |>
  dplyr::select(-timing) |>
  dplyr::mutate(
    fleet = "BFISH"
  ) |>
  tidyr::crossing(timing = alk_years)

opaka_dat_fims <- opaka_dat_fims_raw |>
  dplyr::filter(type != "age_to_length_conversion") |>
  dplyr::bind_rows(opaka_age_to_length) |>
  dplyr::mutate(
    uncertainty = dplyr::if_else(
      type == "length_comp" & fleet == "BFISH",
      "~dmultinom(prob = lengthcomp_proportion, size = 1)",
      uncertainty
    )
  )

data_4_model <- FIMS::FIMSFrame(opaka_dat_fims)


```

The `data_4_model` object contains a `@data` slot that holds a long data frame with:

* 4 fleets: commercial fishery (FRS) and non-commercial fishery (Non_comm), survey (BFISH), and a new fleet for the CPUE of the commercial fishery (FRS_CPUE)
* landings for FRS and Non_comm
* indices for BFISH and FRS_CPUE
* length composition data for the survey (BFISH)
* weight-at-age data
* age-to-length-conversion data


## Run FIMS model

```{r, max.height='100px', attr.output='.numberLines'}
#| label: setup-model
recdevs <- rep$parameters |>
  dplyr::filter(stringr::str_detect(Label, "RecrDev")) |>
  dplyr::select(Label, Value)

init_naa <- (exp(opaka_inputs_bootstrap$ctl$SR_parms["SR_LN(R0)", "INIT"])) * exp(-(0:max(ages)) * 0.135)
init_naa[n_ages] <- init_naa[n_ages] / 0.135

# Create default parameters
default_parameters <- FIMS::setup_default_parameters(
    data = data_4_model,
    growth_module_type = "VonBertalanffy"
  )

# modify the default parameters
parameters <- default_parameters |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(3.5, 3)
    ),
    by = c("module_name", "label")
  ) |>
  # update the growth parameters with initial values from SS3
  # SS3 model uses reference age 2 = 999 which is a switch which makes the 
  # parameter (fixed at 67.5) apply to L-infinity
  # the 67.1182 value in the SS3 estimate of length at age 21
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Growth",
      label = c(
        "length_at_ref_age_1",
        "length_at_ref_age_2",
        "growth_coefficient_K",
        "reference_age_for_length_1",
        "reference_age_for_length_2",
        "length_weight_a", # added length-weight relationship parameters
        "length_weight_b"
      ),
      value = c(6, 67.1182, .242, 0, 21, 1.75e-05, 2.99),
      estimation_type = "constant"
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( #Note: this could probably be combined with Growth block above and put NAs for age for the other parameters
    tibble::tibble(
      module_name = "Growth",
      label = "length_at_age_sd_at_ref_ages",
      age = c(0, 21),
      value = c(0.51, 5.71)
    ),
    by=c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "FRS",
      label = c("slope", "inflection_point"),
      # Used age selectivity values
      value = c(4.5, 1.81),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "BFISH",
      label = c("slope", "inflection_point"),
      value = c(3, 1),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      fleet = "Non_comm",
      label = c("slope", "inflection_point"),
      value = c(4.5, 1.97),
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_M"),
      value = log(0.135)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = "log_init_naa",
      age = 0:max(ages), #fixed init_naa to be from age 0 - max age instead of age 1
      value = log(init_naa),
      estimation_type = "constant"
    ),
    by = c("module_name", "label", "age")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      # Transformed 0.999 to logit where a previous version just used 0.999
      # Wondering if we should use logit(0.75) as noted previously for scamp
      # as the null recruitment model?
      label = c("log_rzero", "logit_steep", "log_sd"),
      value = c(
        # log of R0
        log(1000) + # adding log(1000) to account for R0 units in 1000s in SS3
          opaka_inputs_bootstrap$ctl$SR_parms["SR_LN(R0)", "INIT"],
        # steepness (with logit transformation)
        -log(1.0 - opaka_inputs_bootstrap$ctl$SR_parms["SR_BH_steep", "INIT"]) +
          log(opaka_inputs_bootstrap$ctl$SR_parms["SR_BH_steep", "INIT"] - 0.2),
        # sigmaR
        opaka_inputs_bootstrap$ctl$SR_parms["SR_sigmaR", "INIT"]
      ),
      estimation_type = c("constant", "constant", "constant") #fixing for convergence
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_devs",
      timing = years[-1],
      # The last value of the initial numbers at age is the first
      # recruitment deviation
      value = recdevs$Value
    ),
    by = c("module_name", "label", "timing")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet = "FRS",
      timing = years,
      label = "log_Fmort",
      value = log(rep$exploitation$FRS[-76]) # removing F value for forecast year
    ),
    by = c("module_name", "fleet", "label", "timing")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet = "BFISH",
      label = "log_q",
      value = rep$parameters["LnQ_base_BFISH(2)", "Value"]
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet = "Non_comm",
      label = "log_Fmort",
      timing = years,
      value = log(rep$exploitation$Non_comm[-76]) # removing F value for forecast year
    ),
    by = c("module_name", "fleet", "label", "timing")
  )

# Run the model with optimization
fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  # Model is too big to run on GitHub action if you estimate uncertainty
  FIMS::fit_fims(optimize = TRUE, get_sd = FALSE, number_of_loops = 1)
```

## Examining convergence and parameter estimates

```{r}
obj_fit <- methods::slot(fit,"obj")
opt_par_fit <- FIMS::get_opt(fit)$par
grad_fit <- obj_fit$gr(opt_par_fit)

# Create tibble of parameter estimates
gradient_fit <- tibble::tibble(
      parameter = names(obj_fit$par),
      value = as.numeric(opt_par_fit),
      gradient = as.numeric(grad_fit),
      abs_gradient = abs(gradient)
    ) |>
      dplyr::arrange(parameter)

gradient_fit |>
      dplyr::slice_head(n = 30) |>
      print(width = 200)
gradient_fit |>
      dplyr::slice_tail(n = 30) |>
      print(width = 200)


```

## Plotting Results

```{r}

#
output <- FIMS::get_estimates(fit) |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + FIMS::get_start_year(data_4_model) - 1,
    estimate = estimated,
    fleet = dplyr::recode_values(
      module_id,
      1 ~ "BFISH",
      2 ~ "FRS",
      3 ~ "Non_comm",
      default = NA_character_
    )
  )

fleet_module_lookup <- parameters |>
  dplyr::filter(module_name == "Fleet") |>
  dplyr::distinct(fleet)

index_module_lookup <- fleet_module_lookup |>
  dplyr::filter(fleet %in% c("BFISH", "FRS"))

# year conversion matrix and index plots copied from pacific-hake.qmd
year_conversion_matrix <- dplyr::tibble(
  year = FIMS::get_start_year(data_4_model):(FIMS::get_end_year(data_4_model) + 1)
) |>
  dplyr::mutate(
    index = dplyr::row_number()
  )

# index
stockplotr::plot_timeseries(
  stockplotr::filter_data(
    output |>
      dplyr::filter(fleet == "BFISH") |>
      dplyr::filter(year > 2016), #there are estimated values for years before observed data so removing them
    label_name = "^index_expected$",
    geom = "line"
  ),
  x = "year",
  y = "estimated",
  ylab = "Relative Index of Abundance for BFISH"
) +
  ggplot2::geom_point(
    data = FIMS::get_estimates(fit) |>
      dplyr::mutate(
        fleet = dplyr::recode_values(
          module_id,
          1 ~ "BFISH",
          2 ~ "FRS",
          3 ~ "Non_comm",
          default = NA_character_
        ) 
      )|>
      dplyr::filter(
        fleet == "BFISH",
        label == "index_expected",
        observed > -999
      ) |>
      dplyr::left_join(
        year_conversion_matrix,
        by = c("year_i" = "index")
      ),
    ggplot2::aes(x = year, y = observed)
  ) +
  stockplotr::theme_noaa()

stockplotr::plot_timeseries(
  stockplotr::filter_data(
    output |>
      dplyr::filter(fleet == "FRS"),
    label_name = "^index_expected$",
    geom = "line"
  ),
  x = "year",
  y = "estimated",
  ylab = "Relative CPUE for FRS"
) +
  ggplot2::geom_point(
    data = FIMS::get_estimates(fit) |>
      dplyr::mutate(
        fleet = dplyr::recode_values(
          module_id,
          1 ~ "BFISH",
          2 ~ "FRS",
          3 ~ "Non_comm",
          default = NA_character_
        ) 
      )|>
      dplyr::filter(
        fleet == "FRS",
        label == "index_expected",
        observed > -999
      ) |>
      dplyr::left_join(
        year_conversion_matrix,
        by = c("year_i" = "index")
      ),
    ggplot2::aes(x = year, y = observed)
  ) +
  stockplotr::theme_noaa()


landings_years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(data_4_model)

observed_landings <- data_4_model@data |> 
  dplyr::filter(type == "catch" & fleet != "BFISH") |> 
  dplyr::select(fleet, timing, observed) |> 
  dplyr::rename(fleet = "fleet", year = "timing")

expected_landings <- output |>
  dplyr::filter(label == "landings_expected") |>
  dplyr::mutate(
        fleet = dplyr::recode_values(
          module_id,
          1 ~ "BFISH",
          2 ~ "FRS",
          3 ~ "Non_comm",
          default = NA_character_
        ) 
      )|>
  dplyr::filter(fleet %in% c("FRS", "Non_comm")) |>
  dplyr::transmute(
    fleet = fleet,
    year,
    expected = estimate
  )

catch_results <- observed_landings |>
  dplyr::left_join(expected_landings, by = c("fleet", "year")) |>
  dplyr::filter(observed > -999)

# print(catch_results)

ggplot2::ggplot(catch_results, ggplot2::aes(x = year, y = observed)) +
  ggplot2::geom_point(ggplot2::aes(color = fleet)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Catch (mt)") +
  ggplot2::geom_line(ggplot2::aes(x = year, y = expected, color = fleet)) +
  ggplot2::theme_bw()

# SS3 timeseries includes one extra initial biomass year before the first
# modeled year. In this case that initial comparison year is 1948, so keep
# `Yr > 1947` to align the SS3 biomass series with the FIMS comparison.
ss3_biomass <- rep$timeseries |>
  dplyr::select(Yr, SpawnBio, Bio_all) |>
  dplyr::filter(Yr > 1947) |>
  dplyr::rename(
    year = Yr,
    Bio = Bio_all
  ) |>
  tidyr::pivot_longer(
    cols = c(SpawnBio, Bio),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "SS3")

# FIMS biomass and spawning biomass include one extra initial-year value.
# That first value corresponds to start_year - 1, so plot FIMS on
# (start_year - 1):end_year rather than start_year:end_year.
fims_biomass <- tibble::tibble(
  year = (FIMS::get_start_year(data_4_model) - 1):FIMS::get_end_year(data_4_model),
  SpawnBio = FIMS::get_report(fit)[["spawning_biomass"]][[1]],
  Bio = FIMS::get_report(fit)[["biomass"]][[1]]
) |>
  tidyr::pivot_longer(
    cols = c(SpawnBio, Bio),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "FIMS")

biomass <- dplyr::bind_rows(ss3_biomass, fims_biomass)

ggplot2::ggplot(biomass, ggplot2::aes(x = year, y = value)) +
  ggplot2::geom_line(ggplot2::aes(color = Model)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("") +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::ylim(0, NA)

# SS3 recruitment output is already indexed by report year in `Yr`, so use
# those years directly for the recruitment comparison plot.
ss3_recruits <- rep$recruit |>
  dplyr::select(Yr, pred_recr, raw_dev) |>
  dplyr::rename(
    Year = Yr,
    recruit = pred_recr,
    recdev = raw_dev
  ) |>
  dplyr::mutate(
    recruit = recruit
  ) |>
  tidyr::pivot_longer(
    cols = c(recruit, recdev),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "SS3")

# FIMS expected recruitment is reported on modeled years start_year:end_year.
# Recruitment deviations are not defined for the first modeled year, so pad
# that first year with NA before plotting.
fims_recruit_years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(data_4_model)

fims_recruits <- tibble::tibble(
  Year = fims_recruit_years,
  recruit = FIMS::get_report(fit)[["expected_recruitment"]][[1]][seq_along(fims_recruit_years)],
  recdev = c(
    NA,
    FIMS::get_estimates(fit) |>
      dplyr::filter(label == "log_devs", module_name == "Recruitment") |>
      dplyr::mutate(estimated = exp(estimated)) |>
      dplyr::pull(estimated)
  )
) |>
  tidyr::pivot_longer(
    cols = c(recruit, recdev),
    names_to = "Type",
    values_to = "value"
  ) |>
  dplyr::mutate(Model = "FIMS") |>
  dplyr::filter(!is.na(value))

# Drop undefined recruitment deviations before plotting time series.
recruits <- dplyr::bind_rows(ss3_recruits, fims_recruits) |>
  dplyr::filter(!is.na(value))

ggplot2::ggplot(recruits, ggplot2::aes(x = Year, y = value, color = Model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~Type, scales = "free_y") +
  ggplot2::ggtitle("Recruitment and Recruitment Deviations") +
  ggplot2::theme_bw()
```


```{r}
#| eval: true
#| label: length-composition-plots
## Checking fit to length compositions

lengthcomp <- rbind(
  # length comps from FIMS output
  output |>
    dplyr::filter(label == "lengthcomp_expected", observed != -999) |>
    dplyr::mutate(
      length = comp_lengths[length_i]
    ) |>
    dplyr::select(year, length, estimated, observed) |>
    dplyr::mutate(platform = "FIMS"),

  # get length comps from SS3 output for comparison
  rep$lendbase |>
    dplyr::filter(Fleet == which(rep$FleetNames == "BFISH")) |>
    dplyr::select(year = Yr, length = Bin, estimated = Exp, observed = Obs) |>
    dplyr::mutate(platform = "SS3")
)

lengthcomp |> ggplot2::ggplot(ggplot2::aes(length, estimated, color = platform)) +
  ggplot2::facet_wrap("year") +
  ggplot2::geom_line() +
  ggplot2::geom_point(ggplot2::aes(y = observed, color = "data")) +
  ggplot2::ggtitle("Fit to Length Composition Data") +
  ggplot2::theme_bw()
```

```{r}
#| eval: true
#| label: growth-curve-plots

## Plot Growth Curve

# Pull the report and estimates from the fitted model.
report_fit <- FIMS::get_report(fit)
est_fit <- FIMS::get_estimates(fit)

# Pull model ages.
ages_fit <- FIMS::get_ages(data_4_model)

# Pull mean length-at-age from the report.
growth_laa_values <- unlist(report_fit[["growth_mean_LAA"]])

# Convert the report vector into an age-by-year matrix.
n_ages_fit <- length(ages_fit)
n_growth_years_fit <- length(growth_laa_values) / n_ages_fit

growth_laa_matrix <- matrix(
  growth_laa_values,
  nrow = n_ages_fit,
  ncol = n_growth_years_fit
)

# Put the Growth curve into a plotting table.
growth_curve_fit <- tibble::tibble(
  age = rep(ages_fit, times = n_growth_years_fit),
  year_i = rep(seq_len(n_growth_years_fit), each = n_ages_fit),
  mean_length = as.vector(growth_laa_matrix)
) |>
  dplyr::group_by(age) |>
  dplyr::summarise(
    mean_length = mean(mean_length, na.rm = TRUE),
    min_length = min(mean_length, na.rm = TRUE),
    max_length = max(mean_length, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(platform = "FIMS")

# SS3 mean length at age for comparison
growth_curve_ss3 <- rep$endgrowth |> 
  dplyr::select(Age_Beg, Len_Beg, SD_Beg) |> # beginning of year values, could also use "Mid" instead of "Beg"
  dplyr::rename(age = Age_Beg, mean_length = Len_Beg, sd_length = SD_Beg) |> 
  dplyr::mutate(min_length = NA, max_length = NA, platform = "SS3")

# combine FIMS and SS3 growth curves for plotting
growth_curve_both <- dplyr::bind_rows(growth_curve_fit, growth_curve_ss3)

# Plot mean length-at-age for both models.
ggplot2::ggplot(
  growth_curve_both,
  ggplot2::aes(x = age, y = mean_length, color = platform, fill = platform, linetype = platform)
) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::geom_point(size = 2) +
  ggplot2::labs(
    title = "fit Growth curve",
    subtitle = "Mean length-at-age from FIMS vs SS3",
    x = "Age",
    y = "Mean length-at-age"
  ) +
  ggplot2::theme_bw()
```

```{r}
#| eval: false
#| label: maturity-plots
## Plot Maturity Curve

# Pull proportion mature at age from the FIMS report.
prop_maa_values <- unlist(report_fit[["proportion_mature_at_age"]])

# SS3 maturity values
# the SS3 model has length-based maturity so the derived
# maturity-at-age is shown in the table of ending year growth
# in the Len_Mat column. If it were age-based, we would 
# use the Age_Mat column instead.
maturity_curve_ss3 <- rep$endgrowth |> 
  dplyr::select(Age_Beg, Len_Mat) |> 
  dplyr::filter(Age_Beg <= max(ages_fit)) |> 
  dplyr::rename(age = Age_Beg, mean_mature = Len_Mat) |> 
  dplyr::mutate(platform = "SS3")

# Convert the report vector into an age-by-year matrix.
n_ages_fit <- length(ages_fit)
n_mat_years_fit <- length(prop_maa_values) / n_ages_fit

prop_maa_matrix <- matrix(
  prop_maa_values,
  nrow = n_ages_fit,
  ncol = n_mat_years_fit
)

# Put the maturity curve into a plotting table.
maturity_curve_fit <- tibble::tibble(
  age = rep(ages_fit, times = n_mat_years_fit),
  year_i = rep(seq_len(n_mat_years_fit), each = n_ages_fit),
  mean_mature = as.vector(prop_maa_matrix)
) |>
  dplyr::group_by(age) |>
  dplyr::summarise(
    mean_mature = mean(mean_mature, na.rm = TRUE),
    min_mature = min(mean_mature, na.rm = TRUE),
    max_mature = max(mean_mature, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::mutate(platform = "FIMS")

  # combine FIMS and SS3 growth curves for plotting
maturity_curve_both <- dplyr::bind_rows(maturity_curve_fit, maturity_curve_ss3)

# Plot mean maturity-at-age for both models.
ggplot2::ggplot(
  maturity_curve_both,
  ggplot2::aes(x = age, y = mean_mature, color = platform, fill = platform, linetype = platform)
) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::geom_point(size = 2) +
  ggplot2::labs(
    title = "fit maturity curve",
    subtitle = "Mean proportion mature-at-age from FIMS vs SS3",
    x = "Age",
    y = "Mean mature-at-age"
  ) +
  ggplot2::theme_bw()

```

```{r}
#| label: cleanup
FIMS::clear()
```
