---
title: SWFSC Case Study Pacific mackerel
stock: "SWFSC mackerel"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Pacific mackerel"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: SWFSC
* Analyst: Alexander Jensen

## Simplifications to the original assessment

[SUMMARY TEXT OF DIFFERENCES FROM OPERATIONAL MODEL. 
E.g., The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* Convert two-semester time step to annual
* Sum the catch values for MexCal S1 and S2
* Sum the age comps for MexCal S1 and S2
* Drop PNW fishing fleet data (catch and age comp data)
* Drop spring AT survey index values
* Use expected summary biomass values as AT survey CPUE
* Fix Q=1 for AT survey, rather than have fixed, year-specific values
* No SR_regime parameter estimated (population assumed to start a equilibrium)
* Fixed M at 0.8
* Assume time-invariant logistic selectivity AT survey fleets (time-varying age-0 in benchmark assessment)
* Assume time-invariant logistic selectivity MexCal fleet (time-varying in benchmark assessment)
* Lorenzen age M -> one M value
* No ageing error in FIMS?
* Assume time-invariant weight at age for the fishery and survey
]

## Setting up the data

```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false
load(file.path(data_directory, "sardine_simplified_res.Rdata"))

FIMS::clear()

#--------------------------------------------------------
#Logistic function for later use
logistic <- function(x, slope, inflection_point){
  out <- 1 / (1 + exp(-1 * slope * (x - inflection_point)))
  out <- data.frame(x = x, value = out)
  return(out)
}

#--------------------------------------------------------
# Import data copied+formatted from SS3 model files
load("C:/Users/alex.jensen/Documents/GitHub/Mackerel-SA-2027_Update/Data/FIMS inputs/bench_data.RData")

# Convert to FIMSFrame
data_4_model <- FIMS::FIMSFrame(fims_data)

n_years <- FIMS::get_n_years(data_4_model)
years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(data_4_model)
n_ages <- FIMS::get_n_ages(data_4_model)
ages <- FIMS::get_ages(data_4_model)

# Not sure what to do with this
rzero <- 13.73
M_value <- 0.851 #.8 worked pretty well
init_naa <- exp(rzero) * exp(-(ages - 1) * M_value)
init_naa[n_ages] <- init_naa[n_ages] / M_value # sum of infinite series

# Create default recruitment, growth, and maturity parameters
parameters <- FIMS::setup_default_parameters(data = data_4_model) 

# Move to age-specific selectivity
new_survey_sel <- FIMS::setup_default_Selectivity(
  data = data_4_model,
  fleet = "survey1",
  module_type = "AgeSpecific"
)
new_fish_sel <- FIMS::setup_default_Selectivity(
  data = data_4_model,
  fleet = "fleet1",
  module_type = "AgeSpecific"
)
print(new_survey_sel)

update_parameters <- parameters |>
  dplyr::filter(module_name != "Selectivity") |>
  dplyr::bind_rows(new_survey_sel, new_fish_sel)

# Look at updated default selectivity parameters
update_parameters |>
  dplyr::filter(module_name == "Selectivity") |>
  dplyr::select(module_type, fleet, label, age, value, estimation_type) |>
  print()

# Update other parameter specifications
  # Q will be the most difficult to specify
  # In benchmark, separate base values values are estimated for 2008-2015 and 2016-2021,
    # each with the same informative prior
    # Additionally, we model annual deviations around the base value for 2008-2015
update_parameters <- update_parameters |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "logit_steep",
      value = FIMS::logit(0, 1.0, 0.75), # default of logit-transformed 0.75
      estimation_type = "constant" # fix steepness
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_rzero", "log_sd"),
      value = c(rzero, log(0.75)), # sigmaR = 0.75 
      estimation_type = c("fixed_effects", "constant") # estimate log_rzero, fix log_sd
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"), # already fixed to constant
      value = c(1.01, 1.96) # 1.01 reported in Kuriyama et al. 2023; 1.96 deduced from other data
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet = "survey1",
      label = "log_q",
      value = -1.401, # this is the log of the mean of the base Q values for the two time blocks
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  # NEED TO UPDATE SELECTIVITY PARAMETERIZATION TO ~MATCH BENCHMARK
  dplyr::rows_update( # NOT SURE WHAT TO DO HERE
    ssres[["timeseries"]] |>
        dplyr::select(timing = Yr, value = "F:_1") |>
        dplyr::filter(
          timing >= get_start_year(data_4_model),
          timing <= get_end_year(data_4_model)
        ) |>
        dplyr::mutate(
          module_name = "Fleet",
          fleet = "fleet1",
          label = "log_Fmort",
          estimation_type = "constant",
          value = log(value)
        ),
    by = c("module_name", "fleet", "label", "timing")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_M"),
      value = log(M_value)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_init_naa"),
      age = ages,
      value = log(init_naa),
      estimation_type = "constant"
    ),
    by = c("module_name", "age", "label")
  )
```

## Run FIMS model

```{r}
#| label: setup-model

# Fit the model
fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE)
output <- FIMS::get_estimates(fit) |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + get_start_year(data_4_model) - 1,
    estimate = estimated
  )
dplyr::filter(output, label == "log_q") |> as.data.frame()
dplyr::filter(output, label == "spawning_biomass") |> as.data.frame()
dplyr::filter(output, label == "index_expected") |> as.data.frame()
get_data(data_4_model)|>dplyr::filter(type == "catch")
# Get information about the model and print a few characters to the screen
# log <- FIMS::get_log_module("information")
```

## Add your comparison figures

```{r}
#| label: comparison-plots


#------------------------------------------------------------------------
#------SSB
# Q: Is 'ssb' from  get_report(fit) calculated assuming 2 genders and a 50:50 ratio?
  # If so, it might explain the discrepancy between ss3 (1 gender model) and fims (2 genders)
ssbs <- ssres$timeseries |>
  dplyr::select(Yr, ss3 = SpawnBio) |>
  dplyr::full_join(
    dplyr::filter(output, label == "spawning_biomass") |>
      dplyr::select(year, fims = estimate),
    by = c("Yr" = "year")
  )

ssbs |>
  dplyr::filter(Yr >= 2005, Yr < 2024) |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Spawning biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::xlab("year") +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave(
  "figures/SWFSC-sardine-sb.png",
  width = 7.35,
  height = 4.8
)

# Fishing mortality
ssres$timeseries |>
  dplyr::select(Yr, ss3 = "F:_1") |>
  dplyr::full_join(
    dplyr::filter(output, label == "log_Fmort", module_id == 1) |>
      dplyr::mutate(estimate = exp(estimate)) |>
      dplyr::select(year, fims = estimate),
    by = c("Yr" = "year")
  ) |>
  dplyr::filter(Yr >= 2005, Yr < 2024) |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Fishing mortality") +
  ggplot2::theme_bw() +
  ggplot2::xlab("year") +
  ggplot2::theme(legend.position.inside = c(.9, .9))

#------------------------------------------------------------------------
#------Index fits
index <- ssres$cpue |>
  dplyr::select(year = Yr, obs = Obs, ss3 = Exp)
index$fims <- FIMS::get_report(fit)[["index_expected"]][[2]]
index |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = year)) +
  ggplot2::geom_point(ggplot2::aes(y = obs)) +
  ggplot2::geom_line(ggplot2::aes(y = value, color = variable, group = variable)) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Survey biomass value (mt)")
ggplot2::ggsave("figures/SWFSC-sardine-surveyfit.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#-----Calculate age-1+ biomass
#Multiply numbers at age by weight at age and sum
naa <- get_report(fit)[["numbers_at_age"]][[1]]

naa1 <- tidyr::crossing(c(years, 2024), ages) |>
  dplyr::mutate(naa = naa) |>
  as.data.frame()
names(naa1) <- c("year", "age", "value")
naa1$cohort <- naa1$year - naa1$age

#Format Weight at age (identical WAA for each year and fleet)
waa <- data.frame(
  age = wtatage$age[(wtatage$year == 2005 & wtatage$fleet == "fleet1")],
  waa = wtatage$value[(wtatage$year == 2005 & wtatage$fleet == "fleet1")]
) |>
  dplyr::mutate(age = as.integer(age))

naa1 <- naa1 |>
  dplyr::left_join(waa, by = "age")


naa1 <- naa1 |>
  dplyr::mutate(weight = value * waa)
age1plus <- naa1 |>
  dplyr::filter(age != 0) |>
  dplyr::group_by(year) |>
  dplyr::summarize(summbio = sum(weight))

bio1 <- ssres$timeseries |>
  dplyr::filter(Seas == 1) |>
  dplyr::select(Yr, Bio_smry) |>
  dplyr::mutate(year = Yr, ss3bio = Bio_smry) |>
  dplyr::select(-Yr, -Bio_smry)


age1plus <- age1plus |>
  dplyr::left_join(bio1, by = "year")

names(age1plus) <- c("year", "fims", "ss3")

#Full time series of age-1+ biomass
age1plus |>
  reshape2::melt(id.var = "year") |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave("figures/SWFSC-sardine-age1plusbio.png", width = 6.8, height = 5.25)

#Zoomed in time series of age-1+
age1plus |>
  reshape2::melt(id.var = "year") |>
  dplyr::filter(year >= 2010) |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme(legend.position.inside = c(.9, .9))
ggplot2::ggsave(
  "figures/SWFSC-sardine-age1plusbio_zoomedin.png",
  width = 6.8,
  height = 5.25
)


#------------------------------------------------------------------------
#------Recruitment
# Divide FIMS recruits by 1000 to match reporting in SS3
recs <- ssres$timeseries |>
  dplyr::select(Yr, ss3 = Recruit_0) |>
  dplyr::mutate(fims = c(0, 0, get_report(fit)[["expected_recruitment"]][[1]]) / 1000)

recs |>
  dplyr::filter(Yr >= 2005, Yr < 2024) |>
  reshape2::melt(id.var = "Yr") |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::theme_bw() +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::ylab("Recruits (x1000)")
ggplot2::ggsave("figures/SWFSC-sardine-recruitment.png", width = 6.8, height = 5.25)


#------------------------------------------------------------------------
#------Fixed selectivities
#Are fixed but plot for comparison's sake
##Fishery

### commented out due to error "! object 'value' not found"
# sel_fishery <- logistic(ages,
#                         slope = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "slope", module_id == 1) |>
#                           dplyr::pull(estimated),
#                         inflection_point = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "inflection_point", module_id == 1) |>
#                           dplyr::pull(estimated))

# names(sel_fishery) <- c("age", "fims")

# sel_fishery$ss3 <- ssres$ageselex |>
#   dplyr::filter(Yr == 2005, Factor == "Asel", Fleet == 1) |>
#   dplyr::select(as.character(0:8)) |>
#   t()
# sel_fishery <- sel_fishery |>
#   reshape2::melt(id.var = "age")

# ggplot2::ggplot(
#   sel_fishery,
#   ggplot2::aes(x = age, y = value, group = variable, color = variable)
# ) +
#   ggplot2::geom_point() +
#   ggplot2::geom_line()

#-----------Survey
sel_survey <- logistic(ages,
                       slope = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "slope", module_id == 2) |>
                          dplyr::pull(estimated),
                       inflection_point = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "inflection_point", module_id == 2) |>
                          dplyr::pull(estimated))

names(sel_survey) <- c("age", 'fims')

sel_survey$ss3 <- ssres$ageselex |>
  dplyr::filter(Yr == 2005, Factor == "Asel", Fleet == 2) |>
  dplyr::select(as.character(0:8)) |>
  t()
sel_survey <- sel_survey |>
  reshape2::melt(id.var = "age")

ggplot2::ggplot(
  sel_survey,
  ggplot2::aes(x = age, y = value, group = variable, color = variable)
) +
  ggplot2::geom_point() +
  ggplot2::geom_line()

```

## Add comparison tables

## What was your experience using FIMS? What could we do to improve usability?

Tools to check:
* data were inputted correctly (dimension checks)
* starting values and settings are reasonable
* I had issues installing FIMS with install_github that seemed to be related to R settings

Model:
* streamline configuration of fleets and maybe make it easier to add additional fleets (perhaps clone existing ones then change specific settings)
* Think about ways to modify single model settings based on say different starting values

Output:
* include all parameter values and names (fixed and estimated, also show starting values)
* Code to generate a default set of figures and tables for use in assessment documents/presentations

## What features are most important to add based on this case study?

* The ability to start the model at non-equilibrium conditions  (In SS3 there is a
SR_regime parameter and Initial F to match equilbirium age comps to first year of data age comps)
*Does the model year start in the first year of the data input?

```{r}
#| label: clear
#| output: false
#| warning: false

FIMS::clear()
```
