---
title: SWFSC Case Study Pacific mackerel
stock: "SWFSC mackerel"
previous_model: "SS3"
features: ""
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Pacific mackerel"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: SWFSC
* Analyst: Alexander Jensen

## Simplifications to the original assessment

[SUMMARY TEXT OF DIFFERENCES FROM OPERATIONAL MODEL. 
E.g., The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* Convert two-semester time step to annual
* Sum the catch values for MexCal S1 and S2
* Sum the age comps for MexCal S1 and S2
* Drop PNW fishing fleet data (catch and age comp data)
* Drop spring AT survey index values
* Use expected summary biomass values as AT survey CPUE
* Fix Q=1 for AT survey, rather than have fixed, year-specific values
* No SR_regime parameter estimated (population assumed to start a equilibrium)
* Fixed M at 0.8
* Assume time-invariant logistic selectivity AT survey fleets (time-varying age-0 in benchmark assessment)
* Assume time-invariant logistic selectivity MexCal fleet (time-varying in benchmark assessment)
* Lorenzen age M -> one M value
* No ageing error in FIMS?
* Assume time-invariant weight at age for the fishery and survey
]

## Setting up the data

```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false

FIMS::clear()

#--------------------------------------------------------
# Import data copied+formatted from SS3 model files (fims_data)
load("C:/Users/alex.jensen/Documents/GitHub/Mackerel-SA-2027_Update/Data/FIMS inputs/bench_data.RData")

# Import SS3 output from benchmark model fit
modfit <- readRDS("C:/Users/alex.jensen/Documents/GitHub/Mackerel-SA-2027_Update/Models/2023 Benchmark_v3.30.20/2023 Benchmark_v3.30.20.rds")

# Convert to FIMSFrame
data_4_model <- FIMS::FIMSFrame(fims_data)

n_years <- FIMS::get_n_years(data_4_model)
years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(data_4_model)
n_ages <- FIMS::get_n_ages(data_4_model)
ages <- FIMS::get_ages(data_4_model)

# Specify initial/fixed values based on 2023 benchmark
rzero <- modfit$parameters |>
  dplyr::filter(Label == "SR_LN(R0)") |>
  dplyr::pull(Value)
rzero <- rzero + log(1000)
M_value <- modfit$parameters |>
  dplyr::filter(Label == "NatM_Lorenzen_averageFem_GP_1") |>
  dplyr::pull(Value)
init_naa <- exp(rzero) * exp(-(ages - 1) * M_value)
init_naa[n_ages] <- init_naa[n_ages] / M_value # sum of infinite series
steep <- modfit$parameters |>
  dplyr::filter(Label == "SR_BH_steep") |>
  dplyr::pull(Value)
sigmaR <- modfit$parameters |>
  dplyr::filter(Label == "SR_sigmaR") |>
  dplyr::pull(Value)
logQ <- modfit$parameters |>
  dplyr::filter(Label %in% c("LnQ_base_AT(2)_BLK4repl_2016", "LnQ_base_AT(2)")) |>
  dplyr::mutate(Q = exp(Value)) |>
  dplyr::summarize(meanQ = mean(Q)) |>
  dplyr::mutate(logmeanQ = log(meanQ)) |>
  dplyr::pull(logmeanQ)
sel_survey_init <- modfit$ageselex |>
  dplyr::filter(Label == "2008_2Asel") |>
  tidyr::pivot_longer(cols = `0`:`8`, names_to = "age", values_to = "sel") |>
  dplyr::pull(sel)
sel_survey_init <- ifelse(sel_survey_init == 1, 0.9999, sel_survey_init); print(sel_survey_init)
sel_fish_init <- modfit$ageselex |>
  dplyr::filter(Factor == "Asel", Fleet == 1, Yr %in% years) |>
  tidyr::pivot_longer(cols = `0`:`8`, names_to = "age", values_to = "sel") |>
  dplyr::group_by(age) |>
  dplyr::summarize(avg_sel = mean(sel)) |>
  dplyr::pull(avg_sel); print(sel_fish) 
  # values for ages 1-3 are high similar and greater than 1 (~1.1)
  # not sure why some year-, age-specific selectivity values > 1
sel_fish_init <- ifelse(sel_fish_init > 1, 0.9999, sel_fish_init); print(sel_fish_init)

# Create default recruitment, growth, and maturity parameters
parameters <- FIMS::setup_default_parameters(data = data_4_model) 

# Move to age-specific selectivity
new_survey_sel <- FIMS::setup_default_Selectivity(
  data = data_4_model,
  fleet = "survey1",
  module_type = "AgeSpecific"
)
new_fish_sel <- FIMS::setup_default_Selectivity(
  data = data_4_model,
  fleet = "fleet1",
  module_type = "AgeSpecific"
)
print(new_survey_sel); print(new_fish_sel)
update_parameters <- parameters |>
  dplyr::filter(module_name != "Selectivity") |>
  dplyr::bind_rows(new_survey_sel, new_fish_sel)

# Look at updated default selectivity parameters
update_parameters |>
  dplyr::filter(module_name == "Selectivity") |>
  dplyr::select(module_type, fleet, label, age, value, estimation_type) |>
  print()

# Update other parameter specifications
  # Q will be the most difficult to specify
  # In benchmark, separate base values values are estimated for 2008-2015 and 2016-2021,
    # each with the same informative prior
    # Additionally, we model annual deviations around the base value for 2008-2015
update_parameters <- update_parameters |>
  dplyr::rows_update( # fix h to logit-transformed value specified in benchmark
    tibble::tibble(
      module_name = "Recruitment",
      label = "logit_steep",
      value = FIMS::logit(0, 1.0, steep), # default of logit-transformed 0.75
      estimation_type = "constant" 
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( # change log_devs estimation_type 
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_devs"),
      estimation_type = c("random_effects") # estimate log_rzero, fix log_sd
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( # fix sigmaR to benchmark-specified value; provided benchmark-estimated initial value for R0
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_rzero", "log_sd"),
      value = c(rzero, log(sigmaR)),
      estimation_type = c("fixed_effects", "fixed_effects") # estimate log_rzero, log_sd (for REs)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( # fix maturity curve parameters to benchmark-estimated values
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"), # already fixed to constant
      value = c(1.01, 1.96) # 1.01 reported in Kuriyama et al. 2023; 1.96 deduced from other data
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( # Provide initial value for estimated single Q (no prior)
    tibble::tibble(
      module_name = "Fleet",
      fleet = "survey1",
      label = "log_q",
      value = logQ, # this is the log of the mean of the base Q values for the two time blocks
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet", "label")
  ) |>
  dplyr::rows_update( # Specify survey1 age-specific selectivity parameters (est. 1 param - time-invariant)
    tibble::tibble(
      fleet = "survey1",
      label = "logit_sel_at_age",
      age = ages,
      value = boot::logit(sel_survey_init),
      estimation_type = c("fixed_effects", rep("constant", 8))
    ),
    by = c("fleet", "label", "age")
  ) |>
  dplyr::rows_update( # Specify fleet1 age-specific selectivity parameters (est. 1 param - time-invariant)
    tibble::tibble(
      fleet = "fleet1",
      label = "logit_sel_at_age",
      age = ages, 
      value = boot::logit(sel_fish_init),
      estimation_type = c("fixed_effects", rep("constant", 8))
       # benchmark estimates 4 parameters for ages 0-3
       # for age-specific selectivity, need to fix some age-specific selectivity=1
       # it just so happens that ages 1-3 tend to be similar and highest, so I fix to 1
    ),
    by = c("fleet", "label", "age")
  ) |>
  dplyr::rows_update( # fix Fmort to constant values based on SS3 model fit
    modfit$timeseries |>
        dplyr::select(timing = Yr, value = "F:_1") |>
        dplyr::filter(
          timing >= get_start_year(data_4_model),
          timing <= get_end_year(data_4_model)
        ) |>
        dplyr::mutate(
          module_name = "Fleet",
          fleet = "fleet1",
          label = "log_Fmort",
          estimation_type = "constant",
          value = log(value)
        ),
    by = c("module_name", "fleet", "label", "timing")
  ) |>
  dplyr::rows_update( # fix M to benchmark-estimated value
    tibble::tibble(
      module_name = "Population",
      label = c("log_M"),
      value = log(M_value),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update( # set initial naa (estimated via 'fixed_effects') to values calculated from benchmark R0, M
    tibble::tibble(
      module_name = "Population",
      label = c("log_init_naa"),
      age = ages,
      value = log(init_naa),
      estimation_type = "constant"
    ),
    by = c("module_name", "age", "label")
  )
```

## Run FIMS model

```{r}
#| label: setup-model

# Fit the model
clear()
fit <- update_parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE)
output <- FIMS::get_estimates(fit) |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i + get_start_year(data_4_model) - 1,
    estimate = estimated
  )
fit@sdreport$pdHess
output |> dplyr::filter(module_name == "Selectivity", module_type == "AgeSpecific") |> print()
dplyr::filter(output, label == "log_q") |> as.data.frame() |> print()
dplyr::filter(output, label == "log_sd") |> as.data.frame() |> print() # why is log_sd not included in model output?
dplyr::filter(output, label == "log_rzero") |> as.data.frame() |> print()
dplyr::filter(output, label == "spawning_biomass") |> as.data.frame()

dplyr::filter(output, label == "spawning_biomass") |> as.data.frame()
dplyr::filter(output, label == "index_expected") |> as.data.frame()
get_data(data_4_model)|>dplyr::filter(type == "catch")
# Get information about the model and print a few characters to the screen
# log <- FIMS::get_log_module("information")
```

## Add your comparison figures

```{r}
#| label: comparison-plots


#------------------------------------------------------------------------
#------SSB
# Q: Is 'ssb' from  get_report(fit) calculated assuming 2 genders and a 50:50 ratio?
  # If so, it might explain the discrepancy between ss3 (1 gender model) and fims (2 genders)
ssbs <- modfit$timeseries |>
  dplyr::select(Yr, ss3 = SpawnBio) |>
  dplyr::full_join(
    dplyr::filter(output, label == "spawning_biomass") |>
      dplyr::select(year, fims = estimate),
    by = c("Yr" = "year")
  )

ssbs |>
  dplyr::filter(Yr >= 2008, Yr <= 2021) |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Spawning biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::xlab("year") +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave(
  "figures/SWFSC-mackerel-sb.png",
  width = 7.35,
  height = 4.8
)

#------------------------------------------------------------------------
#-----Calculate age-1+ biomass
#Multiply numbers at age by weight at age and sum
naa <- get_report(fit)[["numbers_at_age"]][[1]]

naa1 <- tidyr::crossing(c(years, 2022), ages) |> # add one year for forecast age-1+ biomass
  dplyr::mutate(naa = naa) |>
  as.data.frame()
names(naa1) <- c("year", "age", "value")
naa1$cohort <- naa1$year - naa1$age

#Format Weight at age (identical WAA for each year and fleet)
waa <- data.frame(
  age = fims_data$age[(fims_data$timing == 2008 & fims_data$fleet == "fleet1" & fims_data$type == "weight_at_age")],
  waa = fims_data$observed[(fims_data$timing == 2008 & fims_data$fleet == "fleet1" & fims_data$type == "weight_at_age")]
) |>
  dplyr::mutate(age = as.integer(age))

naa1 <- naa1 |>
  dplyr::left_join(waa, by = "age")

naa1 <- naa1 |>
  dplyr::mutate(weight = value * waa)
age1plus <- naa1 |>
  dplyr::filter(age != 0) |>
  dplyr::group_by(year) |>
  dplyr::summarize(summbio = sum(weight))

bio1 <- modfit$timeseries |>
  dplyr::select(Yr, Bio_smry) |>
  dplyr::mutate(year = Yr, ss3bio = Bio_smry) |>
  dplyr::select(-Yr, -Bio_smry)


age1plus <- age1plus |>
  dplyr::left_join(bio1, by = "year")

names(age1plus) <- c("year", "fims", "ss3")

#Full time series of age-1+ biomass
age1plus |>
  reshape2::melt(id.var = "year") |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave("figures/SWFSC-mackerel-age1plusbio.png", width = 6.8, height = 5.25)

#Zoomed in time series of age-1+
age1plus |>
  reshape2::melt(id.var = "year") |>
  dplyr::filter(year >= 2010) |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme(legend.position.inside = c(.9, .9))
ggplot2::ggsave(
  "figures/SWFSC-mackerel-age1plusbio_zoomedin.png",
  width = 6.8,
  height = 5.25
)

#------------------------------------------------------------------------
#------Recruitment
# Divide FIMS recruits by 1000 to match reporting in SS3
recs <- modfit$timeseries |>
  dplyr::select(Yr, ss3 = Recruit_0) |>
  dplyr::filter(Yr <= 2022) |>
  dplyr::mutate(fims = c(0, 0, get_report(fit)[["expected_recruitment"]][[1]]) / 1000)

recs |>
  dplyr::filter(Yr >= 2008, Yr <= 2022) |>
  reshape2::melt(id.var = "Yr") |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::theme_bw() +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::ylab("Recruits (x1000)")
ggplot2::ggsave("figures/SWFSC-mackerel-recruitment.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#------Recruitment deviations (TBD)


#------------------------------------------------------------------------
#------Numbers-at-age
naa <- output |> # FIMS-estimated numbers at age for years 2008-2022
  dplyr::filter(label == "numbers_at_age") |>
  dplyr::select(year, age = age_i, fims = estimated) |>
  dplyr::mutate(age = age-1)
naa <- output |> # append FIMS-estimated init_naa for year 2007
  dplyr::filter(label == "log_init_naa") |>
  dplyr::mutate(age_i = age_i-1, 
                fims = exp(estimated),
                year = 2007) |>
  dplyr::select(year, age = age_i, fims) |>
  rbind(naa)
naa <- modfit$natage |> # append SS3-estimated naa for years 2007-2022
  dplyr::filter(Yr >= 2007, Yr <= 2022, `Beg/Mid` == "B") |>
  tidyr::pivot_longer(cols = `0`:`8`, names_to = "age", values_to = "ss3") |>
  dplyr::select(year = Yr, age, ss3) |>
  dplyr::mutate(ss3 = ss3 * 1000,
                age = as.numeric(age)) |>
  dplyr::left_join(naa) |>
  tidyr::pivot_longer(cols = fims:ss3, names_to = "type", values_to = "NAA")
ggplot2::ggplot(
  naa,
  ggplot2::aes(x = as.numeric(age), y = NAA, color = type, group = interaction(type))
) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::scale_linewidth_manual(values=c(1,0.25))+
  ggplot2::scale_linetype_manual(values=c(1,2))+
  ggplot2::facet_wrap(~year, scales = "free_y")
ggplot2::ggsave("figures/SWFSC-mackerel-naa.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#---------Age composition residuals (observed vs expected; TBD)

#------------------------------------------------------------------------
# --------Fishing mortality
modfit$timeseries |>
  dplyr::select(Yr, ss3 = "F:_1") |>
  dplyr::full_join(
    dplyr::filter(output, label == "log_Fmort", module_id == 2) |>
      dplyr::mutate(estimate = exp(estimate)) |>
      dplyr::select(year, fims = estimate),
    by = c("Yr" = "year")
  ) |>
  dplyr::filter(Yr >= 2008, Yr <= 2021) |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Fishing mortality") +
  ggplot2::theme_bw() +
  ggplot2::xlab("year") +
  ggplot2::theme(legend.position.inside = c(.9, .9))
ggplot2::ggsave("figures/SWFSC-mackerel-Fmort.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#------Index fits
all_yrs <- expand.grid(year = c(2008:2021))
index <- modfit$cpue |>
  dplyr::select(year = Yr, obs = Obs, ss3 = Exp)
index <- dplyr::left_join(all_yrs, index)
index$fims <- FIMS::get_report(fit)[["index_expected"]][[1]] # AJ: what's the difference between 1 and 2?
index |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = year)) +
  ggplot2::geom_point(ggplot2::aes(y = obs)) +
  ggplot2::geom_line(ggplot2::aes(y = value, color = variable, group = variable)) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Survey biomass value (mt)")
ggplot2::ggsave("figures/SWFSC-mackerel-surveyfit.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#------Fishery selectivity
sel_fishery <- output |> 
  dplyr::filter(module_name == "Selectivity", module_type == "AgeSpecific", module_id == 2) |> 
  dplyr::mutate(age = age_i - 1,
                fims = boot::inv.logit(estimated)) |>
  dplyr::select(age, fims)
sel_fishery <- modfit$ageselex |>
  dplyr::filter(Yr >= 2008, Yr <= 2021, Fleet == 1, Factor == "Asel") |>
  # dplyr::select(Yr, `0`:`8`) |>
  tidyr::pivot_longer(cols = `0`:`8`, names_to = "age", values_to = "ss3") |>
  dplyr::mutate(age = as.numeric(age)) |>
  dplyr::select(Yr, age, ss3) |>
  dplyr::left_join(sel_fishery) |>
  tidyr::pivot_longer(cols = fims:ss3, names_to = "type", values_to = "Selectivity")
ggplot2::ggplot(
  sel_fishery,
  ggplot2::aes(x = as.numeric(age), y = Selectivity, color = type, group = interaction(type, Yr))
) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::scale_linewidth_manual(values=c(1,0.25))+
  ggplot2::scale_linetype_manual(values=c(1,2))
ggplot2::ggsave("figures/SWFSC-mackerel-fishselectivity.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#-----------Survey selectivity
sel_survey <- output |> 
  dplyr::filter(module_name == "Selectivity", module_type == "AgeSpecific", module_id == 1) |> 
  dplyr::mutate(age = age_i - 1,
                fims = boot::inv.logit(estimated)) |>
  dplyr::select(age, fims)
sel_survey <- modfit$ageselex |>
  dplyr::filter(Yr >= 2008, Yr <= 2021, Fleet == 2, Factor == "Asel") |>
  # dplyr::select(Yr, `0`:`8`) |>
  tidyr::pivot_longer(cols = `0`:`8`, names_to = "age", values_to = "ss3") |>
  dplyr::mutate(age = as.numeric(age)) |>
  dplyr::select(Yr, age, ss3) |>
  dplyr::left_join(sel_survey) |>
  tidyr::pivot_longer(cols = fims:ss3, names_to = "type", values_to = "Selectivity")
ggplot2::ggplot(
  sel_survey,
  ggplot2::aes(x = as.numeric(age), y = Selectivity, color = type, group = interaction(type, Yr))
) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::scale_linewidth_manual(values=c(1,0.25))+
  ggplot2::scale_linetype_manual(values=c(1,2))
ggplot2::ggsave("figures/SWFSC-mackerel-survselectivity.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#-----------Survey catchability
surv_q <- output |> 
  dplyr::filter(label == "log_q", module_id == 1) |>
  dplyr::mutate(fims = exp(estimate)) |>
  dplyr::select(fims)
surv_q <- modfit$cpue |>
  dplyr::select(Year = Yr, ss3 = Calc_Q) |>
  cbind(surv_q) |>
  tidyr::pivot_longer(cols = ss3:fims, values_to = "Catchability", names_to = "type")

surv_q |>
  dplyr::filter(Year >= 2008, Year <= 2022) |>
  ggplot2::ggplot(ggplot2::aes(x = Year, y = Catchability, group = type, color = type)) +
  ggplot2::theme_bw() +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme(legend.position.inside = c(.9, .9))
ggplot2::ggsave("figures/SWFSC-mackerel-survcatchability.png", width = 6.8, height = 5.25)
```

## Add comparison tables

```{r}
#| label: comparison-plots


#------------------------------------------------------------------------
#------Summary table of comparable parameter estimates, recent summary biomass

# Need to run all plots first, and particularly that for age-1+ summary biomass

#ss3
parm_ss3 <- data.frame(Type = "Parameters", # ss3 parameters
                         Component = c("NatM_Lorenzen_averageFem_GP_1", "SR_LN(R0)", "SR_BH_steep", "SR_sigmaR"),
                         ss3=round(modfit$parameters$Value[which(rownames(modfit$parameters) %in% c("NatM_Lorenzen_averageFem_GP_1", "SR_LN(R0)", "SR_BH_steep", "SR_sigmaR"))], 3))
parm_ss3$ss3[parm_ss3$Component == "SR_LN(R0)"] <- parm_ss3$ss3[parm_ss3$Component == "SR_LN(R0)"] + log(1000) # adjust equil. recr. to match fims units
bio_ss3 <- data.frame(Type = "Summary biomass",
                      Component = c("2020", "2021", "2022"),
                      ss3 = age1plus |> dplyr::filter(year %in% c(2020, 2021, 2022)) |> dplyr::pull(ss3))
ss3_smry <- rbind(parm_ss3, bio_ss3)

#fims
fims_M <- output |> dplyr::filter(label == "log_M", age_i == 1, year_i == 1) |> dplyr::pull(estimate) |> exp() |> round(3)
fims_lnR0 <- output |> dplyr::filter(label == "log_rzero") |> dplyr::pull(estimate) |> round(3)
fims_steep <- output |> dplyr::filter(label == "logit_steep") |> dplyr::pull(estimate) |> boot::inv.logit() |> round(3)
fims_sigmaR <- update_parameters |> dplyr::filter(label == "log_sd") |> dplyr::pull(value) |> exp() |> round(3)
parm_fims <- data.frame(Type = "Parameters",
                        Component = c("M", "ln(R0)", "Steepness", "SigmaR"),
                        fims = c(fims_M, fims_lnR0, fims_steep, fims_sigmaR))
bio_fims <- data.frame(Type = "Summary biomass",
                        Component = c("2020", "2021", "2022"),
                        fims = age1plus |> dplyr::filter(year %in% c(2020, 2021, 2022)) |> dplyr::pull(fims))
fims_smry <- rbind(parm_fims, bio_fims)

smry <- cbind(fims_smry, ss3_smry |> dplyr::select(ss3))

knitr::kable(smry)

```

## What was your experience using FIMS? What could we do to improve usability?

Documentation:
* Units for NAA and recruitment would be useful to know. 

Tools to check:
* 

Model:
* 

Output:
* Include estimated SRR log_sd (when log_devs are estimated using random_effects)

## Model fitting takeaways

* Model fitting benefits greatly with estimation of log_init_naa
  * This is closest analogue to estimating early ss3 rec_devs for non-equilibrium start
* Difficulty in model convergence when estimating multiple paramters that affect scaling
  * E.g., log_rzero, log_init_naa, and log_q
  * E.g., log_M, log_rzero, and log_q
  * E.g., log_M, log_init_naa, and log_q

## What other model configurations were explored that may not have worked well?

* Estimating log_devs using 'fixed_effects' with a constant log_sd did not converge well
* Model failed to converge when both 'log_init_naa' and 'log_rzero' were estimated with 'fixed_effects'
* Compared alternative methods of scaling 'fims' biomass: 
  * 1. Estimate log_rzero, fix log_init_naa
  * 2. Fix log_rzero, estimate log_init_naa
  * 2 has the following relative 1:
    * Improved fit to smrybio, recruitment, NAA, index, fishery selectivity 
    * Worsened fit to SSB
    * Similar-enough estimates of survey Q, selectivity
  * Both models overestimated biomass early and late in the time series due to time-invariant Q

## What features are most important to add based on this case study?

* Time-varying age-based selectivity (for fleet1)
* Time-varying catchability (for survey1)
* Prior specification (e.g., informed estimation of catchability, M)

```{r}
#| label: clear
#| output: false
#| warning: false

FIMS::clear()
```
