---
title: SWFSC Case Study Pacific Sardine
format:
  html:
    code-fold: true
---

## The setup

{{< include setup.qmd >}}

```{r}
#| label: setup-objects
common_name <- "Northern subpopulation of Pacific sardine"
```

* R version: `r R_version`
* TMB version: `r TMB_version`
* FIMS commit: `r FIMS_commit`
* Stock name: `r common_name`
* Region: SWFSC
* Analyst: Peter Kuriyama

## Simplifications to the original assessment

The model presented in this case study was changed substantially from the operational version and should not be considered reflective of the `r common_name` stock. These results are intended to be a demonstration and nothing more.

To get the operational model to more closely match a FIMS model the following changes were made:

* Convert two-semester time step to annual
* Sum the catch values for MexCal S1 and S2
* Sum the age comps for MexCal S1 and S2
* Drop PNW fishing fleet data (catch and age comp data)
* Drop spring AT survey index values
* Use expected summary biomass values as AT survey CPUE
* Fix Q=1 for AT survey, rather than have fixed, year-specific values
* No SR_regime parameter estimated (population assumed to start a equilibrium)
* Fixed M at 0.8
* Assume time-invariant logistic selectivity AT survey fleets (time-varying age-0 in benchmark assessment)
* Assume time-invariant logistic selectivity MexCal fleet (time-varying in benchmark assessment)
* Lorenzen age M -> one M value
* No ageing error in FIMS?
* Assume time-invariant weight at age for the fishery and survey

## Setting up the data

```{r}
#| warning: false
#| label: prepare-fims-data
#| output: false
FIMS::clear()

#--------------------------------------------------------
#Logistic function for later use
logistic <- function(x, slope, inflection_point){
  out <- 1 / (1 + exp(-1 * slope * (x - inflection_point)))
  out <- data.frame(x = x, value = out)
  return(out)
}

#--------------------------------------------------------
#Manually enter data

# setwd("C://Users//peter.kuriyama//SynologyDrive/Research//noaa//FIMS")

#-----Catch
catch <- data.frame(year = 2005:2023, catch = c(29188.50, 53107.00, 69929.40,
                                                56317.80, 33546.40, 17466.40, 39383.10, 2585.38, 5705.77, 2558.63, 7.18, 428.26,
                                                347.11, 514.20, 619.04, 653.15, 285.89, 508.02, 152.31))

# ggplot(catch, aes(x = year, y = catch)) + geom_point() +
#   geom_line() + scale_y_continuous(label = comma)


fimscatch <- tibble::tibble(type = "landings", name = "fleet1",
                    age = NA, timing = catch$year, value = catch$catch,
                    unit = "mt", uncertainty = 0.05)

#-----CPUE
cpue <- data.frame(year = 2005:2023, obs = c(649619.0, 899635.0, 956354.0, 863281.0, 652029.0,
                                             504970.0, 395783.0, 293980.0, 182417.0, 89260.1,
                                             46403.0, 40704.0, 44592.1, 48789.1, 53551.8,
                                             59765.8, 68451.7, 71612.5, 68957.9))


# ggplot(cpue, aes(x = year, y = obs)) + geom_point() + geom_line() +
#   scale_y_continuous(label = comma)

fimsindex <- tibble::tibble(type = "index", name = "survey1",
                    age = NA, timing = cpue$year,
                    value = cpue$obs, unit = "mt", uncertainty = .3)
fimsindex$unit <- ""

#-----Age compositions
acomps <- utils::read.csv(file.path(data_directory, "sardine_acomps.csv")) |>
  dplyr::mutate(value_prop = value / Nsamp) # convert age-comp data to proportions, to match fims-demo.Rmd

fimsage <- tibble::tibble(type = "age_comp", name = acomps$name,
                  age = acomps$age, timing = acomps$Yr,
                  value = acomps$value_prop, unit = "proportion", uncertainty = acomps$Nsamp)
#fimsage$uncertainty <- 50 Leave as empirical values

#-----Weight-at-age
  # Q: What are the differences in how SS3 and FIMS process wtatage inputs/units?
  # A: SS3 assumes WAA is expressed in kg, and then converts to mt for biomass calculations
      # SS3 also expresses natage in terms of 1000s/fish
    # FIMS assumes WAA is expressed in mt, but expresses natage in fish
    # WAA input values are supplied to SS3 in kg and FIMS in mt
    # Model now correctly processes inputs and provides comparable NAA, WAA, and biomass outputs to SS3
wtatage <- r4ss::SS_readwtatage(file.path(data_directory, "sardine_wtatage.ss_new")) |>
  dplyr::filter(fleet %in% c(1, 2)) |>
  dplyr::mutate(fleet = ifelse(fleet == 1, "fleet1", "survey1")) |>
  tidyr::pivot_longer(cols = `0`:`10`, names_to = "age", values_to = "value") |>
  dplyr::filter(year != 2024, !(age %in% c("9", "10"))) |> # Trim ages 9 and 10 to match age-comps
  dplyr::mutate(value = value / 1000) # WAA converted from kg to mt
fimswaa <- tibble::tibble(
  type = "weight-at-age",
  name = wtatage$fleet,
  age = wtatage$age,
  timing = wtatage$year,
  value = wtatage$value,
  unit = "mt",
  uncertainty = NA
)

#Combine everything, format
data_4_model <- rbind(fimscatch, fimsindex, fimsage, fimswaa)
data_4_model$age <- as.integer(data_4_model$age)
data_4_model$value <- as.numeric(data_4_model$value)

# Convert to FIMSFrame
data_4_model <- FIMS::FIMSFrame(data_4_model)

n_years <- FIMS::get_n_years(data_4_model)
years <- FIMS::get_start_year(data_4_model):FIMS::get_end_year(data_4_model)
n_ages <- FIMS::get_n_ages(data_4_model)
ages <- FIMS::get_ages(data_4_model)
rzero <- 17 #14.2 is log(R0) in sardine simplified model
M_value <- 0.8 #.8 worked pretty well
init_naa <- exp(rzero) * exp(-(ages - 1) * M_value)
init_naa[n_ages] <- init_naa[n_ages] / M_value # sum of infinite series

# Create default recruitment, growth, and maturity parameters
parameters <- FIMS::create_default_configurations(data = data_4_model) |>
  FIMS::create_default_parameters(data = data_4_model) |>
  tidyr::unnest(data) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Selectivity",
      label = rep(c("inflection_point", "slope"), 2),
      value = c(1, 5, 1.2, 2),
      estimation_type = "constant",
      fleet_name = rep(c("fleet1", "survey1"), each = 2)
    ),
    by = c("module_name", "label", "fleet_name")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "logit_steep",
      value = FIMS::logit(0.2, 1.0, 0.6)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = c("log_rzero", "log_sd"),
      value = c(rzero, log(1.2)),
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Recruitment",
      label = "log_devs",
      value = log(1),
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Maturity",
      label = c("inflection_point", "slope"),
      value = c(1.2, 1.5) # 1.5 is arbitrary guess
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "survey1",
      label = "log_q",
      value = 0,
      estimation_type = "constant"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Fleet",
      fleet_name = "fleet1",
      label = "log_Fmort",
      value = log(0.2)
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_M"),
      value = log(M_value)
    ),
    by = c("module_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Population",
      label = c("log_init_naa"),
      age = ages,
      value = log(init_naa),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "age", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Data",
      fleet_name = "fleet1",
      label = "log_sd",
      value = log(sqrt(log(0.01^2 + 1))),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  ) |>
  dplyr::rows_update(
    tibble::tibble(
      module_name = "Data",
      fleet_name = "survey1",
      label = "log_sd",
      value = log(sqrt(log(0.1^2 + 1))),
      estimation_type = "fixed_effects"
    ),
    by = c("module_name", "fleet_name", "label")
  )
```

## Run FIMS model

```{r}
#| label: setup-model

# Fit the model
fit <- parameters |>
  FIMS::initialize_fims(data = data_4_model) |>
  FIMS::fit_fims(optimize = TRUE)
output <- FIMS::get_estimates(fit) |>
  dplyr::mutate(
    uncertainty_label = "se",
    year = year_i,
    estimate = estimated
  )

# Get information about the model and print a few characters to the screen
log <- FIMS::get_log_module("information")
```

## Add your comparison figures

```{r}
#| label: comparison-plots
load(file.path(data_directory, "sardine_simplified_res.Rdata"))

#------------------------------------------------------------------------
#------SSB
# Q: Is 'ssb' from  get_report(fit) calculated assuming 2 genders and a 50:50 ratio?
  # If so, it might explain the discrepancy between ss3 (1 gender model) and fims (2 genders)
ssbs <- ssres$timeseries |>
  dplyr::select(Yr, ss3 = SpawnBio) |>
  dplyr::mutate(fims = c(0, 0, FIMS::get_report(fit)[["spawning_biomass"]][[1]]))

ssbs |>
  dplyr::filter(Yr >= 2005, Yr < 2024) |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Spawning biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::xlab("year") +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave(
  "figures/SWFSC-sardine-sb.png",
  width = 7.35,
  height = 4.8
)

#------------------------------------------------------------------------
#------Index fits
index <- ssres$cpue |>
  dplyr::select(Yr, Obs, Exp)
names(index) <- c("year", "obs", "ss3")
index$fims <- FIMS::get_report(fit)[["index_expected"]][[2]]
index |>
  tidyr::pivot_longer(names_to = "variable", cols = dplyr::matches("ss3|fims")) |>
  ggplot2::ggplot(ggplot2::aes(x = year)) +
  ggplot2::geom_point(ggplot2::aes(y = obs)) +
  ggplot2::geom_line(ggplot2::aes(y = value, color = variable, group = variable)) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Survey biomass value (mt)")
ggplot2::ggsave("figures/SWFSC-sardine-surveyfit.png", width = 6.8, height = 5.25)

#------------------------------------------------------------------------
#-----Calculate age-1+ biomass
#Multiply numbers at age by weight at age and sum
naa <- get_report(fit)[["numbers_at_age"]][[1]]

naa1 <- tidyr::crossing(c(years, 2024), ages) |>
  dplyr::mutate(naa = naa) |>
  as.data.frame()
names(naa1) <- c("year", "age", "value")
naa1$cohort <- naa1$year - naa1$age

#Format Weight at age (identical WAA for each year and fleet)
waa <- data.frame(
  age = wtatage$age[(wtatage$year == 2005 & wtatage$fleet == "fleet1")],
  waa = wtatage$value[(wtatage$year == 2005 & wtatage$fleet == "fleet1")]
) |>
  dplyr::mutate(age = as.integer(age))

naa1 <- naa1 |>
  dplyr::left_join(waa, by = "age")


naa1 <- naa1 |>
  dplyr::mutate(weight = value * waa)
age1plus <- naa1 |>
  dplyr::filter(age != 0) |>
  dplyr::group_by(year) |>
  dplyr::summarize(summbio = sum(weight))

bio1 <- ssres$timeseries |>
  dplyr::filter(Seas == 1) |>
  dplyr::select(Yr, Bio_smry) |>
  dplyr::mutate(year = Yr, ss3bio = Bio_smry) |>
  dplyr::select(-Yr, -Bio_smry)


age1plus <- age1plus |>
  dplyr::left_join(bio1, by = "year")

names(age1plus) <- c("year", "fims", "ss3")

#Full time series of age-1+ biomass
age1plus |>
  reshape2::melt(id.var = "year") |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position.inside = c(.9, .9))

ggplot2::ggsave("figures/SWFSC-sardine-age1plusbio.png", width = 6.8, height = 5.25)

#Zoomed in time series of age-1+
age1plus |>
  reshape2::melt(id.var = "year") |>
  dplyr::filter(year >= 2010) |>
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = variable, color = variable)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_bw() +
  ggplot2::ylab("Age-1+ biomass (mt)") +
  ggplot2::theme(legend.position.inside = c(.9, .9))
ggplot2::ggsave(
  "figures/SWFSC-sardine-age1plusbio_zoomedin.png",
  width = 6.8,
  height = 5.25
)


#------------------------------------------------------------------------
#------Recruitment
# Divide FIMS recruits by 1000 to match reporting in SS3
recs <- ssres$timeseries |>
  dplyr::select(Yr, Recruit_0) |>
  dplyr::mutate(fims = c(0, 0, get_report(fit)[["expected_recruitment"]][[1]]) / 1000)
names(recs)[2] <- "ss3"

recs |>
  dplyr::filter(Yr >= 2005, Yr < 2024) |>
  reshape2::melt(id.var = "Yr") |>
  ggplot2::ggplot(ggplot2::aes(x = Yr, y = value, group = variable, color = variable)) +
  ggplot2::theme_bw() +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme(legend.position.inside = c(.9, .9)) +
  ggplot2::ylab("Recruits (x1000)")
ggplot2::ggsave("figures/SWFSC-sardine-recruitment.png", width = 6.8, height = 5.25)


#------------------------------------------------------------------------
#------Fixed selectivities
#Are fixed but plot for comparison's sake
##Fishery

sel_fishery <- logistic(ages,
                        slope = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "slope", module_id == 1) |>
                          dplyr::pull(estimated),
                        inflection_point = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "inflection_point", module_id == 1) |>
                          dplyr::pull(estimated))

names(sel_fishery) <- c("age", "fims")

sel_fishery$ss3 <- ssres$ageselex |>
  dplyr::filter(Yr == 2005, Factor == "Asel", Fleet == 1) |>
  dplyr::select(as.character(0:8)) |>
  t()
sel_fishery <- sel_fishery |>
  reshape2::melt(id.var = "age")

ggplot2::ggplot(
  sel_fishery,
  ggplot2::aes(x = age, y = value, group = variable, color = variable)
) +
  ggplot2::geom_point() +
  ggplot2::geom_line()

#-----------Survey
sel_survey <- logistic(ages,
                       slope = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "slope", module_id == 2) |>
                          dplyr::pull(estimated),
                       inflection_point = dplyr::filter(FIMS::get_estimates(fit), module_name == "Selectivity", label == "inflection_point", module_id == 2) |>
                          dplyr::pull(estimated))

names(sel_survey) <- c("age", 'fims')

sel_survey$ss3 <- ssres$ageselex |>
  dplyr::filter(Yr == 2005, Factor == "Asel", Fleet == 2) |>
  dplyr::select(as.character(0:8)) |>
  t()
sel_survey <- sel_survey |>
  reshape2::melt(id.var = "age")

ggplot2::ggplot(
  sel_survey,
  ggplot2::aes(x = age, y = value, group = variable, color = variable)
) +
  ggplot2::geom_point() +
  ggplot2::geom_line()

```

## Add comparison tables

## What was your experience using FIMS? What could we do to improve usability?

Tools to check:
* data were inputted correctly (dimension checks)
* starting values and settings are reasonable
* Perhaps have a model template file that will work as is, then users can modify as necessary
* I had issues installing FIMS with install_github that seemed to be related to R settings

Model:
* streamline configuration of fleets and maybe make it easier to add additional fleets (perhaps clone existing ones then change specific settings)
* Think about ways to modify single model settings based on say different starting values

Output:
* include all parameter values and names (fixed and estimated, also show starting values)
* Code to generate a default set of figures and tables for use in assessment documents/presentations

## What features are most important to add based on this case study?

* The ability to start the model at non-equilibrium conditions  (In SS3 there is a
SR_regime parameter and Initial F to match equilbirium age comps to first year of data age comps)
*Does the model year start in the first year of the data input?

```{r}
#| label: clear
#| output: false
#| warning: false

FIMS::clear()
```
